### 2017-11-16 initial.tex
Initial Version LS Project Proposal

\documentclass{article}
\usepackage[]{float}
\usepackage[utf8]{inputenc}
\usepackage[]{hyperref}
\usepackage[hyperref=true,
    natbib=true,
    style=numeric,
    backend=bibtex]{biblatex}
\bibliography{references.bib}

\title{LS Project Proposal: TCP Congestion Algorithms In Datacenters}
\date{November 16, 2017}
\author{Luc Gommans, Rick van Gorp}

\begin{document}
\maketitle

\section{Introduction}
Multitentant datacenters aim to provide services with high availability and reliability. Some datacenters experience the issue that parts of their networks are sometimes congested, which might result in lower availability. Congestion occurs when a buffer of the receiving network device is full and the network device drops packets. TCP has a mechanism to avoid congestion in the network: TCP congestion control. Multiple algorithms are available for congestion control and show a different behaviour depending on the network's characteristics. In 2010 algorithm DCTCP was described in an article \cite{dctcp-congestion-original} and published as RFC8257 in 2017 \cite{dctcp-congestion}. DCTCP is optimized for datacenters and provides a high-burst tolerance, low latency, and high throughput when the datacenter has a small part of the buffer available \cite{dctcp-congestion}. In 2016 one algorithm was proposed: BBR. BBR is a TCP congestion algorithm created by Google that achieves higher bandwidths and lower latencies compared to other TCP congestion methods \cite{bbr-congestion}. A comparison of BBR with CUBIC \cite{bbr-congestion-comparison} shows that the BBR node pushes the CUBIC node away in bandwidth when using small buffers. The BBR node gets more bandwidth allocated than the CUBIC node.

We propose to do a performance and behaviour analysis of different TCP congestion control algorithms in a datacenter network. Based on the results we will recommend datacenters on measures to keep providing an available and reliable environment for customers when multiple congestion control algorithms active in the network.

\section{Research Questions}
Our main research question is: How can a datacenter provide an available and reliable environment for customers when multiple TCP congestion control algorithms are used in the network?
\\ \\
This main research question divides in five sub questions:
\begin{enumerate}
\item What TCP congestion algorithms are mostly implemented in datacenter environments?
\item How do the TCP congestion algorithms work?
\item How do the TCP congestion algorithms differ in behavior?
\item What is the impact on the availability and service for customers when using multiple TCP congestion algorithms?
\item What connection management measures are able to control the connection when using different congestion control algorithms?
\end{enumerate}

\section{Scope}
During this research we will focus on at least the TCP congestion control algorithms BBR, CUBIC, DCTCP. Additional congestion control algorithms will be determined after researching what the most implemented TCP congestion algorithms are in datacenter environments. We will conduct a behavior analysis and comparison between those algorithms, including the impact of mixed usage of the algorithms on the availability of a service in a datacenter. We will look at the measures a datacenter has to take to keep their service available and reliable for all customers when running multiple congestion algorithms at the same time.

\section{Planning}
\begin{table}[H]
\centering
\label{my-label}
\begin{tabular}{l|p{8cm}}
\textbf{Week} & \textbf{Activity} \\ \hline
1    & Conducting desk research on TCP congestion control algorithms. Conducting desk research on the congestion and packet loss occuring in different scenarios. Conducting desk research on bandwidth management in data centers. Setup the testing environment. \\
2    & Finish test environment setup and conduct all tests.\\
3    & Process the results and start writing the project report.        \\
4    & Finalize our project report and process the final results. Setting up the presentation and practice it.       
\end{tabular}
\caption{Project planning}
\end{table}

\section{Ethical Considerations}
For this research we will create a private setup for testing the performance of TCP congestion control algorithms. This research does only involve TCP congestion control algorithms made available publicly. Therefore we do not expect any ethical issues during this research.

\section{Related work}
The TCP congestion control algorithm BBR is discussed in \cite{bbr-congestion}. The journal contains a performance test of the BBR algorithm and a comparison between the BBR algorithm and the CUBIC algorithm, where a noticeable difference in bandwidth allocation is shown when using small buffers. In \cite{multiple-congestion} a performance analysis of multiple TCP congestion control algorithms is performed. This research does not include the new congestion control algorithms we want to analyze and does not include the datacenter context. In \cite{dctcp-congestion-original} an algorithm is proposed to avoid congestion within datacenters. The authors describe an algorithm that maintains a small buffer occupancy and has an early detection of congestion.

\printbibliography

\end{document}

### 2017-11-17 initial.tex
formatting

@@
 \documentclass{article}
+
 \usepackage[]{float}
@@
 \section{Introduction}
-Multitentant datacenters aim to provide services with high availability and reliability.
-Some datacenters experience the issue that parts of their networks are sometimes congested, which might result in lower availability.
-Congestion occurs when a buffer of the receiving network device is full and the network device drops packets.
-TCP has a mechanism to avoid congestion in the network: TCP congestion control.
-Multiple algorithms are available for congestion control and show a different behaviour depending on the network's characteristics.
-In 2010 algorithm DCTCP was described in an article \cite{dctcp-congestion-original} and published as RFC8257 in 2017 \cite{dctcp-congestion}.
-DCTCP is optimized for datacenters and provides a high-burst tolerance, low latency, and high throughput when the datacenter has a small part of the buffer available \cite{dctcp-congestion}.
-In 2016 one algorithm was proposed: BBR.
-BBR is a TCP congestion algorithm created by Google that achieves higher bandwidths and lower latencies compared to other TCP congestion methods \cite{bbr-congestion}.
-A comparison of BBR with CUBIC \cite{bbr-congestion-comparison} shows that the BBR node pushes the CUBIC node away in bandwidth when using small buffers.
-The BBR node gets more bandwidth allocated than the CUBIC node.
 
-We propose to do a performance and behaviour analysis of different TCP congestion control algorithms in a datacenter network.
-Based on the results we will recommend datacenters on measures to keep providing an available and reliable environment for customers when multiple congestion control algorithms active in the network.
+Multitentant datacenters aim to provide services with high availability and
+reliability. Some datacenters experience the issue that parts of their networks
+are sometimes congested, which might result in lower availability. Congestion
+occurs when a buffer of the receiving network device is full and the network
+device drops packets. TCP has a mechanism to avoid congestion in the network:
+TCP congestion control. Multiple algorithms are available for congestion
+control and show a different behaviour depending on the network's
+characteristics. In 2010 algorithm DCTCP was described in an article
+\cite{dctcp-congestion-original} and published as RFC8257 in 2017
+\cite{dctcp-congestion}. DCTCP is optimized for datacenters and provides a
+high-burst tolerance, low latency, and high throughput when the datacenter has
+a small part of the buffer available \cite{dctcp-congestion}. In 2016 one
+algorithm was proposed: BBR. BBR is a TCP congestion algorithm created by
+Google that achieves higher bandwidths and lower latencies compared to other
+TCP congestion methods \cite{bbr-congestion}. A comparison of BBR with CUBIC
+\cite{bbr-congestion-comparison} shows that the BBR node pushes the CUBIC node
+away in bandwidth when using small buffers. The BBR node gets more bandwidth
+allocated than the CUBIC node.
+
+We propose to do a performance and behaviour analysis of different TCP
+congestion control algorithms in a datacenter network. Based on the results we
+will recommend datacenters on measures to keep providing an available and
+reliable environment for customers when multiple congestion control algorithms
+active in the network.
 
 \section{Research Questions}
-Our main research question is: How can a datacenter provide an available and reliable environment for customers when multiple TCP congestion control algorithms are used in the network?
-\\ \\
+
+Our main research question is: How can a datacenter provide an available and
+reliable environment for customers when multiple TCP congestion control
+algorithms are used in the network?
+
 This main research question divides in five sub questions:
+
 \begin{enumerate}
@@
 \section{Scope}
-During this research we will focus on at least the TCP congestion control algorithms BBR, CUBIC, DCTCP.
-Additional congestion control algorithms will be determined after researching what the most implemented TCP congestion algorithms are in datacenter environments.
-We will conduct a behavior analysis and comparison between those algorithms, including the impact of mixed usage of the algorithms on the availability of a service in a datacenter.
-We will look at the measures a datacenter has to take to keep their service available and reliable for all customers when running multiple congestion algorithms at the same time.
+
+During this research we will focus on at least the TCP congestion control
+algorithms BBR, CUBIC, DCTCP. Additional congestion control algorithms will be
+determined after researching what the most implemented TCP congestion
+algorithms are in datacenter environments. We will conduct a behavior analysis
+and comparison between those algorithms, including the impact of mixed usage of
+the algorithms on the availability of a service in a datacenter. We will look
+at the measures a datacenter has to take to keep their service available and
+reliable for all customers when running multiple congestion algorithms at the
+same time.
 
 \section{Planning}
+
 \begin{table}[H]
@@
 \section{Ethical Considerations}
-For this research we will create a private setup for testing the performance of TCP congestion control algorithms.
-This research does only involve TCP congestion control algorithms made available publicly.
-Therefore we do not expect any ethical issues during this research.
+
+For this research we will create a private setup for testing the performance of
+TCP congestion control algorithms. This research does only involve TCP
+congestion control algorithms made available publicly. Therefore we do not
+expect any ethical issues during this research.
 
 \section{Related work}
+
 The TCP congestion control algorithm BBR is discussed in \cite{bbr-congestion}.
-The journal contains a performance test of the BBR algorithm and a comparison between the BBR algorithm and the CUBIC algorithm, where a noticeable difference in bandwidth allocation is shown when using small buffers.
-In \cite{multiple-congestion} a performance analysis of multiple TCP congestion control algorithms is performed.
-This research does not include the new congestion control algorithms we want to analyze and does not include the datacenter context.
-In \cite{dctcp-congestion-original} an algorithm is proposed to avoid congestion within datacenters.
-The authors describe an algorithm that maintains a small buffer occupancy and has an early detection of congestion.
+The journal contains a performance test of the BBR algorithm and a comparison
+between the BBR algorithm and the CUBIC algorithm, where a noticeable
+difference in bandwidth allocation is shown when using small buffers. In
+\cite{multiple-congestion} a performance analysis of multiple TCP congestion
+control algorithms is performed. This research does not include the new
+congestion control algorithms we want to analyze and does not include the
+datacenter context. In \cite{dctcp-congestion-original} an algorithm is
+proposed to avoid congestion within datacenters. The authors describe an
+algorithm that maintains a small buffer occupancy and has an early detection of
+congestion.
 

### 2017-11-17 initial.tex
textual changes

@@
 
-\title{LS Project Proposal: TCP Congestion Algorithms In Datacenters}
+\title{TCP Congestion Algorithms In Datacenters \\
+	\vspace{0.3cm}
+	{\large LS Project Proposal}
+}
 \date{November 16, 2017}
@@
 
-Multitentant datacenters aim to provide services with high availability and
-reliability. Some datacenters experience the issue that parts of their networks
-are sometimes congested, which might result in lower availability. Congestion
-occurs when a buffer of the receiving network device is full and the network
-device drops packets. TCP has a mechanism to avoid congestion in the network:
+Multitentant datacenters aim to provide services with consistent performance
+and reliability. When customers send a lot of data in a short amount of time,
+networks might become congested, which results in lowered performance.
+
+Congestion occurs when a buffer on the path to the recipient is full and
+packets are dropped. TCP has a mechanism to avoid congestion in the network:
 TCP congestion control. Multiple algorithms are available for congestion
 control and show a different behaviour depending on the network's
-characteristics. In 2010 algorithm DCTCP was described in an article
-\cite{dctcp-congestion-original} and published as RFC8257 in 2017
-\cite{dctcp-congestion}. DCTCP is optimized for datacenters and provides a
+characteristics.
+
+In 2010, the algorithm DCTCP was described in an
+article\cite{dctcp-congestion-original} and published as RFC8257 in
+2017\cite{dctcp-congestion}. DCTCP is optimized for datacenters and provides a
 high-burst tolerance, low latency, and high throughput when the datacenter has
-a small part of the buffer available \cite{dctcp-congestion}. In 2016 one
-algorithm was proposed: BBR. BBR is a TCP congestion algorithm created by
-Google that achieves higher bandwidths and lower latencies compared to other
-TCP congestion methods \cite{bbr-congestion}. A comparison of BBR with CUBIC
-\cite{bbr-congestion-comparison} shows that the BBR node pushes the CUBIC node
-away in bandwidth when using small buffers. The BBR node gets more bandwidth
-allocated than the CUBIC node.
+a small part of the buffer available\cite{dctcp-congestion}. In 2016 another
+algorithm was proposed: BBR. This is a TCP congestion algorithm created by
+Google, which achieves higher bandwidths and lower latencies compared to other
+TCP congestion methods\cite{bbr-congestion}. A comparison of BBR with
+CUBIC\cite{bbr-congestion-comparison} shows that the BBR node pushes the CUBIC
+node away in bandwidth when using small buffers. The BBR node gets more
+bandwidth allocated than the CUBIC node.
 
 We propose to do a performance and behaviour analysis of different TCP
-congestion control algorithms in a datacenter network. Based on the results we
-will recommend datacenters on measures to keep providing an available and
-reliable environment for customers when multiple congestion control algorithms
-active in the network.
+congestion control algorithms in a (simulated) datacenter network. Based on the
+results, we will recommend datacenters on measures to keep providing a reliable
+environment for customers when multiple congestion control algorithms are
+active in the network, for example by using QoS or discouraging certain
+congestion control algorithms.
 
@@
 
-Our main research question is: How can a datacenter provide an available and
-reliable environment for customers when multiple TCP congestion control
-algorithms are used in the network?
+Our main research question is:
+{\it How can a datacenter provide an available and reliable environment for
+customers when multiple TCP congestion control algorithms are used in the
+network?}
 
-This main research question divides in five sub questions:
+\vspace{0.5cm}
+
+This main research question divides in five sub-questions:
 
 \begin{enumerate}
-	\item What TCP congestion algorithms are mostly implemented in datacenter environments?
-	\item How do the TCP congestion algorithms work?
-	\item How do the TCP congestion algorithms differ in behavior?
-	\item What is the impact on the availability and service for customers when using multiple TCP congestion algorithms?
-	\item What connection management measures are able to control the connection when using different congestion control algorithms?
+	\item What TCP congestion algorithms are most commonly implemented in datacenter environments?
+	\item How do these TCP congestion algorithms work?
+	\item How do these TCP congestion algorithms differ in behavior?
+	\item What is the impact on the availability and service for customers when using different, competing congestion algorithms?
+	\item What measures could be taken to split bandwidth equally among tenants?
 \end{enumerate}
@@
 During this research we will focus on at least the TCP congestion control
-algorithms BBR, CUBIC, DCTCP. Additional congestion control algorithms will be
-determined after researching what the most implemented TCP congestion
-algorithms are in datacenter environments. We will conduct a behavior analysis
-and comparison between those algorithms, including the impact of mixed usage of
-the algorithms on the availability of a service in a datacenter. We will look
-at the measures a datacenter has to take to keep their service available and
-reliable for all customers when running multiple congestion algorithms at the
-same time.
+algorithms BBR, CUBIC, and DCTCP. Additional congestion control algorithms will
+be determined after researching what the most implemented TCP congestion
+algorithms are in datacenter environments. We will conduct a behavioral
+analysis and compare those algorithms, including the impact of mixed usage of
+the algorithms on the performance of a service in a datacenter. We will look at
+the measures a datacenter could take to keep their service performant and
+reliable for all tenants.
 
@@
 
-\begin{table}[H]
-	\centering
-	\label{my-label}
-	\begin{tabular}{l|p{8cm}}
-		\textbf{Week} & \textbf{Activity} \\ \hline
-		1    & Conducting desk research on TCP congestion control algorithms.
+\begin{tabular}[H]{ | l | p{10cm} | }
+	\hline
+	\textbf{Week} & \textbf{Activity} \\
+	\hline 1 & Conducting desk research on TCP congestion control algorithms.
 Conducting desk research on the congestion and packet loss occuring in different scenarios.
@@
 \\
-		2    & Finish test environment setup and conduct all tests.\\
-		3    & Process the results and start writing the project report.        \\
-		4    & Finalize our project report and process the final results. Setting up the presentation and practice it.       
+	\hline 2 & Finish test environment setup and conduct all tests.\\
+	\hline 3 & Process the results and start writing the project report. \\
+	\hline 4 & Finalize our project report and process the final results. Setting up the presentation and practice it. \\
+	\hline
 \end{tabular}
-	\caption{Project planning}
-\end{table}
 
@@
 
-For this research we will create a private setup for testing the performance of
-TCP congestion control algorithms. This research does only involve TCP
-congestion control algorithms made available publicly. Therefore we do not
-expect any ethical issues during this research.
+For this research we will create a private setup. This research does only
+involve TCP congestion control algorithms made available publicly. Therefore we
+do not expect any ethical issues during this research.
 
@@
 between the BBR algorithm and the CUBIC algorithm, where a noticeable
-difference in bandwidth allocation is shown when using small buffers. In
-\cite{multiple-congestion} a performance analysis of multiple TCP congestion
+difference in bandwidth allocation is shown when using small buffers.
+
+In \cite{multiple-congestion} a performance analysis of multiple TCP congestion
 control algorithms is performed. This research does not include the new
 congestion control algorithms we want to analyze and does not include the
-datacenter context. In \cite{dctcp-congestion-original} an algorithm is
-proposed to avoid congestion within datacenters. The authors describe an
-algorithm that maintains a small buffer occupancy and has an early detection of
-congestion.
+datacenter context.
+
+In \cite{dctcp-congestion-original} an algorithm is proposed to avoid
+congestion within datacenters. The authors describe an algorithm that maintains
+a small buffer occupancy and has an early detection of congestion.
 

### 2017-11-17 initial.tex
m. Mainly cosmetic changes to planning

@@
 
-\begin{tabular}[H]{ | l | p{10cm} | }
+\begin{tabular}[H]{ | l | p{10.2cm} | }
 	\hline
 	\textbf{Week} & \textbf{Activity} \\
-	\hline 1 & Conducting desk research on TCP congestion control algorithms.
-Conducting desk research on the congestion and packet loss occuring in different scenarios.
-Conducting desk research on bandwidth management in data centers.
-Setup the testing environment.
+	\hline 1 & Conducting desk research on: which TCP congestion control algorithms are used the most, packet loss occuring in different scenarios, and existing bandwidth management in data centers.
+We will also setup the testing environment.
 \\
-	\hline 2 & Finish test environment setup and conduct all tests.\\
+	\hline 2 & Finish test environment setup and conduct tests. \\
 	\hline 3 & Process the results and start writing the project report. \\
-	\hline 4 & Finalize our project report and process the final results. Setting up the presentation and practice it. \\
+	\hline 4 & Process the final results, finalize our project report and prepare the presentation. \\
 	\hline

### 2017-11-17 initial.tex
Main question: s/available/performant/

@@
 Our main research question is:
-{\it How can a datacenter provide an available and reliable environment for
+{\it How can a datacenter provide an available and performant environment for
 customers when multiple TCP congestion control algorithms are used in the

### 2017-11-17 initial.tex
Update initial.tex

--- 2017-11-17 initial.tex
@@
 Our main research question is:
-{\it How can a datacenter provide an available and reliable environment for
+{\it How can a datacenter provide an reliable and performant environment for
 customers when multiple TCP congestion control algorithms are used in the

### 2017-11-20 initial.tex
Added methods section

@@
 
+\section{Methods}
+Measuring the performance of the TCP congestion control algorithms will be done using a simulation setup.
+We have checked whether implementations of CUBIC, BBR and DCTCP are available in the network simulation tools ns-3 or ns-2.
+CUBIC and DCTCP have an implementation for ns-3 \cite{git-tcp-cubic} \cite{git-tcp-dctcp} and also for the Linux kernel.
+BBR is implemented in the Linux kernel version 4.9 and above.
+As BBR only has a complete implementation for the Linux kernel, a virtual test environment will be created to test and compare the performance of the algorithms.
+For algorithms found most common during the research,  it has to be determined whether an implementation is available for those simulation tools.
+
 \section{Planning}

### 2017-11-20 paper/paper.tex
Created folder for paper

--- 2017-11-17 initial.tex
@@
 
-\usepackage[]{float}
+\usepackage{float}
 \usepackage[utf8]{inputenc}
-\usepackage[]{hyperref}
-\usepackage[hyperref=true,
-    natbib=true,
-    style=numeric,
-    backend=bibtex]{biblatex}
-\bibliography{references.bib}
+\usepackage{hyperref}
+%\usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
+%\bibliography{references.bib}
 
@@
 	\vspace{0.3cm}
-	{\large LS Project Proposal}
+	{\large LS Project}
 }
-\date{November 16, 2017}
+\date{\today{}}
 \author{Luc Gommans, Rick van Gorp}
@@
 
+\section{Abstract}
+
 \section{Introduction}
+
+% Copied from proposal
 
@@
 
-\section{Research Questions}
+\section{Related work}
+
+% Copied from proposal
+
+The TCP congestion control algorithm BBR is discussed in \cite{bbr-congestion}.
+The journal contains a performance test of the BBR algorithm and a comparison
+between the BBR algorithm and the CUBIC algorithm, where a noticeable
+difference in bandwidth allocation is shown when using small buffers.
+
+In \cite{multiple-congestion} a performance analysis of multiple TCP congestion
+control algorithms is performed. This research does not include the new
+congestion control algorithms we want to analyze and does not include the
+datacenter context.
+
+In \cite{dctcp-congestion-original} an algorithm is proposed to avoid
+congestion within datacenters. The authors describe an algorithm that maintains
+a small buffer occupancy and has an early detection of congestion.
+
+\section{Method}
+
+\section{Results}
+
+\section{Discussion}
+
+\section{Conclusion}
+
+% For reference, from the proposal:
 
@@
 
-\section{Scope}
-
-During this research we will focus on at least the TCP congestion control
-algorithms BBR, CUBIC, and DCTCP. Additional congestion control algorithms will
-be determined after researching what the most implemented TCP congestion
-algorithms are in datacenter environments. We will conduct a behavioral
-analysis and compare those algorithms, including the impact of mixed usage of
-the algorithms on the performance of a service in a datacenter. We will look at
-the measures a datacenter could take to keep their service performant and
-reliable for all tenants.
-
-\section{Planning}
-
-\begin{tabular}[H]{ | l | p{10.2cm} | }
-	\hline
-	\textbf{Week} & \textbf{Activity} \\
-	\hline 1 & Conducting desk research on: which TCP congestion control algorithms are used the most, packet loss occuring in different scenarios, and existing bandwidth management in data centers.
-We will also setup the testing environment.
-\\
-	\hline 2 & Finish test environment setup and conduct tests. \\
-	\hline 3 & Process the results and start writing the project report. \\
-	\hline 4 & Process the final results, finalize our project report and prepare the presentation. \\
-	\hline
-\end{tabular}
-
-\section{Ethical Considerations}
-
-For this research we will create a private setup. This research does only
-involve TCP congestion control algorithms made available publicly. Therefore we
-do not expect any ethical issues during this research.
-
-\section{Related work}
-
-The TCP congestion control algorithm BBR is discussed in \cite{bbr-congestion}.
-The journal contains a performance test of the BBR algorithm and a comparison
-between the BBR algorithm and the CUBIC algorithm, where a noticeable
-difference in bandwidth allocation is shown when using small buffers.
-
-In \cite{multiple-congestion} a performance analysis of multiple TCP congestion
-control algorithms is performed. This research does not include the new
-congestion control algorithms we want to analyze and does not include the
-datacenter context.
-
-In \cite{dctcp-congestion-original} an algorithm is proposed to avoid
-congestion within datacenters. The authors describe an algorithm that maintains
-a small buffer occupancy and has an early detection of congestion.
-
-\printbibliography
+%\printbibliography
 

### 2017-11-23 paper/paper.tex
Written part of Methods

@@
 \section{Method}
+During this research a performance test of various TCP congestion algorithms will be performed.
+The TCP congestion algorithms will be determined through desk research of the most common algorithms.
+The tests will be conducted in a private environment.
+Based on the results of those tests, recommendations will be given related to the measures a datacenter has to take to keep the traffic equally allocated over customers.
+
+	\subsection{Comparing Performance}
+	Two scenarios will be used to compare the performance of the TCP 				congestion algorithms:
+	\begin{enumerate}
+		\item The user owns one endpoint in the network. In this case only 				the sender side has a congestion algorithm applied.
+		\item The user owns multiple endpoints in the network. In this case a			congestion algorithm is applied at both the client and server side.
+	\end{enumerate}
+
+	The created environments are shown in figure ... and figure ...
 

### 2017-11-27 paper/paper.tex
Update paper.tex
Added figures
Updated methods

@@
 %\usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
+\usepackage{graphicx}
 %\bibliography{references.bib}
@@
 
-	The created environments are shown in figure ... and figure ...
+	The created environments are shown in figure \ref{fig:setup1} and figure \ref{fig:setup2}
+
+	\begin{figure}[H] 
+		\centering
+			\caption{Test setup: Scenario 1}
+  			\includegraphics[scale=0.5]{figs/setup1.png}
+  		\label{fig:setup1}
+	\end{figure}
+
+	\begin{figure}[H] 
+		\centering
+			\caption{Test setup: Scenario 2}
+			\includegraphics[scale=0.5]{figs/setup2.png}
+  		\label{fig:setup2}
+	\end{figure}
+
+	In scenario 1 the server uses the same algorithm as the random node.
+The user node uses a different algorithm for the comparison.
+Tests in this scenario cover behavioral differences between TCP congestion algorithms when a user owns one node within a network.
+In scenario 2 the user server uses the same algorithm as the user node and the random server uses the same algorithm as the random node.
+Tests in this scenario cover behavioral differences between TCP congestion algorithms when a user owns two nodes within a network.
+
+	In both scenarios the TCP congestion algorithms are tested for multiple characteristics:
+	\begin{enumerate}
+		\item TCP Window Size measured over time: 
+		\item Bandwidth allocation: to determine whether using a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more bandwidth to a TCP transport.
+		\item Round Trip Time: to determine the reason of the behaviour of a TCP congestion algorithm in combination with other TCP congestion algorithms.
+	\end{enumerate}
+
+	The tests conducted on the algorithms have different environment variables: packet loss and delay.
+Those environment variables are simulated using netem, which is a network emulation functionality in several Linux distributions.
+Packet loss and delay are chosen as the behavior of most TCP algorithms depend on those variables.
+The values for the environment variables per simulation run are shown in table ...
+
+	\subsection{Traffic Management}
+	Desk research will be conducted to gather information related to traffic management.
+The goal is to find a mechanism that is capable of allocating equally between customers of a datacenter.
+This is done by analyzing the results from the algorithm comparison and the initial desk research on traffic management.
 

### 2017-11-27 paper/paper.tex
minor change

@@
 		\item TCP Window Size measured over time: 
-		\item Bandwidth allocation: to determine whether using a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more bandwidth to a TCP transport.
+		\item Bandwidth allocation: to determine whether using a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more bandwidth to a specific TCP connection.
 		\item Round Trip Time: to determine the reason of the behaviour of a TCP congestion algorithm in combination with other TCP congestion algorithms.

### 2017-11-27 tool/config.py
Started on measurement software

@@ -1,0 +1,3 @@
+class Config:
+    hosts = [{'ip': 'localhost'}]
+

### 2017-11-27 tool/controller.py

@@ -1,0 +1,22 @@
+#!/usr/bin/env python3
+
+import socket, time
+from config import Config
+from shared import *
+
+for host in Config.hosts:
+    sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+    sock.connect((host['ip'], TCPPORT))
+    send(sock, 'test')
+    send(sock, MSG_GETNAME)
+    name = read(sock)
+    print('Connected to ' + name + ', syncing time...')
+
+    send(sock, MSG_SETTIME)
+    time.sleep(0.1)
+    starttime = time.time()
+    send(sock, time.time())
+    d = time.time() - float(read(sock))
+    print('Time sync\'d to {0:f} seconds.'.format(d/2))
+    send(sock, MSG_BYE)
+

### 2017-11-27 tool/disciple.py

@@ -1,0 +1,44 @@
+#!/usr/bin/env python3
+
+import sys, os, socket, datetime, time
+from shared import *
+
+sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+sock.bind(('0.0.0.0', TCPPORT))
+sock.listen(1)
+
+print('Ready to receive commands.')
+
+while True:
+    controller, client_address = sock.accept()
+    while True:
+        data = read(controller)
+        if data == 'test':
+            print('Received connection test at ' + time.ctime())
+
+        elif data == MSG_GETNAME:
+            send(controller, socket.gethostname())
+
+        elif data == MSG_BYE:
+            break
+
+        elif data == MSG_SETALGO:
+            algo = read(controller)
+            os.system('sysctl net.ipv4.tcp_congestion_control=' + algo)
+
+        elif data == MSG_SETTIME:
+            recvtime = read(controller)
+            if os.name == 'nt':
+                t = datetime.fromtimestamp(float(recvtime))
+                recvtime = t.strftime('%H:%I:%S') + recvtime.split('.')[1][:3]
+                os.system('echo ' + recvtime + ' | recvtime')
+            else:
+                os.system('date -s @' + recvtime)
+
+            print('Set the time to ' + recvtime)
+
+            send(controller, time.time())
+
+        else:
+            print('Unrecognized command')
+

### 2017-11-27 tool/shared.py

@@ -1,0 +1,15 @@
+MAXMSGLEN = 100
+TCPPORT = 2
+
+MSG_GETNAME = 'Dear sir/madam, could you please tell me your name?'
+MSG_BYE = 'If you will excuse me, I have to go catch a bus.'
+MSG_SETALGO = 'Could you please change your TCP congestion control algorithm?'
+MSG_SETTIME = 'Could you synchronize the clock at your earliest convenience?'
+
+def send(sock, data):
+    data = str(data)
+    sock.send(bytes(data + (' ' * (MAXMSGLEN - len(data))), 'utf-8'))
+
+def read(sock):
+    return str(sock.recv(MAXMSGLEN), 'utf-8').strip()
+

### 2017-11-30 paper/paper.tex
Added more methods

@@
 \usepackage{hyperref}
-%\usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
+\usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
 \usepackage{graphicx}
-%\bibliography{references.bib}
+\bibliography{references.bib}
 
@@
 % Copied from proposal
+% Edited 27-11-2017
 
@@
 
-We propose to do a performance and behaviour analysis of different TCP
-congestion control algorithms in a (simulated) datacenter network. Based on the
-results, we will recommend datacenters on measures to keep providing a reliable
-environment for customers when multiple congestion control algorithms are
-active in the network, for example by using QoS or discouraging certain
-congestion control algorithms.
+This report includes a performance and behavior analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network.
+Based on the results, we will recommend datacenters on measures to keep providing a reliable environment for customers when multiple congestion control algoritms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
+The main research question is: {\it How can a datacenter provide an reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the
+network?}
 
@@
 
+	\subsection{Selecting TCP congestion algorithms}	
+	In order to select the most common TCP congestion algorithms desk research will be performed.
+Results of scientific reports will show which algorithms are used more often than other algoritms.
+Based on those results a list of TCP congestion algorithms compatible with a multitenant datacenter environment will be composed.
+
 	\subsection{Comparing Performance}
-	Two scenarios will be used to compare the performance of the TCP congestion algorithms:
+	One scenario will be used to compare the performance of the TCP congestion algorithms:
 	\begin{enumerate}
-		\item The user owns one endpoint in the network. In this case only the sender side has a congestion algorithm applied.
-		\item The user owns multiple endpoints in the network. In this case a congestion algorithm is applied at both the client and server side.
+		\item The user owns multiple endpoints in the network.
+In this case a congestion algorithm is applied at both the client and server side.
+This scenario applies to multitentant datacenter situations, where a user owns a server in the datacenter network and a node outside of the datacenter network.
 	\end{enumerate}
 
-	The created environments are shown in figure \ref{fig:setup1} and figure \ref{fig:setup2}
+	The created environment is shown in figure \ref{fig:setup1}.
 
@@
 		\centering
-			\caption{Test setup: Scenario 1}
-  			\includegraphics[scale=0.5]{figs/setup1.png}
+  			\includegraphics[scale=0.5]{figs/setup2.png}
+  			\caption{Test setup - Scenario 1}
   		\label{fig:setup1}
@@
 
-	\begin{figure}[H] 
-		\centering
-			\caption{Test setup: Scenario 2}
-			\includegraphics[scale=0.5]{figs/setup2.png}
-  		\label{fig:setup2}
-	\end{figure}
-
-	In scenario 1 the server uses the same algorithm as the random node.
-The user node uses a different algorithm for the comparison.
-Tests in this scenario cover behavioral differences between TCP congestion algorithms when a user owns one node within a network.
-In scenario 2 the user server uses the same algorithm as the user node and the random server uses the same algorithm as the random node.
-Tests in this scenario cover behavioral differences between TCP congestion algorithms when a user owns two nodes within a network.
+	In scenario 1 the user server uses the same algorithm as the user node and the random server uses the same algorithm as the random node.
+Tests in this scenario cover behavioral differences between TCP congestion algorithms when a user owns a node in the network of a multitenant datacenter.
 
@@
 	\end{enumerate}
-
 The tests conducted on the algorithms have different environment variables: packet loss and delay.
-Those environment variables are simulated using netem, which is a network emulation functionality in several Linux distributions.
+Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions.
 Packet loss and delay are chosen as the behavior of most TCP algorithms depend on those variables.
 The values for the environment variables per simulation run are shown in table ...
+\\
+\\
+	\begin{tabular}[H]{ | l | l | l |}
+	\hline
+	\textbf{Test} & \textbf{Delay} & \textbf{Packet loss} \\
+	\hline 1 & 0ms & 0\%. \\
+	\hline 2 & 0ms & 0.01\% \\
+	\hline 3 & 0ms & 0.10\% \\
+	\hline 4 & 0ms & 1.0\% \\
+	\hline 1 & 0ms & 10\%. \\
+	\hline
+	\end{tabular}
 

### 2017-11-30 paper/paper.tex
Typo

@@
 Based on the results, we will recommend datacenters on measures to keep providing a reliable environment for customers when multiple congestion control algoritms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
-The main research question is: {\it How can a datacenter provide an reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the
+The main research question is: {\it How can a datacenter provide a reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the
 network?}

### 2017-11-30 paper/paper.tex
Added delay and packet loss calcs

@@
 During this research a performance test of various TCP congestion algorithms will be performed.
-The TCP congestion algorithms will be determined through desk research of the most common algorithms.
+The most common TCP congestion algorithms will be determined through desk research.
 The tests will be conducted in a private environment.
@@
 Results of scientific reports will show which algorithms are used more often than other algoritms.
-Based on those results a list of TCP congestion algorithms compatible with a multitenant datacenter environment will be composed.
+Based on those results a list of TCP congestion algorithms will be composed.
 
@@
 	\begin{enumerate}
-		\item TCP Window Size measured over time: 
+		\item TCP Window Size measured over time: to determine whether using a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of a larger window or a smaller window to a specific TCP connection.
 		\item Bandwidth allocation: to determine whether using a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more bandwidth to a specific TCP connection.
@@
 	\end{enumerate}
-The tests conducted on the algorithms have different environment variables: packet loss and delay.
-Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions.
+The tests conducted on the algorithms have two environment variables: packet loss and delay.
+Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions \cite{linux-netem}.
 Packet loss and delay are chosen as the behavior of most TCP algorithms depend on those variables.
-The values for the environment variables per simulation run are shown in table ...
-\\
-\\
-	\begin{tabular}[H]{ | l | l | l |}
+The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.
+
+		\subsubsection{Determining the Packet loss and Delay}
+		The dataset of Verizon \cite{verizon-latency} is used to calculate the packet loss and delay used in this research.
+The dataset includes global information and shows averages per month over the year 2017.
+The packet loss is calculated by taking the average of the Verizon Business Packet Delivery Statistics for Country Specific Metrics for each country.
+Those averages were inverted by subtracting the packet delivery percentage from 100, to show the actual packet loss in percentages and not the packet delivery.
+Based on the ascending sorted values a bar chart was created.
+This bar chart is shown in figure \ref{fig:packet-loss-chart}
+
+	\begin{figure}[H] 
+		\centering
+  			\includegraphics[scale=0.7]{figs/verizon-packetloss.png}
+  			\caption{Average Packet Loss for Country Specific Metrics}
+  		\label{fig:packet-loss-chart}
+	\end{figure}
+
+The country numbers in figure \ref{fig:packet-loss-chart} are only used to show the distribution of the average packet loss.
+The maximum packet loss shown in figure \ref{fig:packet-loss-chart} equals 0.6\% and the minimum equals 0\%.
+However, in case of situations where more packet loss occurs than usual, the double value of the maximum is used: 1.2\%.
+The values for packet loss in this comparison set-up were determined based on the distribution shown in figure \ref{fig:packet-loss-chart} and the manually set maximum of 1.2\%.
+The values are shown in table \ref{table:test-packetloss}.
+
+	\begin{table}[H]
+		\centering
+		\caption{Packet loss percentages used for the comparison setup}
+		\begin{tabular}[H]{ | l |}
 		\hline
-	\textbf{Test} & \textbf{Delay} & \textbf{Packet loss} \\
-	\hline 1 & 0ms & 0\%. \\
-	\hline 2 & 0ms & 0.01\% \\
-	\hline 3 & 0ms & 0.10\% \\
-	\hline 4 & 0ms & 1.0\% \\
-	\hline 1 & 0ms & 10\%. \\
+		\textbf{Packet loss} \\
+		\hline 0\%. \\
+		\hline 0.01\% \\
+		\hline 0.1\% \\
+		\hline 0.6\% \\
+		\hline 1.2\% \\
 		\hline
 		\end{tabular}
+		\label{table:test-packetloss}
+	\end{table}
+
+The delay is calculated by taking the average of the Verizon Business Latency Statistics for Country Specific Metrics for each country.
+Based on the ascending sorted values a bar chart was created.
+This bar chart is shown in figure \ref{fig:verizon-delay-chart}
+
+	\begin{figure}[H] 
+		\centering
+  			\includegraphics[scale=0.7]{figs/verizon-delay.png}
+  			\caption{Average Latency for Country Specific Metrics}
+  		\label{fig:verizon-delay-chart}
+	\end{figure}
+
+The country numbers in figure \ref{fig:verizon-delay-chart} are only used to show the distribution of the average delays.
+The maximum delay shown in figure \ref{fig:verizon-delay-chart} equals 290ms and the minimum equals 8ms (both rounded).
+The values for delay in this comparison set-up were determined based on the distribution shown in figure \ref{fig:verizon-delay-chart}.
+The values are shown in table \ref{table:test-delay}.
+
+	\begin{table}[H]
+		\centering
+		\caption{Packet loss percentages used for the comparison setup}
+		\begin{tabular}[H]{ | l |}
+		\hline
+		\textbf{Delay} \\
+		\hline  8ms\\
+		\hline  64ms  \\
+		\hline  120ms  \\
+		\hline  176ms \\
+		\hline	232ms \\
+		\hline  290ms  \\
+		\hline
+		\end{tabular}
+		\label{table:test-delay}
+	\end{table}
+
+		\subsubsection{Performing time-based automated tests}
+		The tests will be performed using scripts written in Python [\ref{appendix:python}].
+The scripts either act as controller or as disciple.
+The controller is responsible for controlling the tests, by synchronizing the time, exchanging commands to run, modify or stop the tests and gathering data.
+The disciple is the listener and executes all task commands the controller sends.
+The commands include getting the hostname of a disciple, setting the TCP congestion algorithm of a disciple, configuring the time of a disciple, starting the data transfer from a disciple and closing the connection with the controller.
 
@@
 
-%\printbibliography
+\printbibliography
+
+\appendix
+\section{Python}
+\label{appendix:python}
+SCRIPT
 

### 2017-11-30 tool/config.py
Implementing send/receiving of data and measuring speed

@@ -1,3 +1,12 @@
 class Config:
-    hosts = [{'ip': 'localhost'}]
+    hosts = [{'ip': '10.0.0.6'}]
+    setups = [
+        {
+            'hosts': [('10.0.0.5', '10.0.0.6')],
+            'settings': {
+                'duration': 10,
+                'name': 'Testing 10.6->10.5'
+            }
+        }
+    ]
 

### 2017-11-30 tool/controller.py

@@ -18,5 +18,39 @@ for host in Config.hosts:
     send(sock, time.time())
     d = time.time() - float(read(sock))
     print('Time sync\'d to {0:f} seconds.'.format(d/2))
+
     send(sock, MSG_BYE)
+    
+for setup in Config.setups:
+    duration = setup['settings']['duration']
+    name = setup['settings']['name']
+    print("Running setup: " + name)
+
+    for host_from, host_to in setup['hosts']:
+        sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+        sock.connect((host_to, TCPPORT))
+        send(sock, MSG_LISTEN)
+        send(sock, MSG_BYE)
+        time.sleep(0.1) # Otherwise the other host apparently tries to connect too fast
+
+        sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+        sock.connect((host_from, TCPPORT))
+        send(sock, MSG_SEND)
+        send(sock, host_to)
+        send(sock, duration)
+        send(sock, int(time.time()) + 3)
+        send(sock, MSG_BYE)
+
+    print('Sleeping till end of test...')
+    time.sleep(duration + 6)
+
+    for host, role in setup['hosts']:
+        if role == 'receiver':
+            sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+            sock.connect((host, TCPPORT))
+            send(sock, MSG_GETRESULTS)
+            print(read(sock))
+            send(sock, MSG_BYE)
+
+    print('Done running setups')
 

### 2017-11-30 tool/disciple.py

@@ -11,6 +11,7 @@ print('Ready to receive commands.')
 
 while True:
     controller, client_address = sock.accept()
+    print('Connected.')
     while True:
         data = read(controller)
         if data == 'test':
@@ -20,6 +21,7 @@ while True:
             send(controller, socket.gethostname())
 
         elif data == MSG_BYE:
+            print('Disconnected.')
             break
 
         elif data == MSG_SETALGO:
@@ -38,6 +40,59 @@ while True:
             print('Set the time to ' + recvtime)
 
             send(controller, time.time())
+
+        elif data == MSG_SEND:
+            dst = read(controller)
+            duration = int(read(controller))
+            when = int(read(controller))
+
+            ds = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+            ds.connect((dst, TCPPORT + 1))
+
+            while time.time() < when:
+                pass
+
+            until = when + duration
+            data = b'A' * MTU
+            while time.time() < until:
+                ds.send(data)
+
+            ds.close()
+
+        elif data == MSG_LISTEN:
+            ds = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+            ds.bind(('0.0.0.0', TCPPORT + 1))
+            ds.listen(1)
+            print('Listening on data socket, awaiting client.')
+
+            t = int(time.time())
+            timings = {}
+            for i in range(t, t+1000):
+                timings[t] = 0
+
+            client = ds.accept()
+            print('Client connected.')
+            while True:
+                l = len(client.recv(MTU)) ## TODO: "AttributeError: 'tuple' object has no attribute 'recv'"
+                if l == 0:
+                    break
+                timings[int(time.time())] += l
+
+            first = float('inf')
+            for t in timings:
+                count = timings[t]
+                if count > 0 and t < first:
+                    first = t
+
+            results = ''
+            for t in range(first + 1, first + 9):
+                count = timings[t]
+                results += str(count) + ' '
+            results = results.strip()
+
+        elif data == MSG_GETRESULTS:
+            print('Sending results.')
+            send(controller, results)
 
         else:
             print('Unrecognized command')

### 2017-11-30 tool/shared.py

@@ -1,10 +1,14 @@
 MAXMSGLEN = 100
-TCPPORT = 2
+TCPPORT = 1
+MTU = 1400
 
 MSG_GETNAME = 'Dear sir/madam, could you please tell me your name?'
 MSG_BYE = 'If you will excuse me, I have to go catch a bus.'
 MSG_SETALGO = 'Could you please change your TCP congestion control algorithm?'
 MSG_SETTIME = 'Could you synchronize the clock at your earliest convenience?'
+MSG_LISTEN = 'Data will be forthcoming. Observe, hack, make.'
+MSG_SEND = 'ATTACK'
+MSG_GETRESULTS = 'What were the results? :D'
 
 def send(sock, data):
     data = str(data)

### 2017-11-30 tool/controller.py
Transfer speed test working.

@@ -44,13 +44,12 @@ for setup in Config.setups:
     print('Sleeping till end of test...')
     time.sleep(duration + 6)
 
-    for host, role in setup['hosts']:
-        if role == 'receiver':
-            sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-            sock.connect((host, TCPPORT))
-            send(sock, MSG_GETRESULTS)
-            print(read(sock))
-            send(sock, MSG_BYE)
+    for host_from, host_to in setup['hosts']:
+        sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+        sock.connect((host_to, TCPPORT))
+        send(sock, MSG_GETRESULTS)
+        print(read(sock))
+        send(sock, MSG_BYE)
 
     print('Done running setups')
 

### 2017-11-30 tool/disciple.py

@@ -11,7 +11,7 @@ print('Ready to receive commands.')
 
 while True:
     controller, client_address = sock.accept()
-    print('Connected.')
+    print('Connection from {}.'.format(client_address))
     while True:
         data = read(controller)
         if data == 'test':
@@ -67,16 +67,19 @@ while True:
 
             t = int(time.time())
             timings = {}
-            for i in range(t, t+1000):
-                timings[t] = 0
+            for i in range(t, t+MAXTESTDURATION):
+                timings[i] = 0
 
-            client = ds.accept()
-            print('Client connected.')
+            client, addr = ds.accept()
+            print('Client {} connected.'.format(addr))
             while True:
-                l = len(client.recv(MTU)) ## TODO: "AttributeError: 'tuple' object has no attribute 'recv'"
+                l = len(client.recv(MTU))
                 if l == 0:
                     break
                 timings[int(time.time())] += l
+
+            client.close() # Should already be closed, since received_data.length==0 earlier, but just to be sure that the listsock doesn't remain in time-wait
+            ds.close()
 
             first = float('inf')
             for t in timings:

### 2017-11-30 tool/shared.py

@@ -1,6 +1,7 @@
-MAXMSGLEN = 100
+MAXMSGLEN = 100 # bytes
 TCPPORT = 1
-MTU = 1400
+MTU = 1400 # bytes
+MAXTESTDURATION = 1000 # seconds
 
 MSG_GETNAME = 'Dear sir/madam, could you please tell me your name?'
 MSG_BYE = 'If you will excuse me, I have to go catch a bus.'

### 2017-11-30 tool/disciple.py
m. slightly more verbose output

@@ -11,7 +11,7 @@ print('Ready to receive commands.')
 
 while True:
     controller, client_address = sock.accept()
-    print('Connection from {}.'.format(client_address))
+    print('Controller connection from {}.'.format(client_address))
     while True:
         data = read(controller)
         if data == 'test':
@@ -21,7 +21,7 @@ while True:
             send(controller, socket.gethostname())
 
         elif data == MSG_BYE:
-            print('Disconnected.')
+            print('Controller disconnected from socket.')
             break
 
         elif data == MSG_SETALGO:
@@ -80,6 +80,7 @@ while True:
 
             client.close() # Should already be closed, since received_data.length==0 earlier, but just to be sure that the listsock doesn't remain in time-wait
             ds.close()
+            print('Closed listening data socket.')
 
             first = float('inf')
             for t in timings:

### 2017-12-01 tool/controller.py
Enabled setting delay and loss

@@ -1,7 +1,7 @@
 #!/usr/bin/env python3
 
 import socket, time
-from config import Config
+from ctrlconfig import Config
 from shared import *
 
 for host in Config.hosts:

### 2017-12-01 tool/disciple.py

@@ -3,6 +3,19 @@
 import sys, os, socket, datetime, time
 from shared import *
 
+os.system('''
+ip l | grep ifb0;
+if [ $? -ne 0 ]; then
+    modprobe ifb;
+    ip link set dev ifb0 up;
+    tc qdisc add dev eth0 ingress;
+    tc filter add dev eth0 parent ffff: protocol ip u32 match u32 0 0 flowid 1:1 action mirred egress redirect dev ifb0;
+    echo Setup ifb interface;
+else
+    echo ifb interface already setup;
+fi;
+''')
+
 sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
 sock.bind(('0.0.0.0', TCPPORT))
 sock.listen(1)
@@ -48,16 +61,17 @@ while True:
 
             ds = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
             ds.connect((dst, TCPPORT + 1))
+            print('Connected to data sink, waiting until ' + str(when))
 
             while time.time() < when:
                 pass
 
             until = when + duration
-            data = b'A' * MTU
             while time.time() < until:
-                ds.send(data)
+                ds.send(TRANSMIT_DATA)
 
             ds.close()
+            print('Done. Disconnecting from data sink.')
 
         elif data == MSG_LISTEN:
             ds = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
@@ -97,6 +111,12 @@ while True:
         elif data == MSG_GETRESULTS:
             print('Sending results.')
             send(controller, results)
+
+        elif data == MSG_SETCONNPROPS:
+            loss = read(controller)
+            delay = read(controller)
+            print('Setting delay={} loss={}'.format(delay, loss))
+            os.system('tc qdisc add dev ifb0 root netem delay ' + delay + ' loss ' + loss)
 
         else:
             print('Unrecognized command')

### 2017-12-01 tool/shared.py

@@ -1,7 +1,13 @@
+# Settings
 MAXMSGLEN = 100 # bytes
 TCPPORT = 1
 MTU = 1400 # bytes
 MAXTESTDURATION = 1000 # seconds
+# End of settings
+
+import os
+
+TRANSMIT_DATA = os.urandom(MTU)
 
 MSG_GETNAME = 'Dear sir/madam, could you please tell me your name?'
 MSG_BYE = 'If you will excuse me, I have to go catch a bus.'
@@ -10,6 +16,7 @@ MSG_SETTIME = 'Could you synchronize the clock at your earliest convenience?'
 MSG_LISTEN = 'Data will be forthcoming. Observe, hack, make.'
 MSG_SEND = 'ATTACK'
 MSG_GETRESULTS = 'What were the results? :D'
+MSG_SETCONNPROPS = 'Connection properties need changing!'
 
 def send(sock, data):
     data = str(data)

### 2017-12-01 tool/disciple.py
Only setup ifb on posix; correct interface name (eth0->eno1)

@@ -3,18 +3,19 @@
 import sys, os, socket, datetime, time
 from shared import *
 
-os.system('''
-ip l | grep ifb0;
-if [ $? -ne 0 ]; then
-    modprobe ifb;
-    ip link set dev ifb0 up;
-    tc qdisc add dev eth0 ingress;
-    tc filter add dev eth0 parent ffff: protocol ip u32 match u32 0 0 flowid 1:1 action mirred egress redirect dev ifb0;
-    echo Setup ifb interface;
-else
-    echo ifb interface already setup;
-fi;
-''')
+if os.name == 'posix':
+    os.system('''
+    ip l | grep ifb0 > /dev/null;
+    if [ $? -ne 0 ]; then
+        modprobe ifb;
+        ip link set dev ifb0 up;
+        tc qdisc add dev eno1 ingress;
+        tc filter add dev eno1 parent ffff: protocol ip u32 match u32 0 0 flowid 1:1 action mirred egress redirect dev ifb0;
+        echo Setup ifb interface;
+    else
+        echo ifb interface already setup;
+    fi;
+    ''')
 
 sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
 sock.bind(('0.0.0.0', TCPPORT))

### 2017-12-01 tool/ctrlconfig.py
Refactoring controller; implemented version check

@@ -1,12 +1,0 @@
-class Config:
-    hosts = [{'ip': '10.0.0.6'}]
-    setups = [
-        {
-            'hosts': [('10.0.0.5', '10.0.0.6')],
-            'settings': {
-                'duration': 10,
-                'name': 'Testing 10.6->10.5'
-            }
-        }
-    ]
-

### 2017-12-01 paper/paper.tex
Added sections for results

@@
 Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions \cite{linux-netem}.
-Packet loss and delay are chosen as the behavior of most TCP algorithms depend on those variables.
+Packet loss and delay are chosen as the behavior of most TCP algorithms depend on those variables [SRC].
 The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.
@@
 
+		\subsubsection{Testing hardware}
+		The specifications of the hardware used for the tests are shown in table \ref{table:spec1}, table \ref{table:spec2} and table \ref{table:spec3}.
+The switch shown in table \ref{table:spec3} has all router related functionalities, which might impact the network speed, turned off.
+
+		\begin{table}[H]
+			\centering
+			\caption{Node configurations}
+			\begin{tabular}[H]{ | l | l | }
+			\hline
+			\textbf{Hardware} & \textbf{Specification} \\
+			\hline  Model & Dell Optiplex 7010\\
+			\hline  CPU & Intel(R) Core(TM) i5-3570S CPU @ 3.10GHz\\
+			\hline  Memory & 12GB\\
+			\hline  NIC & Intel Corporation 82579LM Gigabit Network Connection (rev 04)\\
+			\hline	NIC Driver & e100e 3.2.6-k\\
+			\hline
+			\end{tabular}
+			\label{table:spec1}
+		\end{table}
+
+		\begin{table}[H]
+			\centering
+			\caption{Switch 1 configuration}
+			\begin{tabular}[H]{ | l | l | }
+			\hline
+			\textbf{Hardware} & \textbf{Specification} \\
+			\hline  Model & 3Com OfficeConnect Gigabit Switch 16\\
+			\hline  Ports & 16x Gigabit Ethernet\\
+			\hline
+			\end{tabular}
+			\label{table:spec2}
+		\end{table}
+
+		\begin{table}[H]
+			\centering
+			\caption{Switch 2 configuration}
+			\begin{tabular}[H]{ | l | l | }
+			\hline
+			\textbf{Hardware} & \textbf{Specification} \\
+			\hline  Model & Sitecom Gigabit Router 300N-XR\\
+			\hline  Ports & 4x Gigabit Ethernet\\
+			\hline
+			\end{tabular}
+			\label{table:spec3}
+		\end{table}
+
 	\subsection{Traffic Management}
@@
 \section{Results}
+	\subsection{TCP Congestion Algorithms}
+	The most common TCP congestion algorithms are CTCP, CUBIC and BIC.
+CTCP is the default congestion algorithm in Microsoft Windows Server Editions and Microsoft Windows 8 and newer [SRC].
+CUBIC is the default congestion algorithm in Linux distributions with kernel versions ...
+and newer [SRC].
+BIC was the default congestion algorithm in earlier Linux distributions with kernel versions: [SRC].
+DCTCP and BBR will also be tested as those protocols are newer.
+DCTCP is specifically designed for datacenters [SRC] and BBR is a new congestion protocol not yet widely deployed [SRC].
+
+		\subsubsection{Algorithm Characteristics}
+		- Includes characteristics of the algorithms. To what do the TCP congestion algorithms respond according to literature? And how do they deal with that (If available in literature).
+
+		\subsubsection{Performance comparison}	
+		- Includes results of tests.
+
+	\subsection{Traffic Management}
+	- Includes what is available for trafic management in a datacenter.
+	- Measures to take to equally allocate bandwidth to customers.
+		\subsubsection{Quality of Service}
 

### 2017-12-01 paper/paper.tex
Added sources

@@
 	Desk research will be conducted to gather information related to traffic management.
-The goal is to find a mechanism that is capable of allocating equally between customers of a datacenter.
+The goal is to find a mechanism that is capable of allocating bandwidth equally between customers of a datacenter.
 This is done by analyzing the results from the algorithm comparison and the initial desk research on traffic management.
@@
 	The most common TCP congestion algorithms are CTCP, CUBIC and BIC.
-CTCP is the default congestion algorithm in Microsoft Windows Server Editions and Microsoft Windows 8 and newer [SRC].
-CUBIC is the default congestion algorithm in Linux distributions with kernel versions ...
-and newer [SRC].
-BIC was the default congestion algorithm in earlier Linux distributions with kernel versions: [SRC].
+CTCP is the default congestion algorithm in Microsoft Windows Server 2008 and newer \cite{cubic-kernel-version}.
+CUBIC is the default congestion algorithm in Linux distributions with kernel versions 2.6.19 and higher \cite{cubic-kernel-version}.
+BIC was the default congestion algorithm in earlier Linux distributions from kernel version 2.6.8 until 2.6.19 \cite{bic-kernel-version} \cite{cubic-kernel-version}.
 DCTCP and BBR will also be tested as those protocols are newer.
-DCTCP is specifically designed for datacenters [SRC] and BBR is a new congestion protocol not yet widely deployed [SRC].
+DCTCP is specifically designed for datacenters \cite{dctcp-congestion} and BBR is a new congestion protocol not yet widely deployed \cite{bbr-congestion}.
 
@@
 \begin{enumerate}
-	\item What TCP congestion algorithms are most commonly implemented in datacenter environments?
+	\item What TCP congestion algorithms are most commonly implemented?
 	\item How do these TCP congestion algorithms work?

### 2017-12-01 tool/disciple.py
Split up setconnprops so that we can set the congestion algo on servers and delay+loss on clients

@@ -112,13 +112,19 @@ while True:
             print('Sending results.')
             send(controller, results)
 
+        elif data == MSG_SETALGO:
+            algo = read(controller)
+            if os.name == 'nt':
+                print('Ignoring setalgo because I am New Technology.')
+            else:
+                print('Setting algo={}'.format(algo))
+                os.system('sysctl net.ipv4.tcp_congestion_control=' + algo)
+
         elif data == MSG_SETCONNPROPS:
             loss = read(controller)
             delay = read(controller)
-            algo = read(controller)
-            print('Setting delay={} loss={} algo={}'.format(delay, loss, algo))
+            print('Setting delay={} loss={}'.format(delay, loss))
             os.system('tc qdisc add dev ifb0 root netem delay ' + delay + ' loss ' + loss)
-            os.system('sysctl net.ipv4.tcp_congestion_control=' + algo)
 
         else:
             print('Unrecognized command')

### 2017-12-01 tool/shared.py

@@ -17,7 +17,8 @@ MSG_SETTIME = 'Could you synchronize the clock at your earliest convenience?'
 MSG_LISTEN = 'Data will be forthcoming. Observe, hack, make.'
 MSG_SEND = 'ATTACK'
 MSG_GETRESULTS = 'What were the results? :D'
-MSG_SETCONNPROPS = 'chprops'
+MSG_SETCONNPROPS = 'chprops plz'
+MSG_SETALGO = 'chalgo plz'
 MSG_GETVERSION = 'How old are you, if I may ask?'
 
 def send(sock, data):

### 2017-12-01 tool/controller.py
Rewrote controller

@@ -4,6 +4,7 @@
 algos = ['cubic', 'ctcp', 'dctcp', 'bic', 'bbr']
 delays = [8, 64, 120, 176, 232, 290]
 losses = [0, 0.01, 0.1, 0.6, 1.2]
+test_duration = 10
 
 hosts = {
     'server1': '10.0.0.6',
@@ -43,5 +44,11 @@ for host in hosts:
 
     send(sock, MSG_BYE)
 
-    #conn_test(host_from, host_to, duration=int(time.time()) + 3)
+for algo1 in algos:
+    for algo2 in algos:
+        for delay in delays:
+            for loss in losses:
+                print('Running algo1={} algo2={} delay={} loss={}'.format(algo1, algo2, delay, loss))
+
+                results.update(runtest(hosts, test_duration, delay, loss, algo1, algo2))
 

### 2017-12-01 tool/controllerlib.py

@@ -8,7 +8,6 @@ def conn_test(host_from, host_to, duration, delay, loss, algo, at):
     send(sock, MSG_SETCONNPROPS)
     send(sock, str(loss) + '%')
     send(sock, str(delay) + 'ms')
-    send(sock, algo.lower())
 
     send(sock, MSG_BYE)
 
@@ -16,10 +15,15 @@ def conn_test(host_from, host_to, duration, delay, loss, algo, at):
 
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
     sock.connect((host_from, TCPPORT))
+
     send(sock, MSG_SEND)
     send(sock, host_to)
     send(sock, duration)
     send(sock, at)
+
+    send(sock, MSG_SETALGO)
+    send(sock, algo.lower())
+
     send(sock, MSG_BYE)
 
     time.sleep(duration + CONNTESTGAP * 2)
@@ -29,6 +33,31 @@ def conn_test(host_from, host_to, duration, delay, loss, algo, at):
     send(sock, MSG_GETRESULTS)
     results = map(int, read(sock).split(' '))
     send(sock, MSG_BYE)
+
+    return results
+
+def runtest(hosts, duration, delay, loss, algo1, algo2):
+    tag = 'algo1={} algo2={} delay={} duration={} '.format(algo1, algo2, delay, duration)
+
+    if algo1 == 'ctcp':
+        s1 = hosts['winserv']
+    else:
+        s1 = hosts['server1']
+
+    if algo2 == 'ctcp':
+        s1 = hosts['winserv']
+    else:
+        s2 = hosts['server2']
+
+    at = time.time() + CONNTESTGAP
+    results[tag + 'run=1 s={} c=1'.format(s1)] = conn_test(s1, hosts['client1'], duration=test_duration, loss=loss, delay=delay, algo=algo1, at=at)
+    results[tag + 'run=1 s={} c=2'.format(s2)] = conn_test(s2, hosts['client2'], duration=test_duration, loss=loss, delay=delay, algo=algo2, at=at)
+    time.sleep(test_duration + CONNTESTGAP * 2)
+
+    at = time.time() + CONNTESTGAP
+    results[tag + 'run=2 s={} c=2'.format(s1)] = conn_test(s1, hosts['client2'], duration=test_duration, loss=loss, delay=delay, algo=algo, at=at)
[…]

### 2017-12-01 paper/paper.tex
Minor changes

@@
 This report includes a performance and behavior analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network.
-Based on the results, we will recommend datacenters on measures to keep providing a reliable environment for customers when multiple congestion control algoritms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
+Based on the results, we will recommend datacenters on measures to take to keep providing a reliable environment for customers when multiple congestion control algoritms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
 The main research question is: {\it How can a datacenter provide a reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the
@@
 \section{Results}
+Type some text here.
+
 	\subsection{TCP Congestion Algorithms}

### 2017-12-04 tool/disciple.py
Fixed datetime issue for NT

@@ -44,7 +44,7 @@ while True:
         elif data == MSG_SETTIME:
             recvtime = read(controller)
             if os.name == 'nt':
-                t = datetime.fromtimestamp(float(recvtime))
+                t = datetime.datetime.fromtimestamp(float(recvtime))
                 recvtime = t.strftime('%H:%I:%S') + recvtime.split('.')[1][:3]
                 os.system('echo ' + recvtime + ' | time')
             else:

### 2017-12-04 tool/controller.py
Fixed import bugs and defs

@@ -7,7 +7,7 @@ losses = [0, 0.01, 0.1, 0.6, 1.2]
 test_duration = 10
 
 hosts = {
-    'server1': '10.0.0.6',
+    'server1': '10.0.0.5',
     'server2': '10.0.0.1',
     'winserv': '10.0.0.3',
     'client1': '10.0.0.2',
@@ -23,6 +23,7 @@ hostnames = {}  # Will be automatically obtained
 
 for host in hosts:
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+    print(hosts[host])
     sock.connect((hosts[host], TCPPORT))
 
     send(sock, MSG_GETNAME)
@@ -43,6 +44,8 @@ for host in hosts:
     print('Time synchronized with {}|{}.'.format(host, name))
 
     send(sock, MSG_BYE)
+
+results = {}
 
 for algo1 in algos:
     for algo2 in algos:

### 2017-12-04 tool/disciple.py
Add error reporting

@@ -17,115 +17,129 @@ if os.name == 'posix':
     fi;
     ''')
 
+controlleraddr = None
+
 sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
 sock.bind(('0.0.0.0', TCPPORT))
 sock.listen(1)
 
 print('Ready to receive commands.')
 
-while True:
-    controller, client_address = sock.accept()
-    print('Controller connection from {}.\n'.format(client_address))
+try:
     while True:
-        data = read(controller)
-        if data == 'test':
-            print('Received connection test at ' + time.ctime())
+        controller, addr = sock.accept()
+        controlleraddr = addr
+        print('Controller connection from {}.\n'.format(addr))
+        while True:
+            data = read(controller)
+            if data == 'test':
+                print('Received connection test at ' + time.ctime())
 
-        elif data == MSG_GETVERSION:
-            send(controller, VERSION)
+            elif data == MSG_GETVERSION:
+                send(controller, VERSION)
 
-        elif data == MSG_GETNAME:
-            send(controller, socket.gethostname())
+            elif data == MSG_GETNAME:
+                send(controller, socket.gethostname())
 
-        elif data == MSG_BYE:
-            print('Controller disconnected from socket.')
-            break
+            elif data == MSG_BYE:
+                print('Controller disconnected from socket.')
+                break
 
-        elif data == MSG_SETTIME:
-            recvtime = read(controller)
-            if os.name == 'nt':
-                t = datetime.fromtimestamp(float(recvtime))
-                recvtime = t.strftime('%H:%I:%S') + recvtime.split('.')[1][:3]
-                os.system('echo ' + recvtime + ' | time')
+            elif data == MSG_SETTIME:
+                recvtime = read(controller)
+                if os.name == 'nt':
+                    t = datetime.fromtimestamp(float(recvtime))
+                    recvtime = t.strftime('%H:%I:%S') + recvtime.split('.')[1][:3]
+                    os.system('echo ' + recvtime + ' | time')
+                else:
+                    os.system('date -s @' + recvtime + ' >/dev/null')
+
+                print('Set the time to ' + recvtime)
+
+                send(controller, time.time())
+
+            elif data == MSG_SEND:
+                dst = read(controller)
+                duration = int(read(controller))
+                when = int(read(controller))
+
+                ds = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+                ds.connect((dst, TCPPORT + 1))
[…]

### 2017-12-04 tool/deploy.sh
v0.1 of deploy.sh

@@ -1,0 +1,9 @@
+#!/usr/bin/env bash
+
+for target in 10.0.0.1 10.0.0.2 10.0.0.4 10.0.0.5; do
+	echo Copying to $target
+	scp disciple.py shared.py $target:
+done
+
+echo todo: windows at .3
+

### 2017-12-04 tool/disciple.py
Set timeout for listener

@@ -88,6 +88,7 @@ try:
                 for i in range(t, t+MAXTESTDURATION):
                     timings[i] = 0
 
+                ds.settimeout(CONNTESTGAP * 3)
                 client, addr = ds.accept()
                 print('Client {} connected.'.format(addr))
                 while True:

### 2017-12-04 tool/deploy.sh
fixing bug in disciple; deploying to windows

@@ -1,9 +1,20 @@
 #!/usr/bin/env bash
 
 for target in 10.0.0.1 10.0.0.2 10.0.0.4 10.0.0.5; do
-	echo Copying to $target
+	echo Copying to $target...
 	scp disciple.py shared.py $target:
+	echo;
 done
 
-echo todo: windows at .3
+echo Copying to Windows...
+
+if [ ! -f ~/mnt/win/desktop.ini ]; then
+	mkdir -p ~/mnt/win
+	# Yes, I am aware that I'm submitting this uber secret secure password to Github.
+	# The target system is never connected to the internet and it's not meant for
+	# security anyway (password is only set to allow us to log in).
+	sudo mount -t cifs -o user=user,pass=123123 //10.0.0.3/Desktop ~/mnt/win
+fi;
+
+sudo cp disciple.py shared.py ~/mnt/win
 

### 2017-12-04 tool/disciple.py

@@ -61,7 +61,7 @@ try:
             elif data == MSG_SEND:
                 dst = read(controller)
                 duration = int(read(controller))
-                when = int(read(controller))
+                when = float(read(controller))
 
                 ds = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
                 ds.connect((dst, TCPPORT + 1))

### 2017-12-04 tool/controller.py
Implemented killing clients

@@ -15,16 +15,26 @@ hosts = {
 }
 # End of settings
 
-import socket, time
+import socket, time, sys
 from shared import *
 from controllerlib import *
+
+kill = False
+if len(sys.argv) == 2 and sys.argv[1] == 'killclients':
+    kill = True
 
 hostnames = {}  # Will be automatically obtained
 
 for host in hosts:
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-    print(hosts[host])
+    if not kill:
+        print('Connecting to ' + hosts[host])
     sock.connect((hosts[host], TCPPORT))
+
+    if kill:
+        print('Killing ' + hosts[host])
+        send(sock, MSG_DIE)
+        continue
 
     send(sock, MSG_GETNAME)
     name = read(sock)
@@ -44,6 +54,9 @@ for host in hosts:
     print('Time synchronized with {}|{}.'.format(host, name))
 
     send(sock, MSG_BYE)
+
+if kill:
+    exit(0)
 
 results = {}
 

### 2017-12-04 tool/deploy.sh

@@ -1,5 +1,7 @@
 #!/usr/bin/env bash
 
+python3 controller.py killclients
+
 for target in 10.0.0.1 10.0.0.2 10.0.0.4 10.0.0.5; do
 	echo Copying to $target...
 	scp disciple.py shared.py $target:

### 2017-12-04 tool/disciple.py

@@ -131,6 +131,9 @@ try:
                 print('Setting delay={} loss={}'.format(delay, loss))
                 os.system('tc qdisc add dev ifb0 root netem delay ' + delay + ' loss ' + loss)
 
+            elif data == MSG_DIE:
+                exit(0)
+
             else:
                 print('Unrecognized command')
 except Exception as e:

### 2017-12-04 tool/shared.py

@@ -20,6 +20,7 @@ MSG_GETRESULTS = 'What were the results? :D'
 MSG_SETCONNPROPS = 'chprops plz'
 MSG_SETALGO = 'chalgo plz'
 MSG_GETVERSION = 'How old are you, if I may ask?'
+MSG_DIE = 'The end is near'
 
 def send(sock, data):
     data = str(data)

### 2017-12-04 tool/controllerlib.py
Fixing bug (var undefined)

@@ -51,6 +51,8 @@ def runtest(hosts, test_duration, delay, loss, algo1, algo2):
     else:
         s2 = hosts['server2']
 
+    results = {}
+
     at = time.time() + CONNTESTGAP
     results[tag + 'run=1 s={} c=1'.format(s1)] = conn_test(s1, hosts['client1'], duration=test_duration, loss=loss, delay=delay, algo=algo1, at=at)
     results[tag + 'run=1 s={} c=2'.format(s2)] = conn_test(s2, hosts['client2'], duration=test_duration, loss=loss, delay=delay, algo=algo2, at=at)

### 2017-12-04 tool/disciple.py
Bugfix disciple

@@ -140,7 +140,7 @@ except Exception as e:
     err = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
     err.settimeout(1)
     try:
-        err.connect((controlleraddr, TCPPORT + 2))
+        err.connect((controlleraddr[0], TCPPORT + 2))
         send(err, 'Error at ' + socket.gethostname() + ': ' + str(e) + '\n\n')
         err.close()
     except:

### 2017-12-04 tool/disciple.py
Kill disciple anyway after error reporting (we will have to restart anyway)

@@ -138,12 +138,13 @@ try:
                 print('Unrecognized command')
 except Exception as e:
     err = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-    err.settimeout(1)
+    err.settimeout(2)
     try:
         err.connect((controlleraddr[0], TCPPORT + 2))
         send(err, 'Error at ' + socket.gethostname() + ': ' + str(e) + '\n\n')
         err.close()
     except:
         print('Error reporting failed')
-        raise e
+
+    raise e
 

### 2017-12-04 tool/controller.py
Set lower timeout on trying to connect to clients

@@ -27,6 +27,7 @@ hostnames = {}  # Will be automatically obtained
 
 for host in hosts:
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+    sock.settimeout(3)
     if not kill:
         print('Connecting to ' + hosts[host])
     sock.connect((hosts[host], TCPPORT))

### 2017-12-04 tool/controller.py
Print results

@@ -66,6 +66,7 @@ for algo1 in algos:
         for delay in delays:
             for loss in losses:
                 print('Running algo1={} algo2={} delay={} loss={}'.format(algo1, algo2, delay, loss))
-
                 results.update(runtest(hosts, test_duration, delay, loss, algo1, algo2))
 
+for result in results:
+    print(result + ": " + results[result])

### 2017-12-04 paper/paper.tex
Updated methods

@@
 This scenario applies to multitentant datacenter situations, where a user owns a server in the datacenter network and a node outside of the datacenter network.
+In this test the user node and random node(s) are at the receiving side and the user server and random server at the sending side.
+The tests will be performed from a multitenant datacenter perspective.
 	\end{enumerate}
@@
 
-	In both scenarios the TCP congestion algorithms are tested for multiple characteristics:
-	\begin{enumerate}
-		\item TCP Window Size measured over time: to determine whether using a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of a larger window or a smaller window to a specific TCP connection.
-		\item Bandwidth allocation: to determine whether using a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more bandwidth to a specific TCP connection.
-		\item Round Trip Time: to determine the reason of the behaviour of a TCP congestion algorithm in combination with other TCP congestion algorithms.
-	\end{enumerate}
+	In scenario 1 the TCP congestion algorithm is tested for fairness in sharing bandwidth with other TCP connections using other TCP congestion algorithms.
+In order to test the fairness of bandwidth shares, the bandwidth allocation per algorithm is measured.
+It can be determined whether a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more or less bandwidth to that specific TCP connection.
+
 The tests conducted on the algorithms have two environment variables: packet loss and delay.
 Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions \cite{linux-netem}.
-Packet loss and delay are chosen as the behavior of most TCP algorithms depend on those variables [SRC].
+\texttt{netem} is applied to the incoming side of the interfaces of the clients to emulate the same packet loss and delay.
+Windows, which is tested as a server, does not support \texttt{netem}.
+Therefore the packet loss and delay are applied to the incoming interfaces of the clients.
+Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables [SRC].
 The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.
+
+%\item TCP Window Size measured over time: to determine whether using a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of a larger window or a smaller window to a specific TCP connection.
+% \item Round Trip Time: to determine the reason of the behaviour of a TCP congestion algorithm in combination with other TCP congestion algorithms.
 
@@
 The disciple is the listener and executes all task commands the controller sends.
-The commands include getting the hostname of a disciple, setting the TCP congestion algorithm of a disciple, configuring the time of a disciple, starting the data transfer from a disciple and closing the connection with the controller.
+The commands include getting the hostname of a disciple, setting the TCP congestion algorithm of a disciple, configuring the time of a disciple, starting the data transfer and measurement tools on the disciple and closing the connection with the controller.
 
@@
 		\subsubsection{Algorithm Characteristics}
+
 		- Includes characteristics of the algorithms. To what do the TCP congestion algorithms respond according to literature? And how do they deal with that (If available in literature).

### 2017-12-04 tool/controller.py
Use a different port for each datastream test; do not report failures to TCPPORT+3 (missing stack traces); other various fixes

@@ -40,6 +40,7 @@ for host in hosts:
     send(sock, MSG_GETNAME)
     name = read(sock)
     hostnames[host] = name
+    hostnames[hosts[host]] = name
 
     send(sock, MSG_GETVERSION)
     version = int(read(sock))
@@ -59,6 +60,8 @@ for host in hosts:
 if kill:
     exit(0)
 
+print('')
+
 results = {}
 
 for algo1 in algos:
@@ -66,7 +69,9 @@ for algo1 in algos:
         for delay in delays:
             for loss in losses:
                 print('Running algo1={} algo2={} delay={} loss={}'.format(algo1, algo2, delay, loss))
-                results.update(runtest(hosts, test_duration, delay, loss, algo1, algo2))
+                results.update(runtest(hosts, hostnames, test_duration, delay, loss, algo1, algo2))
+                print('')
 
 for result in results:
-    print(result + ": " + results[result])
+    print(result + ": " + str(results[result]))
+

### 2017-12-04 tool/controllerlib.py

@@ -2,14 +2,17 @@ from shared import *
 import time
 import socket
 
+dataport = int(40e3)
+
 def conn_test(host_from, host_to, duration, delay, loss, algo, at):
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
     sock.connect((host_to, TCPPORT))
-    send(sock, MSG_LISTEN)
-
     send(sock, MSG_SETCONNPROPS)
     send(sock, str(loss) + '%')
     send(sock, str(delay) + 'ms')
+
+    send(sock, MSG_LISTEN)
+    send(sock, dataport)
 
     send(sock, MSG_BYE)
 
@@ -18,28 +21,32 @@ def conn_test(host_from, host_to, duration, delay, loss, algo, at):
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
     sock.connect((host_from, TCPPORT))
 
-    send(sock, MSG_SEND)
-    send(sock, host_to)
-    send(sock, duration)
-    send(sock, at)
-
     send(sock, MSG_SETALGO)
     send(sock, algo.lower())
 
+    send(sock, MSG_SEND)
+    send(sock, host_to)
+    send(sock, dataport)
+    send(sock, duration)
+    send(sock, at)
+
     send(sock, MSG_BYE)
 
-    time.sleep(duration + CONNTESTGAP * 2)
 
+def get_results(host):
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-    sock.connect((host_to, TCPPORT))
+    sock.connect((host, TCPPORT))
     send(sock, MSG_GETRESULTS)
-    results = map(int, read(sock).split(' '))
+    results = read(sock)
     send(sock, MSG_BYE)
 
     return results
 
-def runtest(hosts, test_duration, delay, loss, algo1, algo2):
-    tag = 'algo1={} algo2={} delay={} duration={} '.format(algo1, algo2, delay, test_duration)
+
+def runtest(hosts, hostnames, test_duration, delay, loss, algo1, algo2):
+    global dataport
+
+    tag = 'delay={} duration={} '.format(delay, test_duration)
 
     if algo1 == 'ctcp':
         s1 = hosts['winserv']
@@ -53,15 +60,25 @@ def runtest(hosts, test_duration, delay, loss, algo1, algo2):
 
     results = {}
 
[…]

### 2017-12-04 tool/disciple.py

@@ -16,135 +16,136 @@ if os.name == 'posix':
         echo ifb interface already setup;
     fi;
     ''')
+    os.system('tc qdisc add dev ifb0 root netem delay 0ms loss 0%')
+    os.system('tc qdisc change dev ifb0 root netem delay 0ms loss 0%')
 
 controlleraddr = None
 
 sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+sock.setsockopt(socket.SOL_SOCKET, socket.SO_REUSEADDR, 1)
 sock.bind(('0.0.0.0', TCPPORT))
 sock.listen(1)
 
 print('Ready to receive commands.')
 
-try:
+while True:
+    controller, addr = sock.accept()
+    controlleraddr = addr
+    print('Controller connection from {}.\n'.format(addr))
     while True:
-        controller, addr = sock.accept()
-        controlleraddr = addr
-        print('Controller connection from {}.\n'.format(addr))
-        while True:
-            data = read(controller)
-            if data == 'test':
-                print('Received connection test at ' + time.ctime())
+        data = read(controller)
+        if data == 'test':
+            print('Received connection test at ' + time.ctime())
 
-            elif data == MSG_GETVERSION:
-                send(controller, VERSION)
+        elif data == MSG_GETVERSION:
+            send(controller, VERSION)
 
-            elif data == MSG_GETNAME:
-                send(controller, socket.gethostname())
+        elif data == MSG_GETNAME:
+            send(controller, socket.gethostname())
 
-            elif data == MSG_BYE:
-                print('Controller disconnected from socket.')
-                break
+        elif data == MSG_BYE:
+            print('Controller disconnected from socket.')
+            break
 
-            elif data == MSG_SETTIME:
-                recvtime = read(controller)
-                if os.name == 'nt':
-                    t = datetime.datetime.fromtimestamp(float(recvtime))
-                    recvtime = t.strftime('%H:%I:%S') + recvtime.split('.')[1][:3]
-                    os.system('echo ' + recvtime + ' | time')
-                else:
-                    os.system('date -s @' + recvtime + ' >/dev/null')
-
-                print('Set the time to ' + recvtime)
-
-                send(controller, time.time())
-
-            elif data == MSG_SEND:
-                dst = read(controller)
-                duration = int(read(controller))
-                when = float(read(controller))
-
-                ds = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
[…]

### 2017-12-04 tool/shared.py

@@ -4,7 +4,7 @@ TCPPORT = 1
 MTU = 1400  # bytes
 MAXTESTDURATION = 1000  # seconds
 CONNTESTGAP = 3  # seconds
-VERSION = 1
+VERSION = 2
 # End of settings
 
 import os

### 2017-12-05 tool/controller.py
Clients can now run on different ports

@@ -11,7 +11,13 @@ hosts = {
     'server2': '10.0.0.1',
     'winserv': '10.0.0.3',
     'client1': '10.0.0.2',
-    'client2': '10.0.0.4',
+    'client2': '10.0.0.2',
+}
+
+# Only non-standard ports need to be mentioned in the ports table.
+defaultport = 1
+ports = {
+    'client2': 2,
 }
 # End of settings
 
@@ -30,7 +36,7 @@ for host in hosts:
     sock.settimeout(3)
     if not kill:
         print('Connecting to ' + hosts[host])
-    sock.connect((hosts[host], TCPPORT))
+    sock.connect((hosts[host], defaultport if host not in ports else ports[host]))
 
     if kill:
         print('Killing ' + hosts[host])
@@ -69,7 +75,7 @@ for algo1 in algos:
         for delay in delays:
             for loss in losses:
                 print('Running algo1={} algo2={} delay={} loss={}'.format(algo1, algo2, delay, loss))
-                results.update(runtest(hosts, hostnames, test_duration, delay, loss, algo1, algo2))
+                results.update(runtest(hosts, hostnames, defaultport, ports, test_duration, delay, loss, algo1, algo2))
                 print('')
 
 for result in results:

### 2017-12-05 tool/controllerlib.py

@@ -4,9 +4,9 @@ import socket
 
 dataport = int(40e3)
 
-def conn_test(host_from, host_to, duration, delay, loss, algo, at):
+def conn_test(host_from, host_to, duration, delay, loss, algo, at, port):
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-    sock.connect((host_to, TCPPORT))
+    sock.connect((host_to, port))
     send(sock, MSG_SETCONNPROPS)
     send(sock, str(loss) + '%')
     send(sock, str(delay) + 'ms')
@@ -19,7 +19,7 @@ def conn_test(host_from, host_to, duration, delay, loss, algo, at):
     time.sleep(0.1)  # Otherwise the other host apparently tries to connect too fast
 
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-    sock.connect((host_from, TCPPORT))
+    sock.connect((host_from, port))
 
     send(sock, MSG_SETALGO)
     send(sock, algo.lower())
@@ -33,9 +33,9 @@ def conn_test(host_from, host_to, duration, delay, loss, algo, at):
     send(sock, MSG_BYE)
 
 
-def get_results(host):
+def get_results(host, port):
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-    sock.connect((host, TCPPORT))
+    sock.connect((host, port))
     send(sock, MSG_GETRESULTS)
     results = read(sock)
     send(sock, MSG_BYE)
@@ -43,7 +43,7 @@ def get_results(host):
     return results
 
 
-def runtest(hosts, hostnames, test_duration, delay, loss, algo1, algo2):
+def runtest(hosts, hostnames, defaultport, ports, test_duration, delay, loss, algo1, algo2):
     global dataport
 
     tag = 'delay={} duration={} '.format(delay, test_duration)
@@ -62,23 +62,23 @@ def runtest(hosts, hostnames, test_duration, delay, loss, algo1, algo2):
 
     print('{}->{} and {}->{}'.format(hostnames[s1], hostnames[hosts['client1']], hostnames[s2], hostnames[hosts['client2']]))
     at = time.time() + CONNTESTGAP
-    conn_test(s1, hosts['client1'], duration=test_duration, loss=loss, delay=delay, algo=algo1, at=at)
-    conn_test(s2, hosts['client2'], duration=test_duration, loss=loss, delay=delay, algo=algo2, at=at)
[…]

### 2017-12-05 tool/disciple.py

@@ -3,6 +3,12 @@
 import sys, os, socket, datetime, time
 from shared import *
 
+if len(sys.argv) == 1 or '-h' in sys.argv or '--help' in sys.argv:
+    print('Usage: python3 disciple.py <port>')
+    exit(1)
+
+port = int(sys.argv[1])
+
 if os.name == 'posix':
     os.system('''
     ip l | grep ifb0 > /dev/null;
@@ -23,7 +29,7 @@ controlleraddr = None
 
 sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
 sock.setsockopt(socket.SOL_SOCKET, socket.SO_REUSEADDR, 1)
-sock.bind(('0.0.0.0', TCPPORT))
+sock.bind(('0.0.0.0', port))
 sock.listen(1)
 
 print('Ready to receive commands.')

### 2017-12-05 tool/shared.py

@@ -1,6 +1,5 @@
 # Settings
 MAXMSGLEN = 100  # bytes
-TCPPORT = 1
 MTU = 1400  # bytes
 MAXTESTDURATION = 1000  # seconds
 CONNTESTGAP = 3  # seconds

### 2017-12-05 tool/shared.py
m. proto version bump

@@ -3,7 +3,7 @@ MAXMSGLEN = 100  # bytes
 MTU = 1400  # bytes
 MAXTESTDURATION = 1000  # seconds
 CONNTESTGAP = 3  # seconds
-VERSION = 2
+VERSION = 3
 # End of settings
 
 import os

### 2017-12-05 tool/controller.py
Numerous changes (committing after a day's work...)

@@ -4,14 +4,14 @@
 algos = ['cubic', 'ctcp', 'dctcp', 'bic', 'bbr']
 delays = [8, 64, 120, 176, 232, 290]
 losses = [0, 0.01, 0.1, 0.6, 1.2]
-test_duration = 10
+test_duration = 20
 
 hosts = {
-    'server1': '10.0.0.5',
-    'server2': '10.0.0.1',
+    'server1': '10.0.0.1',
+    'server2': '10.0.0.2',
     'winserv': '10.0.0.3',
-    'client1': '10.0.0.2',
-    'client2': '10.0.0.2',
+    'client1': '10.0.0.4',
+    'client2': '10.0.0.4',
 }
 
 # Only non-standard ports need to be mentioned in the ports table.
@@ -21,7 +21,7 @@ ports = {
 }
 # End of settings
 
-import socket, time, sys
+import socket, time, sys, os
 from shared import *
 from controllerlib import *
 
@@ -31,12 +31,30 @@ if len(sys.argv) == 2 and sys.argv[1] == 'killclients':
 
 hostnames = {}  # Will be automatically obtained
 
+if not kill:
+    sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+    sock.settimeout(3)
+    sock.connect((hosts['winserv'], defaultport if 'winserv' not in ports else ports['winserv']))
+
+    send(sock, MSG_GETTIME)
+    starttime = time.time()
+    wintime = read(sock)
+    os.system('date -s @' + wintime + ' >/dev/null')
+    print('Corrected time by ' + str(starttime-float(wintime)) + ' seconds')
+    send(sock, MSG_BYE)
+    sock.close()
+
 for host in hosts:
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
     sock.settimeout(3)
     if not kill:
         print('Connecting to ' + hosts[host])
-    sock.connect((hosts[host], defaultport if host not in ports else ports[host]))
+    try:
+        sock.connect((hosts[host], defaultport if host not in ports else ports[host]))
+    except:
+        if not kill:
+            raise
+        continue
 
     if kill:
         print('Killing ' + hosts[host])
@@ -54,14 +72,18 @@ for host in hosts:
         print('Incompatible client detected: {}|{} (v{} instead of {}). Aborting.'.format(host, name, version, VERSION))
         exit(2)
 
[…]

### 2017-12-05 tool/controllerlib.py

@@ -5,6 +5,8 @@ import socket
 dataport = int(40e3)
 
 def conn_test(host_from, host_to, duration, delay, loss, algo, at, port):
+    global dataport
+
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
     sock.connect((host_to, port))
     send(sock, MSG_SETCONNPROPS)
@@ -16,10 +18,10 @@ def conn_test(host_from, host_to, duration, delay, loss, algo, at, port):
 
     send(sock, MSG_BYE)
 
-    time.sleep(0.1)  # Otherwise the other host apparently tries to connect too fast
+    time.sleep(0.1)  # Otherwise the other host tries to connect too fast
 
     sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-    sock.connect((host_from, port))
+    sock.connect((host_from, 1))
 
     send(sock, MSG_SETALGO)
     send(sock, algo.lower())
@@ -31,6 +33,8 @@ def conn_test(host_from, host_to, duration, delay, loss, algo, at, port):
     send(sock, at)
 
     send(sock, MSG_BYE)
+
+    dataport += 1
 
 
 def get_results(host, port):
@@ -44,9 +48,7 @@ def get_results(host, port):
 
 
 def runtest(hosts, hostnames, defaultport, ports, test_duration, delay, loss, algo1, algo2):
-    global dataport
-
-    tag = 'delay={} duration={} '.format(delay, test_duration)
+    tag = 'delay={} loss={} duration={} '.format(delay, loss, test_duration)
 
     if algo1 == 'ctcp':
         s1 = hosts['winserv']
@@ -54,7 +56,7 @@ def runtest(hosts, hostnames, defaultport, ports, test_duration, delay, loss, al
         s1 = hosts['server1']
 
     if algo2 == 'ctcp':
-        s1 = hosts['winserv']
+        s2 = hosts['winserv']
     else:
         s2 = hosts['server2']
 
@@ -62,9 +64,8 @@ def runtest(hosts, hostnames, defaultport, ports, test_duration, delay, loss, al
 
     print('{}->{} and {}->{}'.format(hostnames[s1], hostnames[hosts['client1']], hostnames[s2], hostnames[hosts['client2']]))
     at = time.time() + CONNTESTGAP
-    conn_test(s1, hosts['client1'], duration=test_duration, loss=loss, delay=delay, algo=algo1, at=at, defaultport if 'client1' not in ports else ports['client1'])
[…]

### 2017-12-05 tool/deploy.sh

@@ -2,9 +2,9 @@
 
 python3 controller.py killclients
 
-for target in 10.0.0.1 10.0.0.2 10.0.0.4 10.0.0.5; do
+for target in 10.0.0.1 10.0.0.2 10.0.0.4; do
 	echo Copying to $target...
-	scp disciple.py shared.py $target:
+	scp disciple.py shared.py luc-local@$target:
 	echo;
 done
 

### 2017-12-05 tool/disciple.py

@@ -37,7 +37,8 @@ print('Ready to receive commands.')
 while True:
     controller, addr = sock.accept()
     controlleraddr = addr
-    print('Controller connection from {}.\n'.format(addr))
+    controller.setsockopt(socket.IPPROTO_TCP, socket.TCP_NODELAY, 1)
+    print('\nController connection from {}.'.format(addr))
     while True:
         data = read(controller)
         if data == 'test':
@@ -55,23 +56,39 @@ while True:
 
         elif data == MSG_SETTIME:
             recvtime = read(controller)
+            starttime = time.perf_counter()
+            print('Setting the time to ' + recvtime + ';\n                was ' + str(time.time()))
+
             if os.name == 'nt':
-                t = datetime.datetime.fromtimestamp(float(recvtime))
-                recvtime = t.strftime('%H:%I:%S') + recvtime.split('.')[1][:3]
-                os.system('echo ' + recvtime + ' | time')
+                daysSinceEpoch = int(float(recvtime) / (3600 * 24))
+                remainingSeconds = int(float(recvtime)) - daysSinceEpoch * 3600 * 24
+                h = int(remainingSeconds / 3600)
+                remainingSeconds = remainingSeconds - h * 3600
+                m = int(remainingSeconds / 60)
+                s = remainingSeconds - m * 60
+                frac = recvtime.split('.')[1]
+                t = '{}:{}:{},{}'.format(h + 1, m, s, frac[:2])  # h+1 because of our timezone
+                os.system('echo ' + t + ' | time')
+                print('Converted time to {}.'.format(t))
             else:
                 os.system('date -s @' + recvtime + ' >/dev/null')
 
-            print('Set the time to ' + recvtime)
-
             send(controller, time.time())
-            print("Set time to: " + str(time.time()))
+            settingtimetook = time.perf_counter() - starttime
+            print('Setting time took {} seconds.'.format(settingtimetook))
+
+        elif data == MSG_GETTIME:
+            send(controller, time.time())
 
         elif data == MSG_SEND:
             dst = read(controller)
             dataport = int(read(controller))
             duration = int(read(controller))
             when = float(read(controller))
+
+            if when < time.time():
+                print('Requested time in the past. That isn\'t gonna fly.')
+                exit(4)
 
             ds = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
             ds.connect((dst, dataport))
@@ -90,10 +107,10 @@ while True:
[…]

### 2017-12-05 tool/shared.py

@@ -3,7 +3,7 @@ MAXMSGLEN = 100  # bytes
 MTU = 1400  # bytes
 MAXTESTDURATION = 1000  # seconds
 CONNTESTGAP = 3  # seconds
-VERSION = 3
+VERSION = 4
 # End of settings
 
 import os
@@ -20,6 +20,7 @@ MSG_SETCONNPROPS = 'chprops plz'
 MSG_SETALGO = 'chalgo plz'
 MSG_GETVERSION = 'How old are you, if I may ask?'
 MSG_DIE = 'The end is near'
+MSG_GETTIME = 'Mind telling me the time?'
 
 def send(sock, data):
     data = str(data)

### 2017-12-06 tool/controller.py
added result saving in controller; hardcoded 20s tests in client

@@ -97,18 +97,28 @@ else:
     savefile = ''
 
 savefileout = open('savefile', 'a')
+resultfileout = open('results', 'a')
+resultfileout.write('\n')
 
+num_total_tests = 750
+count = 0
 try:
     for algo1 in algos:
         for algo2 in algos:
+            if algo1 == 'ctcp' or algo2 == 'ctcp':
+                continue # skip for now
+
             for delay in delays:
                 for loss in losses:
+                    count += 1
+
                     config = 'algo1={} algo2={} delay={} loss={}'.format(algo1, algo2, delay, loss)
-                    if config in savefile:
+                    config_alt = 'algo2={} algo1={} delay={} loss={}'.format(algo1, algo2, delay, loss)  # Swap algo1 and 2
+                    if config in savefile or config_alt in savefile:
                         print('Skipping ' + config + ': already in savefile')
                         continue
 
-                    print('Running ' + config)
+                    print('Running ' + config + ' ({}/{}, {} minutes remaining)'.format(count, num_total_tests, (num_total_tests - count) * 2 * 27 / 60))
                     results.update(runtest(hosts, hostnames, defaultport, ports, test_duration, delay, loss, algo1, algo2))
                     savefileout.write(config + '\n')
                     print('')
@@ -117,7 +127,11 @@ except: # Catch keyboardinterrupt -- or error but in all cases, the exception wi
     for result in results:
         print(result + ": " + str(results[result]))
 
+    for result in results:
+        resultfileout.write(result + ": " + str(results[result]) + '\n')
+
     sys.stdout.flush()  # flush stdout so it doesn't mix with stderr that will get printed
+    resultfileout.close()
 
     raise
 

### 2017-12-06 tool/disciple.py

@@ -87,10 +87,11 @@ while True:
             when = float(read(controller))
 
             if when < time.time():
-                print('Requested time in the past. That isn\'t gonna fly.')
+                print('Requested time in the past ({}s). That isn\'t gonna fly.'.format(time.time() - when))
                 exit(4)
 
             ds = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+            ds.settimeout(20 + CONNTESTGAP)
             ds.connect((dst, dataport))
             print('Connected to data sink, waiting until ' + str(when))
 
@@ -117,7 +118,7 @@ while True:
             for i in range(t, t+MAXTESTDURATION):
                 bytecounts[i] = 0
 
-            ds.settimeout(CONNTESTGAP * 3)
+            ds.settimeout(20 + CONNTESTGAP)
             client, addr = ds.accept()
             print('Client {} connected at {}.'.format(addr, time.time()))
             while True:

### 2017-12-06 tool/controller.py
Add testnum=n(r) to results and persist this; write to results2/savefile2; shuffle tests so we get some random sampling instead of a lot of the same ones in case not everything gets done; better ETA estimating; no longer test algo1==algo2; and skip ctcp for now (that might already have been in a previous commit)

@@ -2,7 +2,7 @@
 
 # Settings
 algos = ['cubic', 'ctcp', 'dctcp', 'bic', 'bbr']
-delays = [8, 64, 120, 176, 232, 290]
+delays = [8, 80, 150, 220, 290]
 losses = [0, 0.01, 0.1, 0.6, 1.2]
 test_duration = 20
 
@@ -21,7 +21,7 @@ ports = {
 }
 # End of settings
 
-import socket, time, sys, os
+import socket, time, sys, os, random
 from shared import *
 from controllerlib import *
 
@@ -91,26 +91,29 @@ if kill:
 print('')
 
 results = {}
-if os.path.isfile('savefile'):
-    savefile = open('savefile').read()
+if os.path.isfile('savefile2'):
+    savefile = open('savefile2').read()
 else:
     savefile = ''
 
-savefileout = open('savefile', 'a')
-resultfileout = open('results', 'a')
+savefileout = open('savefile2', 'a')
+resultfileout = open('results2', 'a')
 resultfileout.write('\n')
 
-num_total_tests = 750
+num_total_tests = 0
 count = 0
+todotests = []
 try:
     for algo1 in algos:
         for algo2 in algos:
-            if algo1 == 'ctcp' or algo2 == 'ctcp':
-                continue # skip for now
-
             for delay in delays:
                 for loss in losses:
-                    count += 1
+                    num_total_tests += 1
+                    if algo1 == algo2:
+                        continue
+
+                    if algo1 == 'ctcp' or algo2 == 'ctcp':
+                        continue # skip for now
 
                     config = 'algo1={} algo2={} delay={} loss={}'.format(algo1, algo2, delay, loss)
                     config_alt = 'algo2={} algo1={} delay={} loss={}'.format(algo1, algo2, delay, loss)  # Swap algo1 and 2
@@ -118,10 +121,17 @@ try:
                         print('Skipping ' + config + ': already in savefile')
                         continue
 
-                    print('Running ' + config + ' ({}/{}, {} minutes remaining)'.format(count, num_total_tests, (num_total_tests - count) * 2 * 27 / 60))
-                    results.update(runtest(hosts, hostnames, defaultport, ports, test_duration, delay, loss, algo1, algo2))
-                    savefileout.write(config + '\n')
-                    print('')
+                    todotests.append([algo1, algo2, delay, loss])
+
+    print('Starting {} tests ({} skipped or already done)'.format(len(todotests), num_total_tests))
[…]

### 2017-12-06 tool/controllerlib.py

@@ -3,6 +3,7 @@ import time
 import socket
 
 dataport = int(40e3)
+testnum = int(open('testnum').read())
 
 def conn_test(host_from, host_to, duration, delay, loss, algo, at, port):
     global dataport
@@ -48,7 +49,8 @@ def get_results(host, port):
 
 
 def runtest(hosts, hostnames, defaultport, ports, test_duration, delay, loss, algo1, algo2):
-    tag = 'delay={} loss={} duration={} '.format(delay, loss, test_duration)
+    global testnum
+    tag = 'delay={} loss={} duration={} testnum={} '.format(delay, loss, test_duration, testnum)
 
     if algo1 == 'ctcp':
         s1 = hosts['winserv']
@@ -77,8 +79,14 @@ def runtest(hosts, hostnames, defaultport, ports, test_duration, delay, loss, al
     conn_test(s2, hosts['client1'], duration=test_duration, loss=loss, delay=delay, algo=algo2, at=at, port=defaultport if 'client1' not in ports else ports['client1'])
     time.sleep(test_duration + CONNTESTGAP * 2)
 
+    tag = 'delay={} loss={} duration={} testnum={}r '.format(delay, loss, test_duration, testnum)
     results[tag + 'algo={} run=2 s={} c=2'.format(algo1, s1)] = get_results(hosts['client2'], port=defaultport if 'client2' not in ports else ports['client2'])
     results[tag + 'algo={} run=2 s={} c=1'.format(algo2, s2)] = get_results(hosts['client1'], port=defaultport if 'client1' not in ports else ports['client1'])
+
+    testnum += 1
+    f = open('testnum', 'w')
+    f.write(str(testnum))
+    f.close()
 
     return results
 

### 2017-12-06 tool/shared.py

@@ -2,7 +2,7 @@
 MAXMSGLEN = 100  # bytes
 MTU = 1400  # bytes
 MAXTESTDURATION = 1000  # seconds
-CONNTESTGAP = 3  # seconds
+CONNTESTGAP = 2  # seconds
 VERSION = 4
 # End of settings
 

### 2017-12-07 paper/paper.tex
updates to paper

@@
 
+TCP congestion control algorithms aim to divide the available bandwidth fairly
+among network users. In datacenters, many tenants share a limited amount of bandwidth.
+If too many tenants TODO luc
+
 \section{Introduction}
-
-% Copied from proposal
-% Edited 27-11-2017
 
@@
 and reliability. When customers send a lot of data in a short amount of time,
-networks might become congested, which results in lowered performance.
+networks might become saturated or even congested, and bandwidth needs to be
+distributed fairly among tenants.
 
 Congestion occurs when a buffer on the path to the recipient is full and
-packets are dropped. TCP has a mechanism to avoid congestion in the network:
-TCP congestion control. Multiple algorithms are available for congestion
-control and show a different behaviour depending on the network's
-characteristics.
+packets have to be dropped. TCP uses a congestion control algorithm to avoid
+congestion in the network and distribute bandwidth fairly among users. Several
+algorithms are available to perform this function. Each has a different behaviour
+depending on the network's characteristics.
 
+This report includes a performance and behavior analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network.
+Based on the results, we will recommend datacenters on measures to take to keep providing a reliable environment for customers when multiple congestion control algoritms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
+The main research question is: {\it How can a datacenter provide a reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the
+network?}
+
+\section{Related work}
+
+% Copied from proposal
+
+% todo make this logical
 In 2010, the algorithm DCTCP was described in an
@@
 bandwidth allocated than the CUBIC node.
-
-This report includes a performance and behavior analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network.
-Based on the results, we will recommend datacenters on measures to take to keep providing a reliable environment for customers when multiple congestion control algoritms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
-The main research question is: {\it How can a datacenter provide a reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the
-network?}
-
-\section{Related work}
-
-% Copied from proposal
 
@@
 	In scenario 1 the TCP congestion algorithm is tested for fairness in sharing bandwidth with other TCP connections using other TCP congestion algorithms.
-In order to test the fairness of bandwidth shares, the bandwidth allocation per algorithm is measured.
+In order to test the fairness of congestion algorithms, the bandwidth allocation per algorithm is measured.
 It can be determined whether a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more or less bandwidth to that specific TCP connection.
@@
 		\subsubsection{Algorithm Characteristics}
-
+		The BBR algorithm depends on the parameters BBR.BtlBw and BBR.RTprop.
+Those parameters are estimates defined by the BBR algorithm.
+\\texttt{BtlBw} contains the bandwidth available to data transmissions ...
 		- Includes characteristics of the algorithms. To what do the TCP congestion algorithms respond according to literature? And how do they deal with that (If available in literature).

### 2017-12-07 parser/parser.py
First commit parser for data

@@ -1,0 +1,31 @@
+results = open("newresults", "r")
+
+temp_data = {}
+result_data = {}
+counter = 1
+
+for result in results:
+    if result != "\n":
+        parameters = result.split(":")
+        test_info = parameters[0]
+        test_data = parameters[1].rstrip("\n").lstrip()
+        dictionary = dict(item.split("=") for item in test_info.split(" "))
+        dictionary.update({"bandwidth":test_data})
+        if dictionary["testnum"] not in result_data:
+            result_data[dictionary["testnum"]] = [dictionary]
+        else:
+            result_data[dictionary["testnum"]].append(dictionary)
+
+print("length result_data: " + str(len(result_data)))
+print(result_data)
+            #print(temp_data)
+            #print(len(temp_data))
+            #test_dict = {}
+            #for i in result_data:
+            # if result_data[i]["c"] == "1":
+            # if result_data[i]["run"] == "1":
+            # if result_data[i]["algo"] == "cubic":
+            # print(result_data[i])
+
+
+

### 2017-12-07 parser/parser.py
Added plotting of graphs

@@ -1,8 +1,55 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
 results = open("newresults", "r")
 
-temp_data = {}
+delays = [8, 80, 150, 220, 290]
+losses = [0, 0.01, 0.1, 0.6, 1.2]
+interval = 1 #1 second per bandwidth measurement
 result_data = {}
 counter = 1
+
+def plot_graph(dataset, base_algo, sub_algo, loss, delay):
+    base_plot = {}
+    sub_plot = {} 
+    
+    for result in dataset:
+        if dataset[result][0]["run"] == "1" or dataset[result][0]["run"] == "2": #Always combined same run in dict
+            #Yes, the if statement below is pretty useless. As of now. Not sure if this will have a purpose. 
+            if dataset[result][0]["c"] == "1":
+                t = 0
+                u = 1
+            elif dataset[result][0]["c"] == "2":
+                t = 0 
+                u = 1
+            if dataset[result][t]["algo"] == base_algo and dataset[result][u]["algo"] == sub_algo:
+                print(dataset[result][t]["algo"])
+                if dataset[result][t]["loss"] == loss and dataset[result][t]["delay"] == delay: #The other one will always have the same delay and loss
+                    base_plot[dataset[result][t]["run"]] = [int(x) for x in dataset[result][t]["bandwidth"].split(" ")]
+                    sub_plot[dataset[result][u]["run"]] = [int(x) for x in dataset[result][u]["bandwidth"].split(" ")]
+
+    print("Bandwidth points " + base_algo +  " per 1 second: " + str(base_plot))
+    print("Bandwidth points " + sub_algo + " per 1 second: " + str(sub_plot)) 
+    base_y = np.array(base_plot["1"])
+    sub_y = np.array(sub_plot["1"])
+    base_algor, = plt.plot(base_y, label=base_algo, color="blue")
+    sub_algor, = plt.plot(sub_y, label=sub_algo, color="darkgreen")
+    
+    if len(base_plot) > 1 and len(sub_plot) > 1:
+        base_y_2 = np.array(base_plot["2"])
+        sub_y_2 = np.array(sub_plot["2"])
+        label1 = base_algo + " reverse"
+        label2 = sub_algo + " reverse"
+        base_algo_2, = plt.plot(base_y_2, label=label1, color="purple")
+        sub_algo_2, = plt.plot(sub_y_2, label=label2, color="lightgreen")
[…]

### 2017-12-07 parser/parser.py
Process multiple algorithms and save graphs

@@ -1,8 +1,10 @@
 import matplotlib.pyplot as plt
 import numpy as np
+import os
 
-results = open("newresults", "r")
+results = open("longresults", "r")
 
+algorithms = ['cubic', 'ctcp', 'dctcp', 'bic', 'bbr']
 delays = [8, 80, 150, 220, 290]
 losses = [0, 0.01, 0.1, 0.6, 1.2]
 interval = 1 #1 second per bandwidth measurement
@@ -22,22 +24,29 @@ def plot_graph(dataset, base_algo, sub_algo, loss, delay):
             elif dataset[result][0]["c"] == "2":
                 t = 0 
                 u = 1
+            if len(dataset[result]) < 2:
+                print("Not complete error for algorithm: " + base_algo + " in combination with " + sub_algo) 
+                return False
             if dataset[result][t]["algo"] == base_algo and dataset[result][u]["algo"] == sub_algo:
-                print(dataset[result][t]["algo"])
                 if dataset[result][t]["loss"] == loss and dataset[result][t]["delay"] == delay: #The other one will always have the same delay and loss
                     base_plot[dataset[result][t]["run"]] = [int(x) for x in dataset[result][t]["bandwidth"].split(" ")]
                     sub_plot[dataset[result][u]["run"]] = [int(x) for x in dataset[result][u]["bandwidth"].split(" ")]
+    
+    if "1" not in base_plot:
+        return False
 
     print("Bandwidth points " + base_algo +  " per 1 second: " + str(base_plot))
     print("Bandwidth points " + sub_algo + " per 1 second: " + str(sub_plot)) 
     base_y = np.array(base_plot["1"])
     sub_y = np.array(sub_plot["1"])
-    base_algor, = plt.plot(base_y, label=base_algo, color="blue")
+    base_algor, = plt.plot(base_y, label=base_algo, color="red")
     sub_algor, = plt.plot(sub_y, label=sub_algo, color="darkgreen")
-    
+    plot_height = max(max(base_plot["1"]), max(sub_plot["1"])) + 1000
     if len(base_plot) > 1 and len(sub_plot) > 1:
         base_y_2 = np.array(base_plot["2"])
         sub_y_2 = np.array(sub_plot["2"])
+        print(len(base_plot["1"]))
+        plot_height = max(max(base_plot["1"]), max(sub_plot["1"]), max(base_plot["2"]), max(sub_plot["2"])) + 1000
[…]

### 2017-12-07 parser/parser.py
Fixed issue with reverse tests

@@ -2,7 +2,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import os
 
-results = open("longresults", "r")
+results = open("newres", "r")
 
 algorithms = ['cubic', 'ctcp', 'dctcp', 'bic', 'bbr']
 delays = [8, 80, 150, 220, 290]
@@ -18,19 +18,28 @@ def plot_graph(dataset, base_algo, sub_algo, loss, delay):
     for result in dataset:
         if dataset[result][0]["run"] == "1" or dataset[result][0]["run"] == "2": #Always combined same run in dict
             #Yes, the if statement below is pretty useless. As of now. Not sure if this will have a purpose. 
-            if dataset[result][0]["c"] == "1":
-                t = 0
-                u = 1
-            elif dataset[result][0]["c"] == "2":
-                t = 0 
-                u = 1
+            if "r" in result:
+                if dataset[result][0]["c"] == "1":
+                    client1 = 1
+                    client2 = 0
+                elif dataset[result][0]["c"] == "2":
+                    client1 = 0
+                    client2 = 1
+            else:
+                if dataset[result][0]["c"] == "1":
+                    client1 = 0
+                    client2 = 1
+                elif dataset[result][0]["c"] == "2":
+                    client1 = 1
+                    client2 = 0
+
             if len(dataset[result]) < 2:
                 print("Not complete error for algorithm: " + base_algo + " in combination with " + sub_algo) 
                 return False
-            if dataset[result][t]["algo"] == base_algo and dataset[result][u]["algo"] == sub_algo:
-                if dataset[result][t]["loss"] == loss and dataset[result][t]["delay"] == delay: #The other one will always have the same delay and loss
-                    base_plot[dataset[result][t]["run"]] = [int(x) for x in dataset[result][t]["bandwidth"].split(" ")]
-                    sub_plot[dataset[result][u]["run"]] = [int(x) for x in dataset[result][u]["bandwidth"].split(" ")]
+            if dataset[result][client1]["algo"] == base_algo and dataset[result][client2]["algo"] == sub_algo:
+                if dataset[result][client1]["loss"] == loss and dataset[result][client1]["delay"] == delay: #The other one will always have the same delay and loss
[…]

### 2017-12-08 parser/parser.py
Added run 3 and 4 support

@@ -1,8 +1,9 @@
 import matplotlib.pyplot as plt
+from matplotlib.font_manager import FontProperties
 import numpy as np
 import os
 
-results = open("newres", "r")
+results = open("newresnew", "r")
 
 algorithms = ['cubic', 'ctcp', 'dctcp', 'bic', 'bbr']
 delays = [8, 80, 150, 220, 290]
@@ -16,7 +17,7 @@ def plot_graph(dataset, base_algo, sub_algo, loss, delay):
     sub_plot = {} 
     
     for result in dataset:
-        if dataset[result][0]["run"] == "1" or dataset[result][0]["run"] == "2": #Always combined same run in dict
+        if dataset[result][0]["run"] == "1" or dataset[result][0]["run"] == "2" or dataset[result][0]["run"] == "3" or dataset[result][0]["run"] == "4": #Always combined same run in dict
             #Yes, the if statement below is pretty useless. As of now. Not sure if this will have a purpose. 
             if "r" in result:
                 if dataset[result][0]["c"] == "1":
@@ -32,7 +33,7 @@ def plot_graph(dataset, base_algo, sub_algo, loss, delay):
                 elif dataset[result][0]["c"] == "2":
                     client1 = 1
                     client2 = 0
-
+            
             if len(dataset[result]) < 2:
                 print("Not complete error for algorithm: " + base_algo + " in combination with " + sub_algo) 
                 return False
@@ -43,9 +44,8 @@ def plot_graph(dataset, base_algo, sub_algo, loss, delay):
     
     if "1" not in base_plot:
         return False
-
-    print("Bandwidth points " + base_algo +  " per 1 second: " + str(base_plot))
-    print("Bandwidth points " + sub_algo + " per 1 second: " + str(sub_plot)) 
+    #print("Bandwidth points " + base_algo +  " per 1 second: " + str(base_plot))
+    #print("Bandwidth points " + sub_algo + " per 1 second: " + str(sub_plot)) 
     base_y = np.array(base_plot["1"])
     sub_y = np.array(sub_plot["1"])
     base_algor, = plt.plot(base_y, label=base_algo, color="red")
@@ -54,13 +54,28 @@ def plot_graph(dataset, base_algo, sub_algo, loss, delay):
     if len(base_plot) > 1 and len(sub_plot) > 1:
         base_y_2 = np.array(base_plot["2"])
         sub_y_2 = np.array(sub_plot["2"])
[…]

### 2017-12-10 paper/paper.tex
Update paper.tex

@@
 
-	The created environment is shown in figure \ref{fig:setup1}.
+	The created testing environment is shown in figure \ref{fig:setup1}.
+This is a virtual representation of the network as the client node and random node are running on one physical device.
+The client node and random node are listening each on a different port, which is set by the controller described in section [ENTER SECTION].
+The servers are separate physical devices as one of the servers possible runs Windows when a TCP congestion control algorithm is only available for Windows.
+The software used for the servers and clients is Ubuntu 17.04.
+If Windows is used, the server will be running Windows 10.
 
@@
 	In scenario 1 the TCP congestion algorithm is tested for fairness in sharing bandwidth with other TCP connections using other TCP congestion algorithms.
-In order to test the fairness of congestion algorithms, the bandwidth allocation per algorithm is measured.
+In order to test the fairness of congestion algorithms, the bandwidth allocation per second per algorithm is measured.
 It can be determined whether a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more or less bandwidth to that specific TCP connection.

### 2017-12-11 paper/paper.tex
Minor correction on Methods

@@
 	In order to select the most common TCP congestion algorithms desk research will be performed.
-Results of scientific reports will show which algorithms are used more often than other algoritms.
+This desk research will show which algorithms are used more often than other algoritms by looking for the default algorithms in Operating Systems.
 Based on those results a list of TCP congestion algorithms will be composed.
@@
 Those parameters are estimates defined by the BBR algorithm.
-\\texttt{BtlBw} contains the bandwidth available to data transmissions ...
+\texttt{BtlBw} contains the bandwidth available to data transmissions ...
+\\
+\\
+
 		- Includes characteristics of the algorithms. To what do the TCP congestion algorithms respond according to literature? And how do they deal with that (If available in literature).

### 2017-12-11 paper/paper.tex
Added TCP congestion algorithm characteristics

@@
 		\subsubsection{Algorithm Characteristics}
-		The BBR algorithm depends on the parameters BBR.BtlBw and BBR.RTprop.
+		The BBR algorithm depends on parameters $BtlBw$ and $RT_{prop}$ \cite{bbr-congestion}.
 Those parameters are estimates defined by the BBR algorithm.
-\texttt{BtlBw} contains the bandwidth available to data transmissions ...
-\\
-\\
+$BtlBw$ specifies the bandwidth at the slowest link in each direction and $RT_{prop}$ is the round-trip propagation time.
+$RT_{prop} = RTT - queuing delay - processing delay$, which results in the minimum amount of time for round trip propagation in case there are no queueing or processing delays.
+Using those two parameters, the Bandwidth Delay Product (BDP) is calculated, which is the maximum possible amount of data being sent in a network: $BDP = BtlBw * RT_{prop}$.
+As long as the amount of data in transit is less than the $BDP$ and $BtlBw$, the delivery rate is increased.
+If the amount of data in transit is equal to $BtlBw$ the delivery rate can not go up anymore.
+The $RTT$ can never be lower than $RT_{prop}$...
+explain more on delay perhaps?.
+
+		The CTCP algorithm is a combination of a loss-based and delay-based congestion protocol \cite{compound-tcp-congestion}.
+For the loss-based component, the variable conventional congestion window, $cwnd$, was introduced.
+For the delay-based component, the variable delay window, $dwnd$, was introduced.
+The TCP sending window is determined from: $win = min(cwnd + dwnd, awnd)$, where $awnd$ is the window advertised by the receiver.
+For $cwnd$ on the arrival of an ACK: $cwnd = cwnd + 1/win$.
+The $dwnd$ variable is initially set to zero and is only set when the congestion avoidance is active.
+$dwnd$ is calculated differently for several scenarios.
+If the network path is underutilized: $dwnd(t+1) = dwnd(t) + \alpha*dwnd(t)^{k} - 1$.
+When the network path is congested: $dwnd(t+1) = dwnd(t) - \eta*diff$.
+When packet loss occurs: $dwnd(t+1) = dwnd(t)(1-\beta) - cwnd/2$.
+$k$, $\alpha$, $\eta$ and $\beta$ are tunable parameters for optimization of the algorithm.
+$\eta$ defines the rapidness of reduction of the window when congestion is detected.
+$diff$ is calculated from the expected throughput and the actual throughput, which are again calculated from the window size and Round Trip Time (RTT).
+
+		The DCTCP algorithm consists of three components \cite{dctcp-congestion-original}:
+		\begin{enumerate}
+			\item Simple Marking at the Switch.
+Based on the marking treshold $K$ a packet is marked with the CE flag.
+This only occurs when the queue occupancy is greater than K at arrival time.
+The marking treshold $K$ specifies the minimum value of queue occupancy.
+			\item ECN-Echo at receiver.
+The receiver running the DCTCP algorithm ACKs every packet and sets the ECN-Echo flag only if the packet was marked with CE.
+The ECN-Echo flag is used to notify the sender of congestion in the network.
+			\item Controller at the sender.
+The sender has an estimate of the fraction of packets that are marked with CE.
+This is based on the actual fraction of packets that were marked, the weight given to the new estimate compared to the past weight.
+This results in $\alpha$.
+$\alpha$ is used to indicate the congestion on a network with a maximum of value 1, indicating a congested network, and a minimum value of 0, indicating no congestion in the network.
+DCTCP cuts the window size using $\alpha$: $cwnd = cwnd * (1 - \alpha/2)$.
+This results in a low queue length, but ensuring high throughput.\\
+		\end{enumerate}
 

### 2017-12-11 paper/paper.tex
paper: abstract

--- 2017-12-11 paper/paper.tex
@@
 
-TCP congestion control algorithms aim to divide the available bandwidth fairly
-among network users. In datacenters, many tenants share a limited amount of bandwidth.
-If too many tenants TODO luc
+In datacenters, multiple tenants share a finite amount of bandwidth. When an
+uplink is saturated, packets that overflow available buffer space will be
+dropped and have to be retransmitted by the sender. When one tenant chooses to
+retransmit faster than others, they have an advantage as they will make it
+though more quickly. TCP is a connection-oriented protocol which employs
+congestion control algorithms which aim to divide the available bandwidth
+equally among network users.
+
+Some congestion control algorithms are more aggressive than others, and it has
+been reported that some algorithms cause others to back off excessively.
+Tenants may, perhaps inadvertedly, cause unfair bandwidth distributions during
+times of congestion.
+
+In this research, we compare five popular algorithms under different network
+conditions to see how they influence each other. Among those are the newly
+developed BBR and DCTCP. We found TODO
 

### 2017-12-11 paper/paper.tex
Added BIC algorithm

--- 2017-12-11 paper/paper.tex
@@
 
-		- Includes characteristics of the algorithms. To what do the TCP congestion algorithms respond according to literature? And how do they deal with that (If available in literature).
+		The BIC algorithm controls its windows based on the size of windows \cite{bic-tcp-congestion}. It consists of two components:
+		\begin{enumerate}
+			\item Binary search increase.
+This algorithm computes the midpoint between the maximum window size $W_{max}$ and current minimum window size $W_{min}$.
+If packet loss occurs, $W_{max}$ is set to the midpoint.
+This process repeats until $W_{max}$ and $W_{min}$ are smaller than the minimum increment $S_{min}$.
+			\item Additive increase runs as an addition to the binary search increase.
+If $W_{max} - midpoint > S_{max}$, the window size is increased by $S_{max}$ until $W_{max} - midpoint < S_{max}$.
+$S_{max}$ is the preset maximum increment.
+		\end{enumerate}
+In case the current window size becomes greater than $W_{max}$ a slow start strategy is performed to determine a new $W_{max}$: $W_{max} + \alpha * S_{min}$, where $\alpha$ is a multiplier.
+The multiplier is increased by one until $S_{min} * \alpha >= S_{max}$ or when losses occur.
+In the case of loss a new value for $W_{max}$ will be set.
+
+		The CUBIC algorithm
 

### 2017-12-11 paper/paper.tex
Added CUBIC algorithm

--- 2017-12-11 paper/paper.tex
@@
 
-		The CUBIC algorithm
+		The CUBIC algorithm is a successor of the BIC algorithm \cite{cubic-tcp-congestion}.
+In case of a loss event $W_{max}$, which is the maximum window size, is registered.
+This is followed by a multiplicative decrease of the congestion window by factor $\beta$.
+If an ACK in congestion avoidance is received and the current window size $cwnd$ is less than $W_{max}$, the algorithm uses the concave profile to increase the window size according to the formula: $cwnd = \frac{W(t+RTT)-cwnd}{cwnd}$.
+If $cwnd$ is larger than $W_{max}$, the convex profile is used to increase the window size according to the same formula as the concave profile.
+In both formulas $t$ is the time elapsed from the last window reducation.
 

### 2017-12-11 paper/paper.tex
paper: changed sentence in sub:TCP_Congestion_Algorithms and added another ref

@@
 BIC was the default congestion algorithm in earlier Linux distributions from kernel version 2.6.8 until 2.6.19 \cite{bic-kernel-version} \cite{cubic-kernel-version}.
-DCTCP and BBR will also be tested as those protocols are newer.
-DCTCP is specifically designed for datacenters \cite{dctcp-congestion} and BBR is a new congestion protocol not yet widely deployed \cite{bbr-congestion}.
+DCTCP and BBR will also be tested as those protocols are new and upcoming---they are included in Linux since 4.9 (December 2016) \cite{linux-bbr}.
+DCTCP is specifically designed for datacenters \cite{dctcp-congestion} and BBR is a new algorithm meant for general use \cite{bbr-congestion}.
 

### 2017-12-11 paper/paper.tex
m. paper.tex: remove trailing spacing
m. paper.tex: unindent sub(sub)sections as discussed

@@
 \section{Method}
+
 During this research a performance test of various TCP congestion algorithms will be performed.
@@
 \subsection{Selecting TCP congestion algorithms}
+
 In order to select the most common TCP congestion algorithms desk research will be performed.
@@
 \subsection{Comparing Performance}
+
 One scenario will be used to compare the performance of the TCP congestion algorithms:
+
 \begin{enumerate}
@@
 \subsubsection{Determining the Packet loss and Delay}
+
 The dataset of Verizon \cite{verizon-latency} is used to calculate the packet loss and delay used in this research.
@@
 \subsubsection{Performing time-based automated tests}
+
 The tests will be performed using scripts written in Python [\ref{appendix:python}].
@@
 \subsubsection{Testing hardware}
+
 The specifications of the hardware used for the tests are shown in table \ref{table:spec1}, table \ref{table:spec2} and table \ref{table:spec3}.
@@
 \subsection{Traffic Management}
+
 Desk research will be conducted to gather information related to traffic management.
@@
 \section{Results}
+
 Type some text here.
@@
 \subsection{TCP Congestion Algorithms}
+
 The most common TCP congestion algorithms are CTCP, CUBIC and BIC.
@@
 
-		\subsubsection{Algorithm Characteristics}
+\subsubsection{Theoretical Algorithm Characteristics}
+
 The BBR algorithm depends on parameters $BtlBw$ and $RT_{prop}$ \cite{bbr-congestion}.
@@
 The DCTCP algorithm consists of three components \cite{dctcp-congestion-original}:
+
 \begin{enumerate}
@@
 The BIC algorithm controls its windows based on the size of windows \cite{bic-tcp-congestion}. It consists of two components:
+
 \begin{enumerate}
@@
 \subsubsection{Performance comparison}
+
 - Includes results of tests.
@@
 \subsection{Traffic Management}
+
 - Includes what is available for trafic management in a datacenter.
 - Measures to take to equally allocate bandwidth to customers.
+
 \subsubsection{Quality of Service}

### 2017-12-11 paper/paper.tex
Minor changes to methods

@@
 This scenario applies to multitentant datacenter situations, where a user owns a server in the datacenter network and a node outside of the datacenter network.
-In this test the user node and random node(s) are at the receiving side and the user server and random server at the sending side.
+In this test the clients are at the receiving side and the user server and random server at the sending side.
 The tests will be performed from a multitenant datacenter perspective.
@@
 The created testing environment is shown in figure \ref{fig:setup1}.
-This is a virtual representation of the network as the client node and random node are running on one physical device.
-The client node and random node are listening each on a different port, which is set by the controller described in section [ENTER SECTION].
+The clients are located on one physical device and are listening each on a different port, which is set by the controller described in section [TO DO: ENTER SECTION].
 The servers are separate physical devices as one of the servers possible runs Windows when a TCP congestion control algorithm is only available for Windows.
@@
 
-In scenario 1 the user server uses the same algorithm as the user node and the random server uses the same algorithm as the random node.
+In scenario 1 the user server uses the same algorithm as the first client and the random server uses the same algorithm as the second client.
 Tests in this scenario cover behavioral differences between TCP congestion algorithms when a user owns a node in the network of a multitenant datacenter.
@@
 	\centering
-	\caption{Switch 1 configuration}
+	\caption{Switch configuration}
 	\begin{tabular}[H]{ | l | l | }
@@
 
-\begin{table}[H]
-	\centering
-	\caption{Switch 2 configuration}
-	\begin{tabular}[H]{ | l | l | }
-	\hline
-	\textbf{Hardware} & \textbf{Specification} \\
-	\hline  Model & Sitecom Gigabit Router 300N-XR\\
-	\hline  Ports & 4x Gigabit Ethernet\\
-	\hline
-	\end{tabular}
-	\label{table:spec3}
-\end{table}
-
 \subsection{Traffic Management}
@@
 
-Type some text here.
+TO DO: Type some text here.
 

### 2017-12-14 parser/parser.py
Parser.py now generates HTML pages and displays MiB/s instead of bytes/s

@@ -1,3 +1,7 @@
+#!/usr/bin/env python
+
+# Dependency: python-matplotlib
+
 import matplotlib.pyplot as plt
 from matplotlib.font_manager import FontProperties
 import numpy as np
@@ -37,39 +41,49 @@ def plot_graph(dataset, base_algo, sub_algo, loss, delay):
             if len(dataset[result]) < 2:
                 print("Not complete error for algorithm: " + base_algo + " in combination with " + sub_algo) 
                 return False
+
             if dataset[result][client1]["algo"] == base_algo and dataset[result][client2]["algo"] == sub_algo:
                 if dataset[result][client1]["loss"] == loss and dataset[result][client1]["delay"] == delay: #The other one will always have the same delay and loss
-                    base_plot[dataset[result][client1]["run"]] = [int(x) for x in dataset[result][client1]["bandwidth"].split(" ")]
-                    sub_plot[dataset[result][client2]["run"]] = [int(x) for x in dataset[result][client2]["bandwidth"].split(" ")]
+                    if dataset[result][client1]["bandwidth"].strip() == '' or dataset[result][client2]["bandwidth"].strip() == '':
+                        print('Warn: set without bandwidth, testnum={}'.format(dataset[result][client1]["testnum"]))
+                        continue
+
+                    #base_plot[dataset[result][client1]["run"]] = [int(x) for x in dataset[result][client1]["bandwidth"].strip().split(" ")]
+                    #sub_plot[dataset[result][client2]["run"]] = [int(x) for x in dataset[result][client2]["bandwidth"].strip().split(" ")]
+                    base_plot[dataset[result][client1]["run"]] = [float(x)/(2**20) for x in dataset[result][client1]["bandwidth"].strip().split(" ")]
+                    sub_plot[dataset[result][client2]["run"]] = [float(x)/(2**20) for x in dataset[result][client2]["bandwidth"].strip().split(" ")]
     
     if "1" not in base_plot:
         return False
+
+    filename = base_algo + "/" + base_algo + "_vs_" + sub_algo + "_" + delay + "_" + loss + ".png"
+    print(filename)
+
     #print("Bandwidth points " + base_algo +  " per 1 second: " + str(base_plot))
[…]

### 2017-12-14 parser/parser.py
m. parser.py: remove exit(0) used for debugging

@@ -196,7 +196,6 @@ for delay in delays:
     for loss in losses:
         create_html_delay_loss(delay, loss)
 
-exit(0)
 # Plot graph for base_algorithm and sub_algorithm
 done = {}
 for delay in delays:

### 2017-12-14 paper/paper.tex
Minor change
paper: updated+improved experimental setup description, and other minor changes

@@
 
-One scenario will be used to compare the performance of the TCP congestion algorithms:
+We will analyse one common scenario: multiple servers of different owners
+inside a multitenant datacenter, sends data to clients outside the datacenter.
+This could be users downloading software, the bursts while streaming a video
+or movie, or large resources on a webpage.
 
-\begin{enumerate}
-	\item The user owns multiple endpoints in the network.
-In this case a congestion algorithm is applied at both the client and server side.
-This scenario applies to multitentant datacenter situations, where a user owns a server in the datacenter network and a node outside of the datacenter network.
-In this test the clients are at the receiving side and the user server and random server at the sending side.
-The tests will be performed from a multitenant datacenter perspective.
-\end{enumerate}
+We will create a bottleneck on the path to the user to see how different
+congestion algorithms handle this. The connection will be full duplex and have
+symmetrical speeds, to avoid any congestion in sending acknowledgements.
+Because the server is sending the large volume of data and the client is
+responding with the much smaller acknowledgements, the server will be the one
+applying the congestion control. Therefore we vary the algorithm settings of
+the server and not the client. The client will always be a GNU/Linux machine
+using CUBIC, but we assume (due to the aforementioned reasons) that this does
+not matter.
 
-The created testing environment is shown in figure \ref{fig:setup1}.
-The clients are located on one physical device and are listening each on a different port, which is set by the controller described in section [TO DO: ENTER SECTION].
-The servers are separate physical devices as one of the servers possible runs Windows when a TCP congestion control algorithm is only available for Windows.
-The software used for the servers and clients is Ubuntu 17.04.
-If Windows is used, the server will be running Windows 10.
+The bottleneck is created by having one receiving host, listening on different
+ports, and two transmitting hosts. Each network interface is 1 gigabit per
+second, thus the servers will have to limit their speed in order to avoid
+flooding the client. This setup is shown in figure \ref{fig:setup1}.
+
+We use three physical hosts as servers: two running Ubuntu 17.04 and one
+running Microsoft Windows 10. In each experiment, only two hosts will be used.
+The Windows system is used because there is no working CTCP
+implementation---the default algorithm in Windows since Vista---available for a
+modern Linux kernel.
 
@@
 
-In scenario 1 the user server uses the same algorithm as the first client and the random server uses the same algorithm as the second client.
-Tests in this scenario cover behavioral differences between TCP congestion algorithms when a user owns a node in the network of a multitenant datacenter.
+We will test the TCP congestion algorithm for fairness in sharing bandwidth
+with other TCP connections using other congestion algorithms. In order to test
+the fairness, the bandwidth usage per second per algorithm is measured. If two
+different algorithms use similar amounts of bandwidth, they can be said to be
+fair relative to each other.
 
-In scenario 1 the TCP congestion algorithm is tested for fairness in sharing bandwidth with other TCP connections using other TCP congestion algorithms.
-In order to test the fairness of congestion algorithms, the bandwidth allocation per second per algorithm is measured.
-It can be determined whether a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more or less bandwidth to that specific TCP connection.
-
-The tests conducted on the algorithms have two environment variables: packet loss and delay.
+The tests conducted with the algorithms have two environment variables: packet loss and delay.
 Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions \cite{linux-netem}.
@@
 Therefore the packet loss and delay are applied to the incoming interfaces of the clients.
-Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables [SRC].
+Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables [TODO: SRC].
 The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.
+
+Each test is run at least twice. When results are conflicting or unexpected, we
+ran a test more often and on different hardware.
 
@@
 
-TO DO: Type some text here.
+TODO: Type some text here.
 
@@
 
-\subsubsection{Theoretical Algorithm Characteristics}
+\subsection{Theoretical Algorithm Characteristics}
 
@@
 $\eta$ defines the rapidness of reduction of the window when congestion is detected.
+$\beta$ is the factor used for the multiplicative decrease after loss is detected.
 $diff$ is calculated from the expected throughput and the actual throughput, which are again calculated from the window size and Round Trip Time (RTT).
@@
 
-\subsubsection{Performance comparison}
+\subsection{Experiments' Results}
 
-- Includes results of tests.
+TODO
 

### 2017-12-14 parser/parser.py
parser.py html generator: show datapoints when hovering over image

@@ -17,6 +17,9 @@ result_data = {}
 counter = 1
 
 def plot_graph(dataset, base_algo, sub_algo, loss, delay):
+    global datastrs
+
+    datastr = ''
     base_plot = {}
     sub_plot = {} 
     
@@ -52,11 +55,14 @@ def plot_graph(dataset, base_algo, sub_algo, loss, delay):
                     #sub_plot[dataset[result][client2]["run"]] = [int(x) for x in dataset[result][client2]["bandwidth"].strip().split(" ")]
                     base_plot[dataset[result][client1]["run"]] = [float(x)/(2**20) for x in dataset[result][client1]["bandwidth"].strip().split(" ")]
                     sub_plot[dataset[result][client2]["run"]] = [float(x)/(2**20) for x in dataset[result][client2]["bandwidth"].strip().split(" ")]
+                    datastr += 'base' + dataset[result][client1]["run"] + ': ' + str(base_plot[dataset[result][client1]["run"]]) + '\n'
+                    datastr += 'base' + dataset[result][client2]["run"] + ': ' + str(base_plot[dataset[result][client2]["run"]]) + '\n'
     
     if "1" not in base_plot:
         return False
 
     filename = base_algo + "/" + base_algo + "_vs_" + sub_algo + "_" + delay + "_" + loss + ".png"
+    datastrs[base_algo + "_vs_" + sub_algo + "_" + delay + "_" + loss] = datastr
     print(filename)
 
     #print("Bandwidth points " + base_algo +  " per 1 second: " + str(base_plot))
@@ -141,8 +147,13 @@ def create_html(algo1, algo2, recursed=False):
     exists = 0
     for loss in losses:
         for delay in delays:
-            f.write('<td>{algo1} vs {algo2} with {delay}ms delay and {loss}% loss<br><img src="{algo1}_vs_{algo2}_{delay}_{loss}.png" width=640 height=480></td>'
-                .format(algo1=algo1, algo2=algo2, delay=delay, loss=loss))
+            try:
+                datastr = datastrs['{}_vs_{}_{}_{}'.format(algo1, algo2, delay, loss)]
+            except:
+                datastr = '(unknown)'
+
+            f.write('<td>{algo1} vs {algo2} with {delay}ms delay and {loss}% loss<br><img src="{algo1}_vs_{algo2}_{delay}_{loss}.png" title="{datastr}" width=640 height=480></td>'
+                .format(algo1=algo1, algo2=algo2, delay=delay, loss=loss, datastr=datastr))
 
[…]

### 2017-12-14 parser/parser.py
parser.py fix: add subplot instead of 2x baseplot

@@ -56,7 +56,7 @@ def plot_graph(dataset, base_algo, sub_algo, loss, delay):
                     base_plot[dataset[result][client1]["run"]] = [float(x)/(2**20) for x in dataset[result][client1]["bandwidth"].strip().split(" ")]
                     sub_plot[dataset[result][client2]["run"]] = [float(x)/(2**20) for x in dataset[result][client2]["bandwidth"].strip().split(" ")]
                     datastr += 'base' + dataset[result][client1]["run"] + ': ' + str(base_plot[dataset[result][client1]["run"]]) + '\n'
-                    datastr += 'base' + dataset[result][client2]["run"] + ': ' + str(base_plot[dataset[result][client2]["run"]]) + '\n'
+                    datastr += 'sub' + dataset[result][client2]["run"] + ': ' + str(sub_plot[dataset[result][client2]["run"]]) + '\n'
     
     if "1" not in base_plot:
         return False

### 2017-12-14 paper/paper.tex
Wrote down some experiments' results

@@
 
-TODO
+In our results, we see that all algorithms clearly respond to the simulated
+network circumstances, but not very much to each other. Most algorithms are
+fair towards all other algorithms.
+
+CTCP is the only algorithm that responds excessively to others. Alone, it
+reaches the full gigabit speed; but once any other algorithm runs at the same
+time (i.e. a normal test is performed), CTCP's speed consistently drops. It
+then reaches around 55 megabits ($\pm$5), while the other algorithms reach at %TODO: check 55+-5
+least 800 with good network conditions (8 milliseconds latency, 0\% loss).
+
+On top of this, CTCP also lowers its speed in response to the simulated network
+circumstances: with 1.2\% packet loss and 8 milliseconds latency, it will reach
+around 25 megabits, and with 290 milliseconds latency and no loss, it reaches
+around 1.5 megabits.
+
+BBR is unique in that it barely responds to packet loss but only to delay: even
+with 1.2\% loss, as long as the delay is low (8 milliseconds in our case), it
+reaches 850$\pm$20 megabits per second. Other algorithms will reach much lower
+speeds, but we see no signs of BBR being the cause of that. In tests that
+involve two algorithms that are neither BBR nor CTCP, speeds are not higher
+than with BBR being one of the two algorithms. %TODO check this one (this was gut feeling, but looking at the data, there might be a slight effect that BBR has on things)
 

### 2017-12-14 paper/paper.tex
Added some figures

@@
 network circumstances, but not very much to each other. Most algorithms are
-fair towards all other algorithms.
+fair towards all other algorithms and will use their half of the connection.
+See figures \ref{fig:fair-1} and \ref{fig:fair-2}.
+
+\begin{figure}[H]
+	\makebox[\textwidth]{
+		\centering
+		\begin{minipage}{0.5\paperwidth}
+			\centering
+			\includegraphics[width=0.9\textwidth]{figs/cubic-bic-8-0.png}
+			\caption{CUBIC vs BIC with 8ms latency and 0\% loss}
+			\label{fig:fair-1}
+		\end{minipage}\hfill
+		\begin{minipage}{0.5\paperwidth}
+			\centering
+			\includegraphics[width=0.9\textwidth]{figs/dctcp-bic-8-0.png}
+			\caption{DCTCP vs BIC with 8ms latency and 0\% loss}
+			\label{fig:fair-2}
+		\end{minipage}
+	}
+\end{figure}
 
@@
 
-On top of this, CTCP also lowers its speed in response to the simulated network
+\begin{figure}[H]
+	\centering
+		\includegraphics[scale=0.4]{figs/ctcp-bic-8-0.png}
+		\caption{TODO}
+	\label{fig:packet-loss-chart}
+\end{figure}
+
+Additionally, CTCP also lowers its speed in response to the simulated network
 circumstances: with 1.2\% packet loss and 8 milliseconds latency, it will reach
 around 25 megabits, and with 290 milliseconds latency and no loss, it reaches
-around 1.5 megabits.
+around 1.5 megabits. In neither case does it reach the original 55 megabits.
 
@@
 reaches 850$\pm$20 megabits per second. Other algorithms will reach much lower
-speeds, but we see no signs of BBR being the cause of that. In tests that
-involve two algorithms that are neither BBR nor CTCP, speeds are not higher
-than with BBR being one of the two algorithms. %TODO check this one (this was gut feeling, but looking at the data, there might be a slight effect that BBR has on things)
+speeds (24 megabits is not uncommon), but we see no signs of BBR being the
+cause of that. In tests with BBR nor CTCP, speeds are not higher than when BBR
+is one of the algorithms being tested.
 

### 2017-12-14 paper/paper.tex
paper: minor improvements

@@
 
+TODO in the next section blah, then blah, and finally blah.
+
 \section{Related work}
@@
 
-During this research a performance test of various TCP congestion algorithms will be performed.
+During this research, performance comparison tests of various TCP congestion algorithms will be performed.
 The most common TCP congestion algorithms will be determined through desk research.
-The tests will be conducted in a private environment.
+The tests will be conducted in an isolated environment.
 Based on the results of those tests, recommendations will be given related to the measures a datacenter has to take to keep the traffic equally allocated over customers.
@@
 
-TODO: Type some text here.
+In this section we will first elaborate on which algorithms we looked into, next we look at their internal workings, and finally we show the results of our experiments.
 
@@
 CTCP is the only algorithm that responds excessively to others. Alone, it
-reaches the full gigabit speed; but once any other algorithm runs at the same
+reaches the full gigabit speed; but if any other algorithm runs at the same
 time (i.e. a normal test is performed), CTCP's speed consistently drops. It
@@
 least 800 with good network conditions (8 milliseconds latency, 0\% loss).
+See figure \ref{fig:ctcp-1} for an example of BIC vs. CTCP.
 
@@
 		\includegraphics[scale=0.4]{figs/ctcp-bic-8-0.png}
-		\caption{TODO}
-	\label{fig:packet-loss-chart}
+		\caption{CTCP vs. BIC with 8ms latency and 0\% loss}
+	\label{fig:ctcp-1}
 \end{figure}
@@
 is one of the algorithms being tested.
+
+The full set of results is available online\cite{git-lsproj}.
 

### 2017-12-20 paper/paper.tex
Updated abstract

@@
 though more quickly. TCP is a connection-oriented protocol which employs
-congestion control algorithms which aim to divide the available bandwidth
-equally among network users.
+congestion control algorithms. These algorithms aim to divide the available
+bandwidth equally between network flows.
 
@@
 been reported that some algorithms cause others to back off excessively.
-Tenants may, perhaps inadvertedly, cause unfair bandwidth distributions during
+Tenants may, perhaps inadvertedly, cause inequal bandwidth distributions during
 times of congestion.
 
-In this research, we compare five popular algorithms under different network
-conditions to see how they influence each other. Among those are the newly
-developed BBR and DCTCP. We found TODO
+In this research, we compare five algorithms under different network conditions
+to assess how they influence each other. We found that only CTCP (Windows)
+backs off excessively; all other algorithms behave fairly. The results for CTCP
+are anomalous compared to all other algorithms and we conclude that for this
+algorithm, future work is needed for a reliable assessment.
 

### 2017-12-20 paper/paper.tex
Added traffic shaping

--- 2017-12-14 paper/paper.tex
@@
 \subsection{Traffic Management}
+We propose two possible solutions to manage traffic from a multitenant datacenter to customers when unfair TCP congestion control algorithms are used: Traffic shaping based Monitoring and detection of unaccepted TCP congestion algorithms in the network.
 
-- Includes what is available for trafic management in a datacenter.
-- Measures to take to equally allocate bandwidth to customers.
+\subsubsection{Traffic shaping based Monitoring}
+With the use of a hypothetical TCP congestion control algorithm that is unfair to other TCP congestion control algorithms, the bandwidth allocation to one customer can be higher to other customers using the default congestion control algorithms.
+For example: when a TCP congestion control algorithm shows an aggressive increase and keeps pushing data over the network, other fair TCP congestion algorithms will reduce the sending rate of the data transfer as packet loss occurs.
+One important condition for this unfair congestion control algorithm is that it should not be agressive such that it behaves in a Denial of Service manner.
 
-\subsubsection{Quality of Service}
+To solve this issue, we propose a monitoring solution which monitors the bandwidth usage per server and controls the priority for the traffic sent by a server.The solution uses traffic shaping to control the priority of the traffic per server:
+\begin{enumerate}
+	\item{The monitoring solution detects unfair bandwidth allocation on a congested link. It detects a congested link by checking for delay and packet loss in network segments.}
+	\item{The monitoring solution calculates per server whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage.
+$\alpha$ is deviated bandwidth and is set according to an acceptable deviation value.
+The following formula is used to calculate the standard deviation: $s_{i} = \sqrt{\frac{\sum_{i=1}^N (x_i - \overline{x})^2}{N-1} }$.
+$N$ is the amount of active servers on the link.
+$\overline{x}$ is the average amount of bandwidth for all active servers on the link.
+$x_i$ is the bandwidth of the current server and $s_{i}$ shows the standard deviation in bandwidth for server $i$.
+If $s_{i}$ is a negative value, no priority change will be applied as the bandwidth usage is lower than average for the server.
+If $s_{i} <= \alpha$, then no priority change will be applied.
+If $s_{i} > \alpha$, the priority will be changed to a lower value.}
+	\item{The monitoring solution provides information of the servers where the priority will be set to a lower value to the traffic shaping controller.}
+	\item{The traffic shaping controller sets the priority for the servers to a lower value, so that the traffic coming from the servers that use $\alpha$ more than average bandwidth has less preference than traffic closer to the average bandwidth and lower.}
+	\item{The monitoring solution periodically checks if congestion and delay are not detected.
+If both are not detected, the traffic shaping policy is reversed again as more bandwidth is available to allocate to customers.}
+\end{enumerate}
+
+To do: Not sure about last enumerate item.
+
+\subsubsection{Detection of unaccepted TCP congestion control algorithms}
 
@@
 \section{Conclusion}
+To do: Monitoring $\alpha$ value research. future work.
 

### 2017-12-20 paper/paper.tex
Update related work; give overview of sections in introduction

--- 2017-12-20 paper/paper.tex
@@
 
-TODO in the next section blah, then blah, and finally blah.
+In the next section, we will discuss related work. Section \ref{sec:method}
+describes our methods: how we selected algorithms to test, how we tested their
+performance, and how we chose network conditions to simulate. Section
+\ref{sec:results} describes the results of our research: the algorithms we
+chose to test are described in \ref{sub:chosen-algos}, a theoretical analysis
+of those is described in \ref{sub:algos-theoretical}, and experimental results
+can be found in \ref{sub:experiment-results}. We discuss our results in section
+\ref{sec:discussion} and draw conclusions in section \ref{sec:conclusion}.
 
@@
 
-% Copied from proposal
+In 2010, the algorithm DCTCP was described by Mohammad Alizadeh et
+al.\cite{dctcp-congestion-original} and published as
+RFC8257\cite{dctcp-congestion} in 2017. DCTCP is optimized for datacenters and
+provides a high-burst tolerance, low latency, and high throughput when the
+datacenter has a small part of the buffer available\cite{dctcp-congestion}.
 
-% todo make this logical
-In 2010, the algorithm DCTCP was described in an
-article\cite{dctcp-congestion-original} and published as RFC8257 in
-2017\cite{dctcp-congestion}. DCTCP is optimized for datacenters and provides a
-high-burst tolerance, low latency, and high throughput when the datacenter has
-a small part of the buffer available\cite{dctcp-congestion}. In 2016 another
-algorithm was proposed: BBR. This is a TCP congestion algorithm created by
-Google, which achieves higher bandwidths and lower latencies compared to other
-TCP congestion methods\cite{bbr-congestion}. A comparison of BBR with
-CUBIC\cite{bbr-congestion-comparison} shows that the BBR node pushes the CUBIC
-node away in bandwidth when using small buffers. The BBR node gets more
-bandwidth allocated than the CUBIC node.
+In 2016 another algorithm was proposed: BBR. This is a TCP congestion algorithm
+created by Google, which achieves higher bandwidths and lower latencies
+compared to other TCP congestion methods\cite{bbr-congestion}. A comparison of
+BBR with CUBIC\cite{bbr-congestion-comparison} shows that the BBR node pushes
+the CUBIC node away in bandwidth when using small buffers. The BBR node gets
+more bandwidth allocated than the CUBIC node.
 
-The TCP congestion control algorithm BBR is discussed in \cite{bbr-congestion}.
-The journal contains a performance test of the BBR algorithm and a comparison
-between the BBR algorithm and the CUBIC algorithm, where a noticeable
-difference in bandwidth allocation is shown when using small buffers.
+The TCP congestion control algorithm BBR is discussed by Neal Cardwell et
+al.\cite{bbr-congestion} The article contains a performance test of the BBR
+algorithm and a comparison between the BBR algorithm and the CUBIC algorithm,
+where a noticeable difference in bandwidth allocation is shown when using small
+buffers.
 
-In \cite{multiple-congestion} a performance analysis of multiple TCP congestion
-control algorithms is performed. This research does not include the new
-congestion control algorithms we want to analyze and does not include the
-datacenter context.
+A. Esterhuizen and A.E. Krzesinski\cite{multiple-congestion} analyse the
+performance of multiple TCP congestion control algorithms, specifically: Reno,
+BIC, CUBIC, HighSpeed TCP, TCP-Hybla, TCP-Illinois, TCP Low Priority,
+TCP-Vegas, TCP-Westwood, TCP-YeAH, and Scalable TCP.
 
-In \cite{dctcp-congestion-original} an algorithm is proposed to avoid
-congestion within datacenters. The authors describe an algorithm that maintains
-a small buffer occupancy and has an early detection of congestion.
-
-\section{Method}
+\section{Method}\label{sec:method}
 
@@
 
-\section{Results}
+\section{Results}\label{sec:results}
 
@@
 
-\subsection{TCP Congestion Algorithms}
+\subsection{TCP Congestion Algorithms}\label{sub:chosen-algos}
 
@@
 
-\subsection{Theoretical Algorithm Characteristics}
+\subsection{Theoretical Algorithm Characteristics}\label{sub:algos-theoretical}
 
@@
 
-\subsection{Experiments' Results}
+\subsection{Experiments' Results}\label{sub:experiment-results}
 
@@
 
-\section{Discussion}
+\section{Discussion}\label{sec:discussion}
 
-\section{Conclusion}
+\section{Conclusion}\label{sec:conclusion}
 

### 2017-12-20 paper/paper.tex

--- 2017-12-20 paper/paper.tex
@@
 
-To do: Not sure about last enumerate item.
+Todo: Not sure about last enumerate item.
 
@@
 
-\section{Discussion}
+\section{Discussion}\label{sec:discussion}
 
-\section{Conclusion}
+\section{Conclusion}\label{sec:conclusion}
 To do: Monitoring $\alpha$ value research. future work.

### 2017-12-20 paper/paper.tex
random fixes in paper

@@
 
-The specifications of the hardware used for the tests are shown in table \ref{table:spec1}, table \ref{table:spec2} and table \ref{table:spec3}.
-The switch shown in table \ref{table:spec3} has all router related functionalities, which might impact the network speed, turned off.
+The specifications of the hardware used for the tests are shown in tables \ref{table:spec1} and \ref{table:spec2}.
 
@@
 	\hline  Memory & 12GB\\
-	\hline  NIC & Intel Corporation 82579LM Gigabit Network Connection (rev 04)\\
+	\hline  NIC & Intel 82579LM Gigabit Network Connection (rev 04)\\
 	\hline	NIC Driver & e100e 3.2.6-k\\
@@
 \begin{enumerate}
-	\item{The monitoring solution detects unfair bandwidth allocation on a congested link. It detects a congested link by checking for delay and packet loss in network segments.}
-	\item{The monitoring solution calculates per server whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage.
+	\item The monitoring solution detects unfair bandwidth allocation on a congested link. It detects a congested link by checking for delay and packet loss in network segments.
+	\item The monitoring solution calculates per server whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage.
 $\alpha$ is deviated bandwidth and is set according to an acceptable deviation value.
@@
 If $s_{i} <= \alpha$, then no priority change will be applied.
-If $s_{i} > \alpha$, the priority will be changed to a lower value.}
-	\item{The monitoring solution provides information of the servers where the priority will be set to a lower value to the traffic shaping controller.}
-	\item{The traffic shaping controller sets the priority for the servers to a lower value, so that the traffic coming from the servers that use $\alpha$ more than average bandwidth has less preference than traffic closer to the average bandwidth and lower.}
-	\item{The monitoring solution periodically checks if congestion and delay are not detected.
-If both are not detected, the traffic shaping policy is reversed again as more bandwidth is available to allocate to customers.}
+If $s_{i} > \alpha$, the priority will be changed to a lower value.
+	\item The monitoring solution provides information of the servers where the priority will be set to a lower value to the traffic shaping controller.
+	\item The traffic shaping controller sets the priority for the servers to a lower value, so that the traffic coming from the servers that use $\alpha$ more than average bandwidth has less preference than traffic closer to the average bandwidth and lower.
+	\item The monitoring solution periodically checks if congestion and delay are not detected.
+If both are not detected, the traffic shaping policy is reversed again as more bandwidth is available to allocate to customers.
 \end{enumerate}

### 2017-12-20 paper/paper.tex
Added detection of unaccepted TCP congestion control algos

--- 2017-12-20 paper/paper.tex
@@
 \subsubsection{Detection of unaccepted TCP congestion control algorithms}
+In \cite{tcp-congestion-identification} a method to remotely identify used TCP congestion avoidance algorithms is described. The paper states that 
+they were able to identify all default TCP algorithms, but some non-default algorithms could not be identified.
+The tests performed are limited to remote web servers, so it requires more research to reliably implement such a method.
+If this method is deemed reliable, it can be useful to detect whether an unaccepted TCP congestion control algorithm is used as it will not match any of the default TCP algorithms used.
+Based on that the traffic of the unacceptable TCP congestion control algorithms can be shaped by lowering their priority.
 

### 2017-12-20 paper/paper.tex
Changes to monitoring solution

@@
 For example: when a TCP congestion control algorithm shows an aggressive increase and keeps pushing data over the network, other fair TCP congestion algorithms will reduce the sending rate of the data transfer as packet loss occurs.
-One important condition for this unfair congestion control algorithm is that it should not be agressive such that it behaves in a Denial of Service manner.
+One important condition for this unfair congestion control algorithm is that it should not be that agressive such that it behaves in a Denial of Service manner.
 
@@
 \begin{enumerate}
-	\item The monitoring solution detects unfair bandwidth allocation on a congested link. It detects a congested link by checking for delay and packet loss in network segments.
-	\item The monitoring solution calculates per server whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage.
+	\item The monitoring solution detects periodically whether the bandwidth in the link is fully utilized.
+If this is the case, the next step will be initiated.
+If not, it continues monitoring the link utilization.
+	\item The monitoring solution calculates periodically per server whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage.
 $\alpha$ is deviated bandwidth and is set according to an acceptable deviation value.
@@
 	\item The traffic shaping controller sets the priority for the servers to a lower value, so that the traffic coming from the servers that use $\alpha$ more than average bandwidth has less preference than traffic closer to the average bandwidth and lower.
-	\item The monitoring solution periodically checks if congestion and delay are not detected.
-If both are not detected, the traffic shaping policy is reversed again as more bandwidth is available to allocate to customers.
+	\item The monitoring solution periodically checks if the bandwidth in the link is still fully utilized.
+If this is not the case, the traffic shaping policy is reversed again by setting the priority to the default value.
+If this is the case, the periodical calculations of server bandwidth usage are continued.
 \end{enumerate}
-
-Todo: Not sure about last enumerate item.
 

### 2017-12-20 paper/paper.tex
Added future work + discussion

@@
 \subsubsection{Traffic shaping based Monitoring}
-With the use of a hypothetical TCP congestion control algorithm that is unfair to other TCP congestion control algorithms, the bandwidth allocation to one customer can be higher to other customers using the default congestion control algorithms.
+With the use of a hypothetical TCP congestion control algorithm that is unfair to other TCP congestion control algorithms, the bandwidth allocation to one customer can be higher than for other customers using the default congestion control algorithms.
 For example: when a TCP congestion control algorithm shows an aggressive increase and keeps pushing data over the network, other fair TCP congestion algorithms will reduce the sending rate of the data transfer as packet loss occurs.
@@
 \section{Discussion}\label{sec:discussion}
+The CTCP measurements were done on a Windows 10 machine.
+All other measurements were done on machines running Ubuntu 17.04.
+This might have introduced issues while measuring the CTCP algorithm versus any other congestion algorithm.
+CTCP did not show the expected behavior when running the tests.
+According to our measurements, all other algorithms are fair in sharing bandwidth with other algorithms.
+However, all measurements from CTCP show different results.
+We think this might be caused by the way Python sockets or Python calls are handled in Windows 10.
+We have performed the 0\% loss 0ms delay test on different hardware, which resulted in the same low value of bandwidth usage.
+When we used \texttt{iperf}, it reached the half of 1 Gbps versus other congestion algorithms as expected.
+We also increased the amount of bytes sent each time, where it reached 1 Gbps on its own.
+However, when testing it against another congestion algorithm it still showed the same behavior: performing with a low throughput.
+Therefore, we think the call handling of Python in Windows encounters more delay than Linux, but this requires more testing.
+
+\section{Future work}\label{sec:futurework}
+Running a mixed environment of Linux and Windows where Python is running introduces issues.
+We expect the call handling of Python in Windows encounters more delay than in Linux, but this requires more testing.
+The tests for CTCP might have to be performed again when full speeds can be reached by using the same settings.
+The CTCP tests are currently executed on Windows 10. This might also be performed on a Windows Server instance as it possibly handles the network in a different way.
+
+All tests described in this paper can be performed with more parallel connections, so that even when using a lot of parallel connections with two different congestion control algorithms, it is proven that the congestion control algorithms are fair.
+
+In this paper we have proposed a monitoring solution, which checks for the deviation between the server bandwidth on a link and the average bandwidth used by servers on a link.
+This solution requires performance testing, such that it is capable of running in a datacenter and not causing delay.
+Also an $\alpha$ value has to be determined that works in most environments, without ruling out servers that are not running unfair TCP congestion protocols.
 
 \section{Conclusion}\label{sec:conclusion}
-To do: Monitoring $\alpha$ value research. future work.
+To do: Mention items:
+\begin{enumerate}
+	\item{With current TCP congestion control algorithms behaving fair: No traffic control is currently needed & there is no impact on service and availability for customers}
+	\item{BBR is best performing algorithm of five tested}
+	\item{DCTCP is only suitable when both server and client are in same datacenter}
+	\item{Windows needs further testing}
+\end{enumerate}
 

### 2017-12-20 paper/paper.tex
Conclusion

@@
 \section{Conclusion}\label{sec:conclusion}
+Our main research question is: {\it How can a datacenter provide a reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the network?}
+
+The results have shown that the current default TCP congestion control algorithms are showing a fair behavior: The bandwidth is allocated equally among all nodes.
+Therefore, currently no traffic control is needed, unless someone would create an unfair TCP congestion control algorithm.
+This can be detected by one of the monitoring solutions proposed in this paper.
+
 To do: Mention items:

### 2017-12-20 paper/paper.tex
Conclusion

@@
 Therefore, currently no traffic control is needed, unless someone would create an unfair TCP congestion control algorithm.
-This can be detected by one of the monitoring solutions proposed in this paper.
+This can, after more research, be detected by one of the monitoring solutions proposed in this paper.
+As for the tested TCP congestion control algorithms, there is no impact on the service and availability for customers when using those algorithms mixed in one environment.
 
-To do: Mention items:
-\begin{enumerate}
-	\item{With current TCP congestion control algorithms behaving fair: No traffic control is currently needed & there is no impact on service and availability for customers}
-	\item{BBR is best performing algorithm of five tested}
-	\item{DCTCP is only suitable when both server and client are in same datacenter}
-	\item{Windows needs further testing}
+The BBR congestion control algorithm performs best when delay and loss occur of all TCP congestion control algorithms tested.
+DCTCP is only suitable when both the server and client are in the same datacenter, because the start up of the connection is slow when congestion in the network occurs.
+For datacenters it works well, because it keeps the queue lengths to a minimum, assuring a packet arrives at its destination.
 \end{enumerate}

### 2017-12-20 paper/paper.tex
Discussion and future work WIP

--- 2017-12-20 paper/paper.tex
@@
 
+Because of the exceptional results for CTCP, we performed the test again on
+different hardware. This yielded the same result. To verify that it is CTCP and
+not our software that performs differently on Windows, we setup a small
+experiment using \texttt{iperf}. The setup was the same as our original
+experimental setup with one sender on two hosts and two receivers on one host,
+only we manually started \texttt{iperf} bandwidth tests rather than doing it in
+an automated fashion with Python. In this instance, Windows performed normally
+and did not back off exceptionally when introducing another contender.
+
 BBR is unique in that it barely responds to packet loss but only to delay: even
@@
 \subsubsection{Detection of unaccepted TCP congestion control algorithms}
+
 In \cite{tcp-congestion-identification} a method to remotely identify used TCP congestion avoidance algorithms is described. The paper states that 
@@
 \section{Discussion}\label{sec:discussion}
-The CTCP measurements were done on a Windows 10 machine.
-All other measurements were done on machines running Ubuntu 17.04.
-This might have introduced issues while measuring the CTCP algorithm versus any other congestion algorithm.
-CTCP did not show the expected behavior when running the tests.
-According to our measurements, all other algorithms are fair in sharing bandwidth with other algorithms.
-However, all measurements from CTCP show different results.
-We think this might be caused by the way Python sockets or Python calls are handled in Windows 10.
-We have performed the 0\% loss 0ms delay test on different hardware, which resulted in the same low value of bandwidth usage.
-When we used \texttt{iperf}, it reached the half of 1 Gbps versus other congestion algorithms as expected.
-We also increased the amount of bytes sent each time, where it reached 1 Gbps on its own.
-However, when testing it against another congestion algorithm it still showed the same behavior: performing with a low throughput.
-Therefore, we think the call handling of Python in Windows encounters more delay than Linux, but this requires more testing.
+
+The CTCP measurements were done on a Windows 10 machine because there is no
+up-to-date implementation of CTCP for Linux. All other measurements were done
+on machines running Ubuntu 17.04. This might have introduced issues while
+measuring the CTCP algorithm versus any other congestion algorithm. According
+to our measurements, all other algorithms are fair in sharing bandwidth with
+each other; only CTCP consistently and excessively reduced its bandwidth
+allocation.
+
+Without any concurrent flows, Windows does reach gigabit speed. On different
+hardware, it showed the same, recessive behaviour. We think this might be
+caused by the way Python socket operations are handled in Windows. When
+repeating the test with \texttt{iperf}, which is written in C, this behaviour
+no longer occurs. Even when using the Python equivalent of
+\texttt{while(true)\{ socket.send(data); \}}, the bandwidth allocation did not
+change. We theorize that the difference is caused by something in the
+\texttt{send} call. An alternative theory is that Microsoft Windows for
+workstations have a different implementation than Microsoft Windows Server
+systems.
+
+BBR is the best-performing algorithm in our set, because it barely responds to
+packet loss. As mentioned in the results, we see no hard evidence that this
+causes other algorithms to reduce their speed further, but there are slight
+indications. For example in the test of DCTCP versus BBR, DCTCP reaches around
+0.2 to 0.6 megabytes per second. In the test of DCTCP versus BIC, this is 0.3
+to 0.7 megabytes per second. The difference is large nor statistically
+significant, but it warrants looking into. On the scale of 0.7 megabytes per
+second, a tenth of a megabyte does make a difference.
 
 \section{Future work}\label{sec:futurework}
-Running a mixed environment of Linux and Windows where Python is running introduces issues.
-We expect the call handling of Python in Windows encounters more delay than in Linux, but this requires more testing.
-The tests for CTCP might have to be performed again when full speeds can be reached by using the same settings.
-The CTCP tests are currently executed on Windows 10. This might also be performed on a Windows Server instance as it possibly handles the network in a different way.
 
-All tests described in this paper can be performed with more parallel connections, so that even when using a lot of parallel connections with two different congestion control algorithms, it is proven that the congestion control algorithms are fair.
+Running a mixed environment of Linux and Windows where Python is running
+introduced issues. We expect the send call of Python in Windows is implemented
+differently than in Linux, but this requires more testing. The tests for CTCP
+should be performed again when the reason for the discrepancy is known and can
+be mitigated. The results should also be verified using a Microsoft Windows
+Server instance.
+
+Our tests were done using three nodes in a switched network. To make sure our
+results hold in practice, future work could introduce a routed network with
+many more nodes.
+
+TODO Luc monitoring solution read up
 
@@
 
+TODO bbr turn on halfway through a test, does others' speed lessen?
+
 \section{Conclusion}\label{sec:conclusion}
+
 Our main research question is: {\it How can a datacenter provide a reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the network?}

### 2017-12-20 paper/paper.tex
-

@@
 
-The results have shown that the current default TCP congestion control algorithms are showing a fair behavior: The bandwidth is allocated equally among all nodes.
+The results have shown that the current default TCP congestion control algorithms are showing fair behavior: The bandwidth is allocated equally among all nodes.
 Therefore, currently no traffic control is needed, unless someone would create an unfair TCP congestion control algorithm.
@@
 For datacenters it works well, because it keeps the queue lengths to a minimum, assuring a packet arrives at its destination.
-\end{enumerate}
-
-% For reference, from the proposal:
-
-Our main research question is:
-{\it How can a datacenter provide an reliable and performant environment for
-customers when multiple TCP congestion control algorithms are used in the
-network?}
-
-\vspace{0.5cm}
-
-This main research question divides in five sub-questions:
-
-\begin{enumerate}
-	\item What TCP congestion algorithms are most commonly implemented?
-	\item How do these TCP congestion algorithms work?
-	\item How do these TCP congestion algorithms differ in behavior?
-	\item What is the impact on the availability and service for customers when using different, competing congestion algorithms?
-	\item What measures could be taken to split bandwidth equally among tenants?
 \end{enumerate}

### 2017-12-20 paper/paper.tex
Added sources

@@
 Therefore the packet loss and delay are applied to the incoming interfaces of the clients.
-Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables [TODO: SRC].
+Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables \cite{bbr-congestion}\cite{bic-tcp-congestion}\cite{compound-tcp-congestion}\cite{cubic-tcp-congestion}\cite{dctcp-congestion}.
 The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.

### 2017-12-20 paper/paper.tex
Checked todo: 55 mbps

@@
 time (i.e. a normal test is performed), CTCP's speed consistently drops. It
-then reaches around 55 megabits ($\pm$5), while the other algorithms reach at %TODO: check 55+-5
+then reaches around 57 megabits ($\pm$6.8), while the other algorithms reach at
 least 800 with good network conditions (8 milliseconds latency, 0\% loss).

### 2017-12-20 paper/paper.tex
Minor changes

@@
 time (i.e. a normal test is performed), CTCP's speed consistently drops. It
-then reaches around 57 megabits ($\pm$6.8), while the other algorithms reach at
+then reaches around 57 megabits ($\pm$6.8MiB/s), while the other algorithms reach at
 least 800 with good network conditions (8 milliseconds latency, 0\% loss).

### 2017-12-20 paper/paper.tex
Conclusion

--- 2017-12-20 paper/paper.tex
@@
 Therefore the packet loss and delay are applied to the incoming interfaces of the clients.
-Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables \cite{bbr-congestion}\cite{bic-tcp-congestion}\cite{compound-tcp-congestion}\cite{cubic-tcp-congestion}\cite{dctcp-congestion}.
+Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables [TODO: SRC].
 The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.
@@
 
-The results have shown that the current default TCP congestion control algorithms are showing fair behavior: The bandwidth is allocated equally among all nodes.
-Therefore, currently no traffic control is needed, unless someone would create an unfair TCP congestion control algorithm.
-This can, after more research, be detected by one of the monitoring solutions proposed in this paper.
-As for the tested TCP congestion control algorithms, there is no impact on the service and availability for customers when using those algorithms mixed in one environment.
+For the algorithms BIC, BBR, CUBIC and DCTCP, we conclude that they allocate
+bandwidth fairly and do not suppress each other. There is no impact on the
+service for customers when using those algorithms mixed in one environment.
 
-The BBR congestion control algorithm performs best when delay and loss occur of all TCP congestion control algorithms tested.
-DCTCP is only suitable when both the server and client are in the same datacenter, because the start up of the connection is slow when congestion in the network occurs.
-For datacenters it works well, because it keeps the queue lengths to a minimum, assuring a packet arrives at its destination.
-\end{enumerate}
+CTCP, the default algorithm for Microsoft Windows, shows inconclusive results
+and we can not draw any conclusions with regards to this algorithm.
+
+In our results, BBR shows better performance than any other algorithm when
+packet loss occurs. Finally, DCTCP is only suitable when both the server and
+client are in the same datacenter. There is a considerable ramp-up time when
+congestion occurs. Within datacenters this works well, because with no
+congestion it keeps the queue lengths to a minimum.
 

### 2017-12-20 paper/paper.tex

@@
 time (i.e. a normal test is performed), CTCP's speed consistently drops. It
-then reaches around 55 megabits ($\pm$5), while the other algorithms reach at %TODO: check 55+-5
+then reaches around 57 megabits ($\pm$6.8MiB/s), while the other algorithms reach at
 least 800 with good network conditions (8 milliseconds latency, 0\% loss).

### 2017-12-20 paper/paper.tex
Added BBR to do

@@
 
-TODO bbr turn on halfway through a test, does others' speed lessen?
+BBR appears to perform better in a congested network than the other tested protocols.
+In order to validate that BBR does not push away other TCP congestion algorithms a new testing method can be used.
+The data transfer with a TCP congestion algorithm is started and halfway the data transfer with BBR is started.
+This method shows whether a BBR data transfer has a large influence on the data transfer of other TCP congestion algorithms.
 

### 2017-12-20 paper/paper.tex
traffic management

--- 2017-12-20 paper/paper.tex
@@
 \subsection{Traffic Management}
-We propose two possible solutions to manage traffic from a multitenant datacenter to customers when unfair TCP congestion control algorithms are used: Traffic shaping based Monitoring and detection of unaccepted TCP congestion algorithms in the network.
+
+Our results show that traffic management is currently not necessary. However, a
+tenant might purposefully install an unfair algorithm. In this section we
+discuss different possible approaches a datacenter could take.
+
+We propose two possible solutions to manage traffic from a multitenant
+datacenter to customers when unfair TCP congestion control algorithms are used:
+traffic shaping based monitoring, and detection of unaccepted TCP congestion
+algorithms in the network.
 
 \subsubsection{Traffic shaping based Monitoring}
-With the use of a hypothetical TCP congestion control algorithm that is unfair to other TCP congestion control algorithms, the bandwidth allocation to one customer can be higher than for other customers using the default congestion control algorithms.
-For example: when a TCP congestion control algorithm shows an aggressive increase and keeps pushing data over the network, other fair TCP congestion algorithms will reduce the sending rate of the data transfer as packet loss occurs.
-One important condition for this unfair congestion control algorithm is that it should not be that agressive such that it behaves in a Denial of Service manner.
 
-To solve this issue, we propose a monitoring solution which monitors the bandwidth usage per server and controls the priority for the traffic sent by a server.The solution uses traffic shaping to control the priority of the traffic per server:
+With the use of a hypothetical TCP congestion control algorithm that is unfair towards other flows, the bandwidth allocation to one customer would be higher than to other customers.
+When a TCP flow shows an aggressive increase and keeps pushing data over the network, other (fair) algorithms will reduce the transmission rate as packet loss occurs.
+One important condition for this unfair congestion control algorithm is that it should not be agressive to the extent of causing a denial of service.
+
+To mitigate this, the bandwidth usage per server could be monitored, prioritizing the traffic sent by some servers. The solution uses traffic shaping to control the priority of the traffic per server:
+
 \begin{enumerate}
-	\item The monitoring solution detects periodically whether the bandwidth in the link is fully utilized.
-If this is the case, the next step will be initiated.
+	\item The monitoring solution measures periodically whether the bandwidth in the link is fully utilized.
+If this is the case, the next step will be performed.
 If not, it continues monitoring the link utilization.
@@
 
-In \cite{tcp-congestion-identification} a method to remotely identify used TCP congestion avoidance algorithms is described. The paper states that 
-they were able to identify all default TCP algorithms, but some non-default algorithms could not be identified.
-The tests performed are limited to remote web servers, so it requires more research to reliably implement such a method.
-If this method is deemed reliable, it can be useful to detect whether an unaccepted TCP congestion control algorithm is used as it will not match any of the default TCP algorithms used.
-Based on that the traffic of the unacceptable TCP congestion control algorithms can be shaped by lowering their priority.
+Peng Yang et al. describe a method to remotely identify used TCP congestion
+avoidance algorithms\cite{tcp-congestion-identification}. The paper states that
+they were able to identify all default TCP algorithms, but some non-default
+algorithms could not be identified. The tests performed are limited to remote
+web servers, so it requires more research to reliably implement such a method
+in a generic manner. If this method is deemed reliable, it can be useful to
+detect whether an unfair TCP congestion control algorithm is used as it will
+not match any of the default TCP algorithms used. Based on that the traffic of
+the unacceptable TCP congestion control algorithms can be shaped by lowering
+their priority.
 

### 2017-12-20 paper/paper.tex
paragraphs have a blank line in between instead of identation; removed unused appendix; changed "subsub:performing time-based automated tests"

@@
 \bibliography{references.bib}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{\baselineskip}
 
@@
 
-The tests will be performed using scripts written in Python [\ref{appendix:python}].
-The scripts either act as controller or as disciple.
-The controller is responsible for controlling the tests, by synchronizing the time, exchanging commands to run, modify or stop the tests and gathering data.
-The disciple is the listener and executes all task commands the controller sends.
-The commands include getting the hostname of a disciple, setting the TCP congestion algorithm of a disciple, configuring the time of a disciple, starting the data transfer and measurement tools on the disciple and closing the connection with the controller.
+The tests will be performed using scripts written in Python\cite{git-lsproj}.
+All nodes run a client, which performs actions as instructed by the controller.
+The controller is responsible for controlling the tests: synchronizing time
+among nodes, setting the TCP congestion algorithm of a node, and requesting one
+node to open a listening port and instructing another to start sending at a
+given time. Nodes report their measurements back to the controller.
 
@@
 
-TODO Luc monitoring solution read up
-
 In this paper we have proposed a monitoring solution, which checks for the deviation between the server bandwidth on a link and the average bandwidth used by servers on a link.
@@
 
-\appendix
-\section{Python}
-\label{appendix:python}
-SCRIPT
-
 \end{document}

### 2017-12-20 paper/paper.tex
spell check british english

@@
 been reported that some algorithms cause others to back off excessively.
-Tenants may, perhaps inadvertedly, cause inequal bandwidth distributions during
+Tenants may, perhaps inadvertently, cause unequal bandwidth distributions during
 times of congestion.
@@
 
-This report includes a performance and behavior analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network.
-Based on the results, we will recommend datacenters on measures to take to keep providing a reliable environment for customers when multiple congestion control algoritms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
-The main research question is: {\it How can a datacenter provide a reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the
+This report includes a performance and behaviour analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network.
+Based on the results, we will recommend datacenters on measures to take to keep providing a reliable environment for customers when multiple congestion control algorithms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
+The main research question is: {\it How can a datacenter provide a reliable and performing environment for customers when multiple TCP congestion control algorithms are used in the
 network?}
@@
 al.\cite{dctcp-congestion-original} and published as
-RFC8257\cite{dctcp-congestion} in 2017. DCTCP is optimized for datacenters and
+RFC8257\cite{dctcp-congestion} in 2017. DCTCP is optimised for datacenters and
 provides a high-burst tolerance, low latency, and high throughput when the
@@
 In order to select the most common TCP congestion algorithms desk research will be performed.
-This desk research will show which algorithms are used more often than other algoritms by looking for the default algorithms in Operating Systems.
+This desk research will show which algorithms are used more often than other algorithms by looking for the default algorithms in Operating Systems.
 Based on those results a list of TCP congestion algorithms will be composed.
@@
 This could be users downloading software, the bursts while streaming a video
-or movie, or large resources on a webpage.
+or movie, or large resources on a web page.
 
@@
 Therefore the packet loss and delay are applied to the incoming interfaces of the clients.
-Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables \cite{bbr-congestion}\cite{bic-tcp-congestion}\cite{compound-tcp-congestion}\cite{cubic-tcp-congestion}\cite{dctcp-congestion}.
+Packet loss and delay are chosen as the behaviour of most TCP algorithms depends on those variables \cite{bbr-congestion}\cite{bic-tcp-congestion}\cite{compound-tcp-congestion}\cite{cubic-tcp-congestion}\cite{dctcp-congestion}.
 The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.
@@
 
-The dataset of Verizon \cite{verizon-latency} is used to calculate the packet loss and delay used in this research.
-The dataset includes global information and shows averages per month over the year 2017.
+The data set of Verizon \cite{verizon-latency} is used to calculate the packet loss and delay used in this research.
+The data set includes global information and shows averages per month over the year 2017.
 The packet loss is calculated by taking the average of the Verizon Business Packet Delivery Statistics for Country Specific Metrics for each country.
@@
 All nodes run a client, which performs actions as instructed by the controller.
-The controller is responsible for controlling the tests: synchronizing time
+The controller is responsible for controlling the tests: synchronising time
 among nodes, setting the TCP congestion algorithm of a node, and requesting one
@@
 The goal is to find a mechanism that is capable of allocating bandwidth equally between customers of a datacenter.
-This is done by analyzing the results from the algorithm comparison and the initial desk research on traffic management.
+This is done by analysing the results from the algorithm comparison and the initial desk research on traffic management.
 
@@
 $BtlBw$ specifies the bandwidth at the slowest link in each direction and $RT_{prop}$ is the round-trip propagation time.
-$RT_{prop} = RTT - queuing delay - processing delay$, which results in the minimum amount of time for round trip propagation in case there are no queueing or processing delays.
+$RT_{prop} = RTT - queuing delay - processing delay$, which results in the minimum amount of time for round trip propagation in case there are no queuing or processing delays.
 Using those two parameters, the Bandwidth Delay Product (BDP) is calculated, which is the maximum possible amount of data being sent in a network: $BDP = BtlBw * RT_{prop}$.
@@
 $dwnd$ is calculated differently for several scenarios.
-If the network path is underutilized: $dwnd(t+1) = dwnd(t) + \alpha*dwnd(t)^{k} - 1$.
+If the network path is underutilised: $dwnd(t+1) = dwnd(t) + \alpha*dwnd(t)^{k} - 1$.
 When the network path is congested: $dwnd(t+1) = dwnd(t) - \eta*diff$.
 When packet loss occurs: $dwnd(t+1) = dwnd(t)(1-\beta) - cwnd/2$.
-$k$, $\alpha$, $\eta$ and $\beta$ are tunable parameters for optimization of the algorithm.
+$k$, $\alpha$, $\eta$ and $\beta$ are tunable parameters for optimisation of the algorithm.
 $\eta$ defines the rapidness of reduction of the window when congestion is detected.
@@
 	\item Simple Marking at the Switch.
-Based on the marking treshold $K$ a packet is marked with the CE flag.
+Based on the marking threshold $K$ a packet is marked with the CE flag.
 This only occurs when the queue occupancy is greater than K at arrival time.
-The marking treshold $K$ specifies the minimum value of queue occupancy.
+The marking threshold $K$ specifies the minimum value of queue occupancy.
 	\item ECN-Echo at receiver.
@@
 If $cwnd$ is larger than $W_{max}$, the convex profile is used to increase the window size according to the same formula as the concave profile.
-In both formulas $t$ is the time elapsed from the last window reducation.
+In both formulas $t$ is the time elapsed from the last window reduction.
 
@@
 When a TCP flow shows an aggressive increase and keeps pushing data over the network, other (fair) algorithms will reduce the transmission rate as packet loss occurs.
-One important condition for this unfair congestion control algorithm is that it should not be agressive to the extent of causing a denial of service.
+One important condition for this unfair congestion control algorithm is that it should not be aggressive to the extent of causing a denial of service.
 
-To mitigate this, the bandwidth usage per server could be monitored, prioritizing the traffic sent by some servers. The solution uses traffic shaping to control the priority of the traffic per server:
+To mitigate this, the bandwidth usage per server could be monitored, prioritising the traffic sent by some servers. The solution uses traffic shaping to control the priority of the traffic per server:
 
 \begin{enumerate}
-	\item The monitoring solution measures periodically whether the bandwidth in the link is fully utilized.
+	\item The monitoring solution measures periodically whether the bandwidth in the link is fully utilised.
 If this is the case, the next step will be performed.
-If not, it continues monitoring the link utilization.
+If not, it continues monitoring the link utilisation.
 	\item The monitoring solution calculates periodically per server whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage.
@@
 	\item The traffic shaping controller sets the priority for the servers to a lower value, so that the traffic coming from the servers that use $\alpha$ more than average bandwidth has less preference than traffic closer to the average bandwidth and lower.
-	\item The monitoring solution periodically checks if the bandwidth in the link is still fully utilized.
+	\item The monitoring solution periodically checks if the bandwidth in the link is still fully utilised.
 If this is not the case, the traffic shaping policy is reversed again by setting the priority to the default value.
@@
 \texttt{while(true)\{ socket.send(data); \}}, the bandwidth allocation did not
-change. We theorize that the difference is caused by something in the
+change. We theorise that the difference is caused by something in the
 \texttt{send} call. An alternative theory is that Microsoft Windows for
@@
 
-Our main research question is: {\it How can a datacenter provide a reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the network?}
+Our main research question is: {\it How can a datacenter provide a reliable and performing environment for customers when multiple TCP congestion control algorithms are used in the network?}
 

### 2017-12-20 paper/paper.tex
m. order of references

@@
 Therefore the packet loss and delay are applied to the incoming interfaces of the clients.
-Packet loss and delay are chosen as the behaviour of most TCP algorithms depends on those variables \cite{bbr-congestion}\cite{bic-tcp-congestion}\cite{compound-tcp-congestion}\cite{cubic-tcp-congestion}\cite{dctcp-congestion}.
+Packet loss and delay are chosen as the behaviour of most TCP algorithms depends on those variables \cite{bbr-congestion}\cite{dctcp-congestion}\cite{cubic-tcp-congestion}\cite{compound-tcp-congestion}\cite{bic-tcp-congestion}.
 The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.

### 2017-12-22 paper/paper.tex
Added some more clarification referring to the BBR paper

@@
 This report includes a performance and behaviour analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network.
+Mario Hock et al.
+describe in a paper that BBR shows unfair behaviour compared to the CUBIC congestion control algorithm \cite{bbr-congestion-comparison}.
+Therefore the fairness between TCP congestion control algorithms will also be included in the analysis.
 Based on the results, we will recommend datacenters on measures to take to keep providing a reliable environment for customers when multiple congestion control algorithms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
@@
 BBR is the best-performing algorithm in our set, because it barely responds to
-packet loss. As mentioned in the results, we see no hard evidence that this
+packet loss.
+However, we were unable to reproduce the results written in the paper by Mario Hock et al \cite{bbr-congestion-comparison}.
+We used multiple values to simulate packet loss as well as delay, which simulate having small buffers.
+BBR is a TCP congestion control protocol that reacts to the minimum RTT in combination with the Bottleneck Bandwidth.
+We have tested multiple delay values that appear in the real world.
+The delay values are directly related to the minimum RTT possible on the link.
+As mentioned in the results, we see no hard evidence that this
 causes other algorithms to reduce their speed further, but there are slight

### 2017-12-23 paper/paper.tex
Minor change in wording
Minor wording change

@@
 packet loss.
-However, we were unable to reproduce the results written in the paper by Mario Hock et al \cite{bbr-congestion-comparison}.
+However, we were unable to draw the same conclusions to our results as the conclusions written in the paper by Mario Hock et al \cite{bbr-congestion-comparison}.
 We used multiple values to simulate packet loss as well as delay, which simulate having small buffers.
@@
 
-For the algorithms BIC, BBR, CUBIC and DCTCP, we conclude that they allocate
+For the algorithms BIC, BBR, CUBIC and DCTCP, we conclude that, based on the setup we have used, they allocate
 bandwidth fairly and do not suppress each other. There is no impact on the

### 2017-12-23 paper/paper.tex
Fixed error in image and added explanation in results for some algorithms

@@
 See figures \ref{fig:fair-1} and \ref{fig:fair-2}.
+In figure \ref{fig:fair-1} we see that in the first test run, the increase of BIC is more aggressive than CUBIC's.
+This is related to the difference in increase functions used by both TCP congestion algorithms.
+In figure \ref{fig:fair-2} we also see that the increase of BIC is more aggressive than DCTCP's.
+BIC's increase function is characterised to be more aggressive and that can be seen in the graph as the increase line is more steep.
+The differences in those increase functions, however, do not show any implications on the fairness of the algorithms as the bandwidth allocation is near to equal for both algorithms.
 
@@
 			\centering
-			\includegraphics[width=0.9\textwidth]{figs/cubic-bic-8-0.png}
+			\includegraphics[width=0.9\textwidth]{figs/bic_vs_cubic_8_0.png}
 			\caption{CUBIC vs BIC with 8ms latency and 0\% loss}
@@
 with 1.2\% loss, as long as the delay is low (8 milliseconds in our case), it
-reaches 850$\pm$20 megabits per second. Other algorithms will reach much lower
+reaches 850$\pm$20 megabits per second. BBR is a delay based algorithm, so it mainly reacts based on delay. Other algorithms will reach much lower
 speeds (24 megabits is not uncommon), but we see no signs of BBR being the

### 2017-12-23 paper/paper.tex
Changed description of results more to be compared with the paper of Mario Hock et al

@@
 with 1.2\% loss, as long as the delay is low (8 milliseconds in our case), it
-reaches 850$\pm$20 megabits per second. BBR is a delay based algorithm, so it mainly reacts based on delay. Other algorithms will reach much lower
+reaches 850$\pm$20 megabits per second.
+A comparison between BBR and CUBIC is shown in figure \ref{fig:bbr-cubic}.
+We see in figure \ref{fig:bbr-cubic} that, compared to CUBIC, the bandwidth usage by the BBR data transfer is higher when 1.2\% packet loss is introduced.
+However, the link speed of 1Gbps is not fully utilized as is in the paper written by Mario Hock et al.
+\cite{bbr-congestion-comparison}.
+We also do not see the CUBIC data transfer responding to the BBR data transfer by reducing its speed in any way.
+In another test, where CUBIC is compared to BIC with the same parameters, we see CUBIC utilizing the same amount of bandwidth as in the BBR-CUBIC comparison.
+This is shown in figure \ref{fig:bic-cubic}.
+In figure \ref{fig:bbr-cubic-0} we show a comparison of a CUBIC data transfer with a BBR data transfer with parameters 8ms latency and 0\% packet loss.
+This time the total link speed is utilized and we see no unfairness between CUBIC and BBR.
+We therefore see no signs in our results of BBR being the cause of the CUBIC data transfer being slower.
+In our results other algorithms also reach much lower
 speeds (24 megabits is not uncommon), but we see no signs of BBR being the
@@
 is one of the algorithms being tested.
+
+\begin{figure}[H]
+	\makebox[\textwidth]{
+		\centering
+		\begin{minipage}{0.5\paperwidth}
+			\centering
+			\includegraphics[width=0.9\textwidth]{figs/cubic_vs_bbr_8_12.png}
+			\caption{CUBIC vs. BBR with 8ms latency and 1.2\% loss}
+			\label{fig:bbr-cubic}
+		\end{minipage}\hfill
+		\begin{minipage}{0.5\paperwidth}
+			\centering
+			\includegraphics[width=0.9\textwidth]{figs/cubic_vs_bic_8_12.png}
+			\caption{CUBIC vs BIC with 8ms latency and 1.2\% loss}
+			\label{fig:bic-cubic}
+		\end{minipage}
+	}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+		\includegraphics[scale=0.4]{figs/cubic_vs_bbr_8_0.png}
+		\caption{CUBIC vs. BBR with 8ms latency and 0\% loss}
+	\label{fig:bbr-cubic-0}
+\end{figure}
 

### 2017-12-24 paper/paper.tex
Paper: formatting, minor fixes, and release 1.0

@@
 \usepackage[utf8]{inputenc}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
@@
 \bibliography{references.bib}
+\usepackage[top=3.5cm, bottom=3.5cm, left=4.3cm, right=4.3cm]{geometry}
 
@@
 \title{TCP Congestion Algorithms In Datacenters \\
-	\vspace{0.3cm}
-	{\large LS Project}
+	{\large Large Systems $\cdot$ Project Report}
 }
@@
 
-\section{Abstract}
+\section*{Abstract}
 
@@
 
+\newpage
 \section{Introduction}
@@
 
-This report includes a performance and behaviour analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network.
-Mario Hock et al.
-describe in a paper that BBR shows unfair behaviour compared to the CUBIC congestion control algorithm \cite{bbr-congestion-comparison}.
-Therefore the fairness between TCP congestion control algorithms will also be included in the analysis.
-Based on the results, we will recommend datacenters on measures to take to keep providing a reliable environment for customers when multiple congestion control algorithms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms.
-The main research question is: {\it How can a datacenter provide a reliable and performing environment for customers when multiple TCP congestion control algorithms are used in the
+This report includes a performance and behaviour analysis of different TCP
+congestion control algorithms in a (simulated) multitenant datacenter network.
+Mario Hock et al. describe in a paper that BBR shows unfair behaviour when
+compared to the CUBIC congestion control
+algorithm\cite{bbr-congestion-comparison}. Therefore the fairness between
+congestion control algorithms will also be included in the analysis. Based on
+the results, we will recommend datacenters on measures to take to keep
+providing a reliable environment for customers when multiple congestion control
+algorithms are active in the network, for example using Quality of Service or
+prohibiting certain TCP congestion control algorithms. The main research question
+is: {\it How can a datacenter provide a reliable and performing environment for
+customers when multiple TCP congestion control algorithms are used in the
 network?}
@@
 can be found in \ref{sub:experiment-results}. We discuss our results in section
-\ref{sec:discussion} and draw conclusions in section \ref{sec:conclusion}.
+\ref{sec:discussion}, mention possible future work in section
+\ref{sec:futurework}, and finally draw conclusions in section \ref{sec:conclusion}.
 
@@
 
-During this research, performance comparison tests of various TCP congestion algorithms will be performed.
-The most common TCP congestion algorithms will be determined through desk research.
-The tests will be conducted in an isolated environment.
-Based on the results of those tests, recommendations will be given related to the measures a datacenter has to take to keep the traffic equally allocated over customers.
+During this research, performance comparison tests of various TCP congestion
+control algorithms will be performed. The most common TCP congestion algorithms will be
+determined through desk research. The tests will be conducted in an isolated
+environment. Based on the results of those tests, recommendations will be given
+related to the measures a datacenter has to take to keep the traffic equally
+allocated over customers.
 
@@
 
-In order to select the most common TCP congestion algorithms desk research will be performed.
-This desk research will show which algorithms are used more often than other algorithms by looking for the default algorithms in Operating Systems.
-Based on those results a list of TCP congestion algorithms will be composed.
+In order to select the most common TCP congestion algorithms desk research will
+be performed. This desk research will show which algorithms are used more often
+than other algorithms by looking for the default algorithms in Operating
+Systems. Based on those results, a list of TCP congestion algorithms will be
+composed.
 
@@
 We will create a bottleneck on the path to the user to see how different
-congestion algorithms handle this. The connection will be full duplex and have
-symmetrical speeds, to avoid any congestion in sending acknowledgements.
-Because the server is sending the large volume of data and the client is
-responding with the much smaller acknowledgements, the server will be the one
-applying the congestion control. Therefore we vary the algorithm settings of
-the server and not the client. The client will always be a GNU/Linux machine
-using CUBIC, but we assume (due to the aforementioned reasons) that this does
-not matter.
+congestion control algorithms handle this. The connection will be full-duplex
+and have symmetrical speeds, to avoid any congestion in sending
+acknowledgements. Because the server is sending the large volume of data and
+the client is responding with the much smaller acknowledgements, the server
+will be the one applying congestion control. Therefore we vary the algorithm
+settings of the server and not the client. The client will always be a
+GNU/Linux machine using CUBIC, but we assume (due to the aforementioned
+reasons) that this does not matter.
 
@@
 We use three physical hosts as servers: two running Ubuntu 17.04 and one
-running Microsoft Windows 10. In each experiment, only two hosts will be used.
+running Microsoft Windows 10. In each experiment, only two hosts will be active.
 The Windows system is used because there is no working CTCP
@@
 
-We will test the TCP congestion algorithm for fairness in sharing bandwidth
+We will test the congestion algorithm for fairness in sharing bandwidth
 with other TCP connections using other congestion algorithms. In order to test
@@
 The tests conducted with the algorithms have two environment variables: packet loss and delay.
-Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions \cite{linux-netem}.
+Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions\cite{linux-netem}.
 \texttt{netem} is applied to the incoming side of the interfaces of the clients to emulate the same packet loss and delay.
@@
 Therefore the packet loss and delay are applied to the incoming interfaces of the clients.
-Packet loss and delay are chosen as the behaviour of most TCP algorithms depends on those variables \cite{bbr-congestion}\cite{dctcp-congestion}\cite{cubic-tcp-congestion}\cite{compound-tcp-congestion}\cite{bic-tcp-congestion}.
-The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.
+Packet loss and delay are chosen as the behaviour of most TCP algorithms depends on those variables\cite{bbr-congestion}\cite{dctcp-congestion}\cite{cubic-tcp-congestion}\cite{compound-tcp-congestion}\cite{bic-tcp-congestion}.
+The values for the environment variables are chosen based on measurements performed by Verizon\cite{verizon-latency}.
 
@@
 
-The data set of Verizon \cite{verizon-latency} is used to calculate the packet loss and delay used in this research.
+The data set of Verizon\cite{verizon-latency} is used to calculate the packet loss and delay used in this research.
 The data set includes global information and shows averages per month over the year 2017.
@@
 
-The country numbers in figure \ref{fig:packet-loss-chart} are only used to show the distribution of the average packet loss.
-The maximum packet loss shown in figure \ref{fig:packet-loss-chart} equals 0.6\% and the minimum equals 0\%.
-However, in case of situations where more packet loss occurs than usual, the double value of the maximum is used: 1.2\%.
-The values for packet loss in this comparison set-up were determined based on the distribution shown in figure \ref{fig:packet-loss-chart} and the manually set maximum of 1.2\%.
-The values are shown in table \ref{table:test-packetloss}.
+The country numbers in figure \ref{fig:packet-loss-chart} are only used to show
+the distribution of the average packet loss. The maximum average packet loss
+shown in figure \ref{fig:packet-loss-chart} equals 0.6\% and the minimum equals
+0\%. However, to account for situations where more packet loss occurs than usual,
+the double value of the maximum is used: 1.2\%. The values for packet loss in
+this comparison set-up were determined based on the distribution shown in
+figure \ref{fig:packet-loss-chart} and the manually set maximum of 1.2\%. The
+values are shown in table \ref{table:test-packetloss}.
 
@@
 	\centering
-	\caption{Packet loss percentages used for the comparison setup}
 	\begin{tabular}[H]{ | l |}
@@
 	\end{tabular}
+	\caption{Packet loss percentages used for the comparison setup}
 	\label{table:test-packetloss}
@@
 
-The delay is calculated by taking the average of the Verizon Business Latency Statistics for Country Specific Metrics for each country.
-Based on the ascending sorted values a bar chart was created.
-This bar chart is shown in figure \ref{fig:verizon-delay-chart}
+\vspace{0.9cm}
+
+The delay is calculated by taking the average of the Verizon Business Latency
+Statistics for Country Specific Metrics for each country. Based on the
+ascending sorted values a bar chart was created. This chart is shown in figure
+\ref{fig:verizon-delay-chart}.
 
@@
 
-The country numbers in figure \ref{fig:verizon-delay-chart} are only used to show the distribution of the average delays.
-The maximum delay shown in figure \ref{fig:verizon-delay-chart} equals 290ms and the minimum equals 8ms (both rounded).
-The values for delay in this comparison set-up were determined based on the distribution shown in figure \ref{fig:verizon-delay-chart}.
-The values are shown in table \ref{table:test-delay}.
+The country numbers in figure \ref{fig:verizon-delay-chart} are only used to
+show the distribution of the average delays. The maximum delay shown in figure
+\ref{fig:verizon-delay-chart} equals 290ms and the minimum equals 8ms (both
+rounded). The values for delay in this comparison set-up were determined based
+on the distribution shown in figure \ref{fig:verizon-delay-chart}. The values
+are shown in table \ref{table:test-delay}.
 
@@
 	\end{tabular}
+	\caption{Packet loss percentages used for the comparison setup}
 	\label{table:test-delay}
@@
 The most common TCP congestion algorithms are CTCP, CUBIC and BIC.
-CTCP is the default congestion algorithm in Microsoft Windows Server 2008 and newer \cite{cubic-kernel-version}.
-CUBIC is the default congestion algorithm in Linux distributions with kernel versions 2.6.19 and higher \cite{cubic-kernel-version}.
-BIC was the default congestion algorithm in earlier Linux distributions from kernel version 2.6.8 until 2.6.19 \cite{bic-kernel-version} \cite{cubic-kernel-version}.
-DCTCP and BBR will also be tested as those protocols are new and upcoming---they are included in Linux since 4.9 (December 2016) \cite{linux-bbr}.
-DCTCP is specifically designed for datacenters \cite{dctcp-congestion} and BBR is a new algorithm meant for general use \cite{bbr-congestion}.
+CTCP is the default congestion algorithm in Microsoft Windows Server 2008 and newer\cite{cubic-kernel-version}.
+CUBIC is the default congestion algorithm in Linux distributions with kernel versions 2.6.19 and higher\cite{cubic-kernel-version}.
+BIC was the default congestion algorithm in earlier Linux distributions from kernel version 2.6.8 until 2.6.19\cite{bic-kernel-version}\cite{cubic-kernel-version}.
+DCTCP and BBR will also be tested as those protocols are new and upcoming---they are included in Linux since 4.9 (December 2016)\cite{linux-bbr}.
+DCTCP is specifically designed for datacenters\cite{dctcp-congestion} and BBR is a new algorithm meant for general use\cite{bbr-congestion}.
 
@@
 
-The BBR algorithm depends on parameters $BtlBw$ and $RT_{prop}$ \cite{bbr-congestion}.
+The BBR algorithm depends on parameters $BtlBw$ and $RT_{prop}$\cite{bbr-congestion}.
 Those parameters are estimates defined by the BBR algorithm.
@@
 If the amount of data in transit is equal to $BtlBw$ the delivery rate can not go up anymore.
-The $RTT$ can never be lower than $RT_{prop}$...
-explain more on delay perhaps?.
+The $RTT$ can never be lower than $RT_{prop}$.
 
-The CTCP algorithm is a combination of a loss-based and delay-based congestion protocol \cite{compound-tcp-congestion}.
+The CTCP algorithm is a combination of a loss-based and delay-based congestion protocol\cite{compound-tcp-congestion}.
 For the loss-based component, the variable conventional congestion window, $cwnd$, was introduced.
@@
 
-The DCTCP algorithm consists of three components \cite{dctcp-congestion-original}:
+The DCTCP algorithm\cite{dctcp-congestion-original} consists of three components:
 
@@
 
-The BIC algorithm controls its windows based on the size of windows \cite{bic-tcp-congestion}. It consists of two components:
+The BIC algorithm controls its windows based on the size of windows\cite{bic-tcp-congestion}. It consists of two components:
 
@@
 
-The CUBIC algorithm is a successor of the BIC algorithm \cite{cubic-tcp-congestion}.
+The CUBIC algorithm is a successor of the BIC algorithm\cite{cubic-tcp-congestion}.
 In case of a loss event $W_{max}$, which is the maximum window size, is registered.
@@
 			\centering
-			\includegraphics[width=0.9\textwidth]{figs/bic_vs_cubic_8_0.png}
+			\includegraphics[width=0.9\textwidth]{figs/cubic_vs_bic_8_0.png}
 			\caption{CUBIC vs BIC with 8ms latency and 0\% loss}
@@
 	\centering
-		\includegraphics[scale=0.4]{figs/ctcp-bic-8-0.png}
+		\includegraphics[scale=0.55]{figs/ctcp-bic-8-0.png}
 		\caption{CTCP vs. BIC with 8ms latency and 0\% loss}
@@
 
+%\newgeometry{top=1cm,bottom=1.5cm}
+
 BBR is unique in that it barely responds to packet loss but only to delay: even
@@
 We see in figure \ref{fig:bbr-cubic} that, compared to CUBIC, the bandwidth usage by the BBR data transfer is higher when 1.2\% packet loss is introduced.
-However, the link speed of 1Gbps is not fully utilized as is in the paper written by Mario Hock et al.
-\cite{bbr-congestion-comparison}.
+However, the link speed of 1Gbps is not fully utilized as is in the paper written by Mario Hock et al.\cite{bbr-congestion-comparison}.
 We also do not see the CUBIC data transfer responding to the BBR data transfer by reducing its speed in any way.
@@
 	\centering
-		\includegraphics[scale=0.4]{figs/cubic_vs_bbr_8_0.png}
+		\includegraphics[scale=0.6]{figs/cubic_vs_bbr_8_0.png}
 		\caption{CUBIC vs. BBR with 8ms latency and 0\% loss}
@@
 The full set of results is available online\cite{git-lsproj}.
+
+%\restoregeometry
 
@@
 
-With the use of a hypothetical TCP congestion control algorithm that is unfair towards other flows, the bandwidth allocation to one customer would be higher than to other customers.
+With the use of a hypothetical congestion control algorithm which is unfair towards other flows, the bandwidth allocation to one customer would be higher than to other customers.
 When a TCP flow shows an aggressive increase and keeps pushing data over the network, other (fair) algorithms will reduce the transmission rate as packet loss occurs.
@@
 
-To mitigate this, the bandwidth usage per server could be monitored, prioritising the traffic sent by some servers. The solution uses traffic shaping to control the priority of the traffic per server:
+To mitigate this, the bandwidth usage per server could be monitored, prioritising the traffic sent by some servers. The monitor uses traffic shaping to control the priority of the traffic per server:
 
 \begin{enumerate}
-	\item The monitoring solution measures periodically whether the bandwidth in the link is fully utilised.
+	\item The monitor measures periodically whether the bandwidth in the link is fully utilised.
 If this is the case, the next step will be performed.
 If not, it continues monitoring the link utilisation.
-	\item The monitoring solution calculates periodically per server whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage.
-$\alpha$ is deviated bandwidth and is set according to an acceptable deviation value.
+	\item The monitor calculates periodically, per server, whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage.
+$\alpha$ is the deviated bandwidth and is set according to an acceptable deviation value.
 The following formula is used to calculate the standard deviation: $s_{i} = \sqrt{\frac{\sum_{i=1}^N (x_i - \overline{x})^2}{N-1} }$.
@@
 If $s_{i} > \alpha$, the priority will be changed to a lower value.
-	\item The monitoring solution provides information of the servers where the priority will be set to a lower value to the traffic shaping controller.
+	\item The monitor provides information of the servers where the priority will be set to a lower value to the traffic shaping controller.
 	\item The traffic shaping controller sets the priority for the servers to a lower value, so that the traffic coming from the servers that use $\alpha$ more than average bandwidth has less preference than traffic closer to the average bandwidth and lower.
-	\item The monitoring solution periodically checks if the bandwidth in the link is still fully utilised.
+	\item The monitor periodically checks if the bandwidth in the link is still fully utilised.
 If this is not the case, the traffic shaping policy is reversed again by setting the priority to the default value.
@@
 
-Peng Yang et al. describe a method to remotely identify used TCP congestion
-avoidance algorithms\cite{tcp-congestion-identification}. The paper states that
-they were able to identify all default TCP algorithms, but some non-default
-algorithms could not be identified. The tests performed are limited to remote
-web servers, so it requires more research to reliably implement such a method
-in a generic manner. If this method is deemed reliable, it can be useful to
-detect whether an unfair TCP congestion control algorithm is used as it will
-not match any of the default TCP algorithms used. Based on that the traffic of
-the unacceptable TCP congestion control algorithms can be shaped by lowering
-their priority.
+Peng Yang et al.\cite{tcp-congestion-identification} describe a method to
+remotely identify used TCP congestion avoidance algorithms. The paper states
+that they were able to identify all default TCP algorithms, but some
+non-default algorithms could not be identified. The tests performed are limited
+to remote web servers, so it requires adaption to implement such a method in a
+generic manner. If this method is deemed reliable, it can be useful to detect
+whether an unfair algorithm is used as it will not match
+any of the default algorithms used. Based on that, the traffic of the
+unacceptable TCP congestion control algorithms can be shaped by lowering their
+priority.
 
@@
 
-Without any concurrent flows, Windows does reach gigabit speed. On different
-hardware, it showed the same, recessive behaviour. We think this might be
-caused by the way Python socket operations are handled in Windows. When
-repeating the test with \texttt{iperf}, which is written in C, this behaviour
-no longer occurs. Even when using the Python equivalent of
+Without any concurrent flows, Windows does reach the full link speed (1Gbps).
+On different hardware, it showed the same, recessive behaviour. We
+think this might be caused by the way Python socket operations are handled in
+Windows. When repeating the test with \texttt{iperf}, which is written in C,
+this behaviour no longer occurs. Even when using the Python equivalent of
 \texttt{while(true)\{ socket.send(data); \}}, the bandwidth allocation did not
@@
 BBR is the best-performing algorithm in our set, because it barely responds to
-packet loss.
-However, we were unable to draw the same conclusions to our results as the conclusions written in the paper by Mario Hock et al \cite{bbr-congestion-comparison}.
-We used multiple values to simulate packet loss as well as delay, which simulate having small buffers.
-BBR is a TCP congestion control protocol that reacts to the minimum RTT in combination with the Bottleneck Bandwidth.
-We have tested multiple delay values that appear in the real world.
-The delay values are directly related to the minimum RTT possible on the link.
-As mentioned in the results, we see no hard evidence that this
-causes other algorithms to reduce their speed further, but there are slight
-indications. For example in the test of DCTCP versus BBR, DCTCP reaches around
-0.2 to 0.6 megabytes per second. In the test of DCTCP versus BIC, this is 0.3
-to 0.7 megabytes per second. The difference is large nor statistically
+packet loss. However, we were unable to draw the same conclusions to our
+results as the conclusions written in the paper by Mario Hock et
+al\cite{bbr-congestion-comparison}. We used multiple values to simulate packet
+loss as well as delay, which simulate having small buffers. BBR is a TCP
+congestion control protocol that reacts to the minimum RTT in combination with
+the Bottleneck Bandwidth. We have tested multiple delay values that appear in
+the real world. The delay values are directly related to the minimum RTT
+possible on the link. As mentioned in the results, we see no hard evidence that
+this causes other algorithms to reduce their speed further, but there are
+slight indications. For example in the test of DCTCP versus BBR, DCTCP reaches
+around 0.2 to 0.6 megabytes per second. In the test of DCTCP versus BIC, this
+is 0.3 to 0.7 megabytes per second. The difference is large nor statistically
 significant, but it warrants looking into. On the scale of 0.7 megabytes per
@@
 
-Running a mixed environment of Linux and Windows where Python is running
-introduced issues. We expect the send call of Python in Windows is implemented
-differently than in Linux, but this requires more testing. The tests for CTCP
-should be performed again when the reason for the discrepancy is known and can
-be mitigated. The results should also be verified using a Microsoft Windows
-Server instance.
+Running a mixed environment of Linux and Windows introduced issues, at least in
+combination with Python as benchmark system. We expect that in Windows, the
+Python send call on sockets is implemented differently than in Linux, but this
+requires more testing. The tests for CTCP should be performed again when the
+reason for the discrepancy is known and can be mitigated. The results should
+also be verified using a Microsoft Windows Server instance.
 
@@
 
-BBR appears to perform better in a congested network than the other tested protocols.
-In order to validate that BBR does not push away other TCP congestion algorithms a new testing method can be used.
-The data transfer with a TCP congestion algorithm is started and halfway the data transfer with BBR is started.
-This method shows whether a BBR data transfer has a large influence on the data transfer of other TCP congestion algorithms.
+BBR performs better in a congested network than any other tested algorithm.
+While we did not see clear signs of other algorithms responding to BBR's higher
+bandwidth share, this should be tested further to be certain. We suggest
+starting a transfer with one algorithm under the desired network circumstances,
+determine its bandwidth, and then start a parallel flow using BBR. If the
+bandwidth of the original flow does not decrease, BBR is indeed fair towards
+this algorithm.
 
@@
 packet loss occurs. Finally, DCTCP is only suitable when both the server and
-client are in the same datacenter. There is a considerable ramp-up time when
-congestion occurs. Within datacenters this works well, because with no
-congestion it keeps the queue lengths to a minimum.
+client are in the same datacenter: there is a considerable ramp-up time when
+congestion occurs. Within datacenters DCTCP works well, because without
+congestion it keeps queue lengths to a minimum.
 

### 2017-12-24 paper/paper.tex
paper: consistent "vs." usage and fixing our names in the references

@@
 \usepackage[utf8]{inputenc}
-\usepackage[hidelinks]{hyperref}
+\usepackage[hidelinks, pdfauthor={Luc Gommans and Rick van Gorp}, pdftitle={TCP Congestion Algorithms in Datacenters}]{hyperref}
 \usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
@@
 			\includegraphics[width=0.9\textwidth]{figs/cubic_vs_bic_8_0.png}
-			\caption{CUBIC vs BIC with 8ms latency and 0\% loss}
+			\caption{CUBIC vs. BIC with 8ms latency and 0\% loss}
 			\label{fig:fair-1}
@@
 			\includegraphics[width=0.9\textwidth]{figs/dctcp-bic-8-0.png}
-			\caption{DCTCP vs BIC with 8ms latency and 0\% loss}
+			\caption{DCTCP vs. BIC with 8ms latency and 0\% loss}
 			\label{fig:fair-2}
@@
 			\includegraphics[width=0.9\textwidth]{figs/cubic_vs_bic_8_12.png}
-			\caption{CUBIC vs BIC with 8ms latency and 1.2\% loss}
+			\caption{CUBIC vs. BIC with 8ms latency and 1.2\% loss}
 			\label{fig:bic-cubic}
